### 2018-12-03 notes/Paper.tex
added notes for paper

\documentclass[prd,preprintnumbers,twocolumn,eqsecnum,floatfix,letter]{revtex4}
\usepackage{color}
\usepackage{calc}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{amssymb,amsmath}
\usepackage{tensor}
\usepackage{bm}
\usepackage{times}
\usepackage[varg]{txfonts}
\usepackage[colorlinks, pdfborder={0 0 0}]{hyperref}
\definecolor{LinkColor}{rgb}{0.75, 0, 0}
\definecolor{CiteColor}{rgb}{0, 0.5, 0.5}
\definecolor{UrlColor}{rgb}{0, 0, 0.75}
\hypersetup{linkcolor=LinkColor}
\hypersetup{citecolor=CiteColor}
\hypersetup{urlcolor=UrlColor}
\maxdeadcycles=1000
\allowdisplaybreaks
\textwidth 7.5 in 
\hoffset -1 cm 
\newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
\newcommand{\ajith}[1]{\textcolor{red}{\textit{Ajith:#1}}}
\newcommand{\ashok}[1]{\textcolor{cyan}{\textit{Ashok:#1}}}

\begin{document}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ber}{\begin{eqnarray}}
\newcommand{\eer}{\end{eqnarray}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\newcommand{\etal}{\emph{et al.}}

\newcommand{\Sl}{S_\ell}
\newcommand{\Sigmal}{\Sigma_\ell}
\newcommand{\Flux}{\mathcal{F}}
\newcommand{\LNh}{\hat{\mathbf{L}}_N}
\newcommand{\LN}{\mathbf{L}_N}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\bJ}{\mathbf{J}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\flow}{f_\mathrm{low}}
\newcommand{\fcut}{f_\mathrm{cut}}

\newcommand{\bchi}{\bm{\chi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bchia}{\bm{\chi}_a}
\newcommand{\bchis}{\bm{\chi}_s}
\newcommand{\chis}{\chi_s}
%\newcommand{\bchi}{\mathbf{\chi}}
\newcommand{\chia}{\chi_a}
\newcommand{\chiadL}{\bchia \cdot \LNh}
\newcommand{\chisdL}{\bchis \cdot \LNh}
\newcommand{\chisSqr}{\bchis^2}
\newcommand{\chiaSqr}{\bchia^2}
\newcommand{\chisDchia}{\bchis \cdot \bchia}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\pc}{{+,\times}}

\newcommand{\LIGO}{LIGO Laboratory, California Institute of Technology, 
Pasadena, CA 91125, USA}
\newcommand{\CIT}{Theoretical Astrophysics, California Institute of
Technology, Pasadena, CA 91125, USA}
\newcommand{\UWM}{UWM}

\title{Notes on the SpinTaylorF2 project}
\author{P.~Ajith}\email{ajith@icts.res.in}
\affiliation{International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bangalore 560012, India}
\author{Ashok Choudhary}
\affiliation{International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bangalore 560012, India}

\begin{abstract}
\end{abstract}

\maketitle

\section{Computing the spin and angular momentum vectors in the frequency domain}

\subsection{Performing the evolution in Cartesian coordinates}

The spin angular momentum vectors $\bS_i (v)$ and the unit vector $ \LNh (v)$ along the Newtonian angular momentum the frequency domain can be computed by using the evolution equations: 
\begin{eqnarray}
\frac{d \bS_i}{dv} & = & \frac{d \bS_i}{dt}  \frac{dt}{dv}, \\
\frac{d \LNh}{dv}  & = & \frac{d \LNh}{dt} \frac{dt}{dv},  
\label{eq:precessionEqFreqDom}
\end{eqnarray}
%
where 
\begin{equation}
\frac{d \LNh}{dt}  = - \frac{1}{||\mathbf{L}||} \, \frac{d}{dt} (\bS_1 + \bS_2),
\label{eq:orbitPrec}
\end{equation}
% 
%
\begin{equation}
\frac{d \bS_i}{dt}  =  \bm{\Omega}_i \times \bS_i \,,~ i = 1,2, 
\label{eq:spinPrec}
\end{equation}
%
and 
% 
\begin{widetext}
\begin{align}\label{eq:EbyF}
\frac{dt}{dv} & = \frac{5m}{32 \, v^9  \, \eta} \left\{1 + v^2 \left[\frac{11 \eta }{4}+\frac{743}{336} \right] 
    + v^3 \left[\frac{113}{12} \left(\biggl( 1-\frac{76\eta}{113}\biggr) \, \chisdL + \delta \, \chiadL \right)-4 \pi \right] \right. \nonumber \\ 
    &\quad + v^4 \left[(\chiadL)^2 \left(30 \eta -\frac{719}{96}\right)-\frac{719 \chiadL \, \chisdL \delta
   }{48}+\chiaSqr \left(\frac{233}{96}-10 \eta \right)+\frac{233 \chisDchia \delta}{48} \right. \nonumber\\
   &\qquad \left. +(\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
   }{24}+\frac{233}{96}\right)+\frac{617 \eta ^2}{144}+\frac{5429 \eta }{1008}+\frac{3058673}{1016064} \right] \nonumber \\ 
    &\quad + v^5 \left[\chiadL \delta  \left(\frac{7 \eta }{2}+\frac{146597}{2016}\right)+\chisdL \left(-\frac{17 \eta
   ^2}{2}-\frac{1213 \eta }{18}+\frac{146597}{2016}\right)+\frac{13 \pi  \eta }{8}-\frac{7729 \pi }{672} \right]  \nonumber\\ 
    &\quad + v^6 \left[\frac{1712 \gamma_E }{105}+\frac{25565 \eta ^3}{5184}-\frac{15211 \eta
   ^2}{6912}-\frac{451 \pi ^2 \eta }{48}+\frac{3147553127 \eta }{12192768}+\frac{32 \pi
   ^2}{3}-\frac{10817850546611}{93884313600}+\frac{1712 \ln (4v)}{105} \right]  \nonumber\\ 
    &\quad \left. + v^7 \left[\frac{14809 \pi  \eta ^2}{3024}-\frac{75703 \pi  \eta }{6048}-\frac{15419335 \pi }{1016064} \right] \right\}. 
\end{align}
\end{widetext}
% 
(\ajith{Need to add the BH absorption term above}). The (orbit averaged) precession frequencies $\bm{\Omega}_i$ appearing in Eq.~(\ref{eq:spinPrec}) are given by 
\begin{align}
\label{eq:Omega}
\bm{\Omega}_1 & =  \frac{v^5}{m} \left\{\left(\frac{3}{4}+\frac{\eta}{2}-\frac{3\delta}{4}\right) \LNh \right. \nonumber \\
     & \quad +  \frac{v}{2m^2} \left[-3 \, (\bS_2 + q \, \bS_1).\LNh ~ \LNh + \bS_2 \right]  \nonumber \\
     & \quad +  \left. v^2 \left(\frac{9}{16} + \frac{5\eta}{4} - \frac{\eta^2}{24} - \frac{9\delta}{16} + \frac{5\delta \eta}{8} \right)  \LNh  \right\}, \nonumber \\
%
\bm{\Omega}_2 & =  \frac{v^5}{m} \left\{\left(\frac{3}{4}+\frac{\eta}{2}+\frac{3\delta}{4}\right) \LNh \right. \nonumber \\
     & \quad +  \frac{v}{2m^2} \left[-3 \, (\bS_1 + q^{-1} \, \bS_2).\LNh ~ \LNh + \bS_1 \right]  \nonumber \\
     & \quad +  \left. v^2 \left(\frac{9}{16} + \frac{5\eta}{4} - \frac{\eta^2}{24} + \frac{9\delta}{16} - \frac{5\delta \eta}{8} \right)  \LNh  \right\}. 
\end{align}

%
Above, $v \equiv (m \omega)^{1/3}$ is the PN expansion parameter, $\omega$ is the orbital frequency, $m \equiv m_1 + m_2$ is the total mass of the binary, $q \equiv m_2/m_1$ is the mass ratio, $\eta \equiv m_1 m_2/m^2$ is the symmetric mass ratio, $\delta \equiv (m_1-m_2)/m$ the asymmetric mass ratio, $\gamma_E$ the Euler's constant, and $\LNh$ the unit vector along the Newtonian orbital angular momentum $\LN \equiv m^2\eta/v ~ \LNh$. The spin variables $\bchis$ and $\bchia$ are related to the (dimensionless) spin vectors of the binary as: 
%
\begin{equation}
\bchis = (\bchi_1 + \bchi_2)/2, ~~~ \bchia = (\bchi_1 - \bchi_2)/2,
\end{equation}
%
where $\bchi_i \equiv \bS_i/m_i^2$.

\subsection{Performing the evolution in spherical polar coordinates}

The time evolution equation of $i \equiv \arccos L_{Nz}$ and $ \alpha \equiv \arctan (L_{Ny}/L_{Nx})$ are given by:
\begin{equation}
\frac{d\iota}{dt} = \frac{-1}{\|L\| \sin\iota}\sum\limits_{k = 1}^{2}m^2_{k}\| \boldsymbol{\chi}_{k}\|\sin\theta_{s_{k}}\frac{d\theta_{s_{k}}}{dt}
\end{equation}  
and
\begin{widetext}
\begin{align}
\begin{split}
\frac{d\alpha}{dt} =& \frac{\cos\alpha}{\sin\iota}\Bigg(\frac{-1}{\|\boldsymbol{L}\|}\sum\limits_{k = 1}^{2} m^{2}_{k}\|\boldsymbol{{\chi}}_{k}\|\left[\frac{d\theta_{S_{k}}}{dt}\cos\theta_{s_{k}}\sin\phi_{s_{k}} + \frac{d\phi_{S_{k}}}{dt}\sin\theta_{s_{k}}\cos\phi_{s_{k}}\right]\Bigg) \\ 
- & \frac{\sin\alpha}{\sin\iota}\Bigg(\frac{-1}{\|\boldsymbol{L}\|}\sum\limits_{k = 1}^{2} m^{2}_{k}\|\boldsymbol{{\chi}}_{k}\|\left[\frac{d\theta_{S_{k}}}{dt}\cos\theta_{s_{k}}\cos\phi_{s_{k}} - \frac{d\phi_{S_{k}}}{dt}\sin\theta_{s_{k}}\sin\phi_{s_{k}}\right]\Bigg)
\end{split}
\end{align}
 \end{widetext}

 The angles corresponding to spin vectors($S_{1}$ and $S_{2}$)can also be written as:
 \begin{eqnarray}
 \frac{d\theta_{s}}{dt} = \frac{1}{\|\boldsymbol{\chi}\|\cos\theta_{s}}\left[
 \cos\phi_{s}(\Omega_{y}\chi_{z}-\Omega_{z}\chi_{y}) + \sin\phi_{s}(\Omega_{z}\chi_{x}-\Omega_{x}\chi_{z}) \right], \nonumber \\ 
  \frac{d\phi_{s}}{dt} = \frac{1}{\|\boldsymbol{\chi}\|\sin\theta_{s}}\left[
  \cos\phi_{s}(\Omega_{z}\chi_{x}-\Omega_{x}\chi_{z}) - \sin\phi_{s}(\Omega_{y}\chi_{z}-\Omega_{z}\chi_{y}) \right], \nonumber \\ 
i = 1,2 \end{eqnarray}
where the expression for $\Omega$ are given in \ref{eq:Omega}

\section{Computing the GW polarizations in the radiation frame}

In the absence of precession, the phase evolution of the (dominant harmonic) GWs is twice the orbital phase $\varphi(t)$. But the precession of the orbital plane introduces additional modulation in the orbital phase. If we define $\Phi(t)$ as the orbital phase with respect to the line of ascending nodes (the point at which the orbit crosses the $x-y$ plane from below) then the the phase evolution of the (dominant harmonic) GWs is given by $2\Phi(t)$, where $\Phi(t)$ is given by 
%
\begin{equation}
\label{eq:dPhiByDt}
\frac{d\Phi}{dt} = \omega - \frac{d\alpha}{dt} \cos i, 
\end{equation}
%
where $\alpha$ and $i$ are the angles describing the evolution of the orbital angular momentum vector $\LNh$ in the Finn-Chernoff coordinate system:
%
\begin{equation}
\alpha \equiv \arctan (\hat{L}_{Ny}/\hat{L}_{Nx}), ~~~ i \equiv \arccos (\hat{L}_{Nz}). 
\end{equation}
%
Time domain GW polarizations in the radiation frame can be written as 
%
\begin{eqnarray}
\label{eq:polzns_timedom}
h_{+,\times}(t) & = & \frac{-m \eta v^2}{d} ~\Bigg\{ C_{+,\times}\cos 2[\Phi(t)+\Phi_{0}] \nonumber \\ 
 & + & S_{+,\times}\sin 2[\Phi(t)+\Phi_{0}] \Bigg\}, \nonumber \\ 
\end{eqnarray} 
% 
which can be rewritten as 
%
\begin{eqnarray}
h_{+,\times}(t) = & A_{+,\times}(t) \, \cos \Psi_{+,\times}(t),
\end{eqnarray} 
%
where 
\begin{eqnarray}
\label{eq:polz_timedom_ampphase}
A_{+,\times}(t) & = & \frac{-m \eta v^2}{d} \sqrt{C_{+,\times}^2 + S_{+,\times}^2}, \nonumber \\ 
\Psi_{+,\times}(t) & = & 2 [\Phi(t) + \Phi_0] - \arctan \left(\frac{S_{+,\times}}{C_{+,\times}} \right),
\end{eqnarray}
where 
\begin{eqnarray}
C_+(t) & = & 2 \cos^2\alpha (\cos^2\Theta \cos^2 i+1)
	 -  \sin^2\alpha (\cos 2\Theta+\cos 2i+2) \nonumber \\ 
	& + & \cos \alpha \sin 2\Theta \sin 2i+2 \sin^2 \Theta \sin^2 i , \nonumber \\ 
S_+(t) & = & -2 \sin\alpha (\cos\alpha (\cos 2\Theta+3) \cos i + 2 \sin\Theta \cos\Theta \sin i), \nonumber \\ 
C_\times(t) & = & \sin \alpha (2 \cos \alpha \cos \Theta (\cos 2i+3)+4 \sin \Theta \sin i \cos i), \nonumber \\ 
S_\times(t) & = & 4 \cos 2\alpha \cos \Theta \cos i+ 4 \cos \alpha \sin \Theta \sin i.
\end{eqnarray}

\section{Computation of the Fourier transform}

In this section, we compute the Fourier transform of the time domain GW polarizations given in Eq.~(\ref{eq:polz_timedom_ampphase}).

\subsection{Stationary phase approximation}

We compute the Fourier transform of Eq.~(\ref{eq:polz_timedom_ampphase}) using stationary phase approximation (SPA). According to the SPA, the power in a Fourier frequency $f$ should mostly come from the saddle point $t_f$ given by the relation 
%
\begin{equation}
 f = F_{+,\times}(t_f).
 \label{eq:saddle_point}
\end{equation}
%
Above $F_{+,\times}(t_f)$ is the instantaneous GW frequency of the $+,\times$ polarization in the radiation frame. That is, 
\begin{equation}
\label{eq:saddle_point}
2 \pi F_{+,\times}(t) =  \frac{d\Psi_{+,\times}(t)}{dt} = 2 \omega_\mathrm{orb}(t) - \xi_{+,\times}(t), 
\end{equation}
where 
\begin{equation}
\xi_{+,\times}(t) = 2 \frac{d\alpha}{dt} \cos i - \frac{d}{dt} \arctan\left(\frac{S_{+,\times}}{C_{+,\times}}\right).
\end{equation}
% 
Then the Fourier transform can be written as as 
%
\begin{equation}
\tilde{h}_{+,\times}(f) = \mathcal{A}_{+,\times}(t_f) ~ e^{i[\Psi_{f+,\times}(t_f)-\pi/4]},
\end{equation}
where 
\begin{equation}
\label{eq:fourier_ampl}
\mathcal{A}_{+,\times}(t_f) = \frac{A_{+,\times}(t_f)}{2 \sqrt{\dot{F}_{+,\times}(t_f)}},
\end{equation}
and 
\begin{equation}
\label{eq:fourier_phase}
\Psi_{f\pc}(t_f) = 2\pi f t_{f}- 2 \varphi_\mathrm{orb}(t_f) + \int_{t_0}^{t_f} \xi_\pc(t) \, dt 
\end{equation}

The time-derivative of the instantaneous frequency appearing in Eq.~(\ref{eq:fourier_ampl}) can be written as 
%
\begin{eqnarray}
\frac{dF_\pc(t_f)}{dt} & = & \frac{dF_\pc}{dv} \frac{dv}{dt} \nonumber \\
 & = & -\frac{3v_f^2}{\pi m^2} \frac{\mathcal{F}(v_f)}{E'(v_f)}\left[1-\frac{m}{6v_f^2}\frac{d\xi_{+,\times}(t_f)}{dv} \right].
\end{eqnarray}
%
Using the binomial expansion of $(1-x)^{-1/2} \simeq (1+x/2)$, 
%
\begin{eqnarray}
\frac{1}{\sqrt{\dot{F}_{+,\times}(t_f)}} \simeq \sqrt{\frac{\pi}{3}} \frac{m}{v_f} \left[\frac{-E'(v_f)}{\mathcal{F}(v_f)} \right]^{1/2} \left[1+\frac{m}{12v_f^2} \frac{d\xi_{+,\times}(t_f)}{dv}\right]
\end{eqnarray}
%
Plugging this in Eq.~(\ref{eq:fourier_ampl}),
%
\begin{eqnarray}
\mathcal{A}_{+,\times}(t_f) & = &  \frac{m^2\eta v_f}{2d} \sqrt{\frac{\pi}{3}} \left[C_{+,\times}(t_f)^2 + S_{+,\times}(t_f)^2 \right]^{1/2} \nonumber \\ 
& & \left[\frac{-E'(v_f)}{\mathcal{F}(v_f)} \right]^{1/2} \left[1+ \frac{m}{12 v_f^2} \frac{d\xi_{+,\times}}{dv}\right]
\label{eq:spa_amp}
\end{eqnarray}
%
\subsection{Computing the saddle point}

For each Fourier frequency $f$, compute the corresponding saddle point $v_f$ (value of $v$ at the saddle point) by solving Eq.(\ref{eq:saddle_point}), which can be rewritten as 
%
\begin{equation}
v_f = v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f)}{2\pi f} \right]^{1/3} \simeq  v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f)}{6\pi f} \right]
\label{eq:saddle_point2}
\end{equation}
% 
where $v_f^0 = (\pi m f)^{1/3}$. Below, we describe the different approximations used in solving this equation.

\paragraph{\textbf{Approximation A}}: We neglect the correction to the saddle point due to the modulational effect of precession, by neglecting the contribution from $\xi_\pc$ in Eq.~(\ref{eq:saddle_point2}). We take $v_f^0$ as the saddle point corresponding to the Fourier frequency $f$.

Comparison of the amplitude and phase of the this version of SPA with the FFT is given in Figures~\ref{fig:fft_spa_comparison_approx_1_nonprec} and \ref{fig:fft_spa_comparison_approx_1_prec}. In each plot, the legend SPA$_1$ refers to the approximation in which the correction terms $\frac{m}{12 v_f^2} \frac{d\xi_{+,\times}}{dv}$ in Eq.(\ref{eq:spa_amp}) and $\int_{t_0}^{t_f} \xi_\pc(t) \, dt$ in Eq.(\ref{eq:fourier_phase}) are neglected, while SPA$_2$ refers to the version computed using the complete terms in  Eq.(\ref{eq:spa_amp}) and Eq.(\ref{eq:fourier_phase}).

\paragraph{\textbf{Approximation B}}: A slightly more accurate approximation of the saddle point is given by 
%
\begin{equation}
v_f \simeq  v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f^0)}{6\pi f} \right],
\label{eq:saddle_point2}
\end{equation}
%
where the correction term is computed at $v_f = v_f^0$. This refined approximation is being compared against the FFT.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tb]
\begin{center}
\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_1.00_m_20.00_spin1[0.00,0.00,0.98]_spin2[0.00,0.00,0.98]_iota1.57}.pdf}
\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_4.00_m_20.00_spin1[0.00,0.00,0.98]_spin2[0.00,0.00,0.98]_iota1.57}.pdf}
\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_10.00_m_20.00_spin1[0.00,0.00,0.98]_spin2[0.00,0.00,0.98]_iota1.57}.pdf}
\caption{Comparison of different SPA approximations with numerical FFTs in the case of some non-precessing-spin binaries. Left and middle panels show the amplitude and phase in the Fourier domain, while the right panels show the difference between the FFT and SPA phase. Parameters of the waveforms are shown in the titles.}
\label{fig:fft_spa_comparison_approx_1_nonprec}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tb]
\begin{center}
\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_1.00_m_20.00_spin1[0.98,0.00,0.00]_spin2[0.98,0.00,0.00]_iota1.57}.pdf}
\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_4.00_m_20.00_spin1[0.98,0.00,0.00]_spin2[0.98,0.00,0.00]_iota1.57}.pdf}
\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_10.00_m_20.00_spin1[0.98,0.00,0.00]_spin2[0.98,0.00,0.00]_iota1.57}.pdf}
\caption{Comparison of different SPA approximations with numerical FFTs in the case of some highly precessing binaries. Left and middle panels show the amplitude and phase in the Fourier domain, while the right panels show the difference between the FFT and SPA phase. Parameters of the waveforms are shown in the titles.}
\label{fig:fft_spa_comparison_approx_1_prec}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\appendix

\section{Ingredients for the Fourier domain waveforms}
The orbital phase is given by 
\begin{widetext}
\begin{align}
\begin{split}
\varphi_\mathrm{orb}(f) = &\frac{3}{256\eta v^{5}}\Bigg\lbrace 1 + v^{2}\Bigg\lbrack\frac{55\eta}{9}+\frac{3715}{756}\Bigg\rbrack + v^{3}\Bigg\lbrack\frac{113\delta\boldsymbol{\chi_{a}.L}}{3} +\frac{113\boldsymbol{\chi_{s}.L}}{3} - \frac{76\eta\boldsymbol{\chi_{s}.L}}{3} -16\pi\Bigg\rbrack + \\& v^{4}\Bigg\lbrack\frac{15293365}{508032}+\frac{27145 \eta }{504}+\frac{3085 \eta ^2}{72}-\frac{3595 \boldsymbol{\chi_{a}.L}^2}{48}+300 \eta  \boldsymbol{\chi_{a}.L}^2+\frac{1165 \chi_{a}^{2}}{48}- 100 \eta  \chi_{a}^{2}-\\& \frac{3595 \delta  \boldsymbol{\chi_{a}.L} \boldsymbol{\chi_{s}.L}}{24}-\frac{3595 \boldsymbol{\chi_{s}.L}^2}{48}-\frac{5 \eta  \boldsymbol{\chi_{s}.L}^2}{12}+\frac{1165 \delta  \chi_{s}.\chi_{a}}{24}+\frac{1165 \chi_{s}^{2}}{48}+\frac{35 \eta \chi _{s}^{2}}{12}\Bigg\rbrack + \\& v^{5}\Bigg\lbrack
-\frac{5 \delta  (146597+7056 \eta ) \boldsymbol{\chi_{a}.L}}{2268}-\frac{5 \left(3 \pi  (-7729+1092 \eta )+\left(146597-135856 \eta -17136 \eta ^2\right) \boldsymbol{\chi_{s}.L}\right)}{2268}\Bigg\rbrack \\&(3 \text{Log}[u]+1) +
v^{6}\Bigg\lbrack \frac{11583231236531}{4694215680}-\frac{15737765635 \eta}{3048192}+\frac{76055 \eta^2}{1728}-\frac{127825 \eta^3}{1296}-\frac{640 \pi ^2}{3}+\\&\frac{2255 \eta \pi ^2}{12}- \frac{6848 \text{$\gamma $E}}{21}-\frac{6848 \text{Log}[4]}{21}-\frac{6848 \text{Log}[u]}{21}\Bigg\rbrack + v^{7}\Bigg\lbrack\frac{77096675 \pi }{254016}+\frac{378515 \eta \pi }{1512}-\frac{74045 \eta^2 \pi }{756}\Bigg\rbrack
\end{split}
\label{eqn:fphase}
\end{align}
\end{widetext}

\bibliography{Spin}

\end{document}

### 2018-12-13 notes/Paper.tex
added notes on paper

@@
 \newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
-\newcommand{\ajith}[1]{\textcolor{red}{\textit{Ajith:#1}}}
-\newcommand{\ashok}[1]{\textcolor{cyan}{\textit{Ashok:#1}}}
+\newcommand{\ashok}[1]{\textcolor{red}{\textit{Ashok:#1}}}
+\newcommand{\sean}[1]{\textcolor{cyan}{\textit{Sean:#1}}}
 
@@
 
-\newcommand{\LIGO}{LIGO Laboratory, California Institute of Technology, 
-Pasadena, CA 91125, USA}
-\newcommand{\CIT}{Theoretical Astrophysics, California Institute of
-Technology, Pasadena, CA 91125, USA}
-\newcommand{\UWM}{UWM}
-
-\title{Notes on the SpinTaylorF2 project}
-\author{P.~Ajith}\email{ajith@icts.res.in}
-\affiliation{International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bangalore 560012, India}
-\author{Ashok Choudhary}
-\affiliation{International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bangalore 560012, India}
+\title{Spin effect on Christodoulou memory for merging blackhole binaries}
+\author{Ashok Choudhary}\email{aschoudhary@mix.wvu.edu}
+\author{Sean T. McWilliams}\email{sean.mcwilliams@mail.wvu.edu}
+\affiliation{Department of Physics and Astronomy, West Virginia University, Morgantown, WV 26506, USA}
 
 \begin{abstract}
+	The Christodoulou memory, which is a nonlinear memory effect sourced by the gravitational wave stress tensor, produce a growing, nonoscillatory change in the gravitational wave "plus" polarization.
+This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed.
+While the memory contribution during early in-spiral is well described by Post-Newtonian approximation, the contribution at merger must be obtained from Numerical relativity simulations.
+The Post-Newtonian corrections upto 3PN order to the gravitational wave memory for quasicircular, inspiralling compact binaries for non spinning has been computed by Favata et al.[Physical Review D 80, 024002 (2009)].
+Here we include the Spin contribution at leading order and calculate the memory effect upto 3PN order.
+We also compute memory contribution at merger from Numerical relativity using Weyl curvature component $\psi_4$ after removing the unphysical low frequency contribution due to spectral leakage.
+We use the method described in [Class.
+Quantum Grav.
+28 195015] for integration which control the amplification of the unphysical frequency resulting from spectral leakage.
 \end{abstract}
@@
 
-\section{Computing the spin and angular momentum vectors in the frequency domain}
-
-\subsection{Performing the evolution in Cartesian coordinates}
-
-The spin angular momentum vectors $\bS_i (v)$ and the unit vector $ \LNh (v)$ along the Newtonian angular momentum the frequency domain can be computed by using the evolution equations: 
-\begin{eqnarray}
-\frac{d \bS_i}{dv} & = & \frac{d \bS_i}{dt}  \frac{dt}{dv}, \\
-\frac{d \LNh}{dv}  & = & \frac{d \LNh}{dt} \frac{dt}{dv},  
-\label{eq:precessionEqFreqDom}
-\end{eqnarray}
-%
+\section{Introduction}
+Gravitational wave memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it.
+Nonlinear memory effect also called Christodoulou memory, arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave.
+It can be understood in the following way.
+Consider the Einstein field equation in harmonic gauge.
+[add ref]
+\begin{subequations}
+\begin{align}	
+\Box \bar{h}^{\alpha \beta} & = -16\pi (-g)(T^{\alpha\beta} + t_{LL}^{\alpha\beta})-\bar{h}^{\alpha\mu}_{,\nu}\bar{h}^{\beta\nu}_{,\mu}+\bar{h}^{\mu\nu}\bar{h}^{\alpha\beta}_{,\mu\nu}\\
+\bar{h}^{\alpha \beta},_{\beta} & = 0
+\end{align}
+\end{subequations}
 where
 \begin{equation}
-\frac{d \LNh}{dt}  = - \frac{1}{||\mathbf{L}||} \, \frac{d}{dt} (\bS_1 + \bS_2),
-\label{eq:orbitPrec}
+\bar{h}^{\alpha \beta} = \eta^{\alpha\beta}-\sqrt{-g}g^{\alpha\beta} 
 \end{equation}
-% 
-%
+is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator.
+The Landau-Lifshitz term is the Gravitational wave stress tensor[add ref].
 \begin{equation}
-\frac{d \bS_i}{dt}  =  \bm{\Omega}_i \times \bS_i \,,~ i = 1,2, 
-\label{eq:spinPrec}
+	T^{gw}_{jk}=\frac{1}{32\pi}\langle h^{TT}_{ab,j}h^{TT}_{ab,k}\rangle\approx T^{gw}_{00}n_{j}n_{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n_{j}n_{k}
 \end{equation}
-%
-and 
+where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab,j}=\bar{h}^{TT}_{ab,j}$.
+When we apply standard Green's function to right hand side of equation (1.1a), we obtain the following correction term to GW field [add ref]
+\begin{equation}
+	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'.N})}d\Omega'\Bigg]^{TT}
+\end{equation}
+where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy.
+
+\section{POST-NEWTONIAN WAVEFORM POLARIZATION AND MODE DECOMPOSITION}
+Following Kidder [add ref] we introduce a asymptotically-flat radiative coordinate system described by spherical coordinates $(T, R, \Theta, \Phi)$ and their orthogonal basis vectors $(\vec{e}_T, \vec{e}_R, \vec{e}_{\Theta}, \vec{e}_{\Phi})$.
+The coordinate origin is center-of-mass of the source.
+The retarded time in radiative coordinates is $T_R = T - R$
+\\
+The mode decomposition of gravitational wave polarization is given by
+\begin{equation}
+	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
+\end{equation}
+where
+\begin{equation}
+	h^{lm}= \frac{G}{\sqrt{2}\, R \, c^{l+2}}\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
+\end{equation}
+
+Here $U^{lm}(T_R)$ and $V^{lm}(T_R)$ are "scalar" mass and current multipoles respectively. The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
+\begin{equation}
+	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
+\end{equation} 
+Here
+\begin{align}
+	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
+	&\times \sum_{k=k_i}^{k_f}\frac{(-1)^k(\sin\frac{\Theta}{2})^{2k+s-1}(\cos\frac{\Theta}{2})^{1l+m-s-2k}}{k!(l+m-k)!(l-s-k)!(s-m+k)!}
+\end{align}
+where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$.
+
+\section{MEMORY CONTRIBUTION TO THE RADIATIVE-MASS MULTIPOLES}
+The GW polarization can be expanded on the basis of spin-weighted spherical harmonics.
+\begin{equation}
+	\sqrt{2}R(h_{+} - ih_{\times}) = \sum_{l=2}^{\infty}\sum_{m=-l}^{l}(U_{lm} - \mathit{i}V_{lm})_2Y^{lm}(\Theta, \Phi)
+\end{equation}
+Here $(R,\Theta,\Phi)$ are the spherical coordinates pointing from the source to the observer. The coefficient $U_{lm}$ and $V_{lm}$ are mass and current multipole moments. \\
+The memory contribution to the waveform polarization can be evaluated in the the following ways as described in [add referece]
+
+\subsection{Computing angular integral}
+
+\begin{equation}
+	\int d\Omega n_{L},
+\end{equation}
+
+\begin{equation}
+	\int d\Omega _{-2}Y_{l'm'}(\theta,\phi)_{-2}Y_{l''m''}^{*}(\theta, \phi)Y_{lm}^{*}(\theta, \phi)
+\end{equation}
+
+\begin{align}
+	U_{lm}^{(mem)(1)}=&R^{2}\sqrt{\frac{2(l-2)!}{(l+2)!}}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}(-1)^{m+m'}\\
+	&\times\Bigg\langle \dot{h}_{l'm'}\dot{h}_{l'm'}^{*}\Bigg\rangle G^{2-20}_{l'm'lm'-m''-m}
+\end{align}
+
+\subsection{Time derivatives of the memory multipole moments}
+
+\begin{equation}
+	\int_{-\infty}^{T_{R}} x^{n}dt=\int_{-\infty}^{T_{R}}\frac{x^{n}}{\dot{x}}dx
+\end{equation}
+
+\subsection{Computing time integral over the past history of the source}
+sesadsd
+
 % 
 \begin{widetext}
-\begin{align}\label{eq:EbyF}
-\frac{dt}{dv} & = \frac{5m}{32 \, v^9  \, \eta} \left\{1 + v^2 \left[\frac{11 \eta }{4}+\frac{743}{336} \right] 
-    + v^3 \left[\frac{113}{12} \left(\biggl( 1-\frac{76\eta}{113}\biggr) \, \chisdL + \delta \, \chiadL \right)-4 \pi \right] \right. \nonumber \\ 
-    &\quad + v^4 \left[(\chiadL)^2 \left(30 \eta -\frac{719}{96}\right)-\frac{719 \chiadL \, \chisdL \delta
-   }{48}+\chiaSqr \left(\frac{233}{96}-10 \eta \right)+\frac{233 \chisDchia \delta}{48} \right. \nonumber\\
-   &\qquad \left. +(\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
-   }{24}+\frac{233}{96}\right)+\frac{617 \eta ^2}{144}+\frac{5429 \eta }{1008}+\frac{3058673}{1016064} \right] \nonumber \\ 
-    &\quad + v^5 \left[\chiadL \delta  \left(\frac{7 \eta }{2}+\frac{146597}{2016}\right)+\chisdL \left(-\frac{17 \eta
-   ^2}{2}-\frac{1213 \eta }{18}+\frac{146597}{2016}\right)+\frac{13 \pi  \eta }{8}-\frac{7729 \pi }{672} \right]  \nonumber\\ 
-    &\quad + v^6 \left[\frac{1712 \gamma_E }{105}+\frac{25565 \eta ^3}{5184}-\frac{15211 \eta
-   ^2}{6912}-\frac{451 \pi ^2 \eta }{48}+\frac{3147553127 \eta }{12192768}+\frac{32 \pi
-   ^2}{3}-\frac{10817850546611}{93884313600}+\frac{1712 \ln (4v)}{105} \right]  \nonumber\\ 
-    &\quad \left. + v^7 \left[\frac{14809 \pi  \eta ^2}{3024}-\frac{75703 \pi  \eta }{6048}-\frac{15419335 \pi }{1016064} \right] \right\}. 
-\end{align}
-\end{widetext}
-% 
-(\ajith{Need to add the BH absorption term above}). The (orbit averaged) precession frequencies $\bm{\Omega}_i$ appearing in Eq.~(\ref{eq:spinPrec}) are given by 
-\begin{align}
-\label{eq:Omega}
-\bm{\Omega}_1 & =  \frac{v^5}{m} \left\{\left(\frac{3}{4}+\frac{\eta}{2}-\frac{3\delta}{4}\right) \LNh \right. \nonumber \\
-     & \quad +  \frac{v}{2m^2} \left[-3 \, (\bS_2 + q \, \bS_1).\LNh ~ \LNh + \bS_2 \right]  \nonumber \\
-     & \quad +  \left. v^2 \left(\frac{9}{16} + \frac{5\eta}{4} - \frac{\eta^2}{24} - \frac{9\delta}{16} + \frac{5\delta \eta}{8} \right)  \LNh  \right\}, \nonumber \\
-%
-\bm{\Omega}_2 & =  \frac{v^5}{m} \left\{\left(\frac{3}{4}+\frac{\eta}{2}+\frac{3\delta}{4}\right) \LNh \right. \nonumber \\
-     & \quad +  \frac{v}{2m^2} \left[-3 \, (\bS_1 + q^{-1} \, \bS_2).\LNh ~ \LNh + \bS_1 \right]  \nonumber \\
-     & \quad +  \left. v^2 \left(\frac{9}{16} + \frac{5\eta}{4} - \frac{\eta^2}{24} + \frac{9\delta}{16} - \frac{5\delta \eta}{8} \right)  \LNh  \right\}. 
-\end{align}
-
-%
-Above, $v \equiv (m \omega)^{1/3}$ is the PN expansion parameter, $\omega$ is the orbital frequency, $m \equiv m_1 + m_2$ is the total mass of the binary, $q \equiv m_2/m_1$ is the mass ratio, $\eta \equiv m_1 m_2/m^2$ is the symmetric mass ratio, $\delta \equiv (m_1-m_2)/m$ the asymmetric mass ratio, $\gamma_E$ the Euler's constant, and $\LNh$ the unit vector along the Newtonian orbital angular momentum $\LN \equiv m^2\eta/v ~ \LNh$.
-The spin variables $\bchis$ and $\bchia$ are related to the (dimensionless) spin vectors of the binary as:
-%
-\begin{equation}
-\bchis = (\bchi_1 + \bchi_2)/2, ~~~ \bchia = (\bchi_1 - \bchi_2)/2,
-\end{equation}
-%
-where $\bchi_i \equiv \bS_i/m_i^2$.
-
-\subsection{Performing the evolution in spherical polar coordinates}
-
-The time evolution equation of $i \equiv \arccos L_{Nz}$ and $ \alpha \equiv \arctan (L_{Ny}/L_{Nx})$ are given by:
-\begin{equation}
-\frac{d\iota}{dt} = \frac{-1}{\|L\| \sin\iota}\sum\limits_{k = 1}^{2}m^2_{k}\| \boldsymbol{\chi}_{k}\|\sin\theta_{s_{k}}\frac{d\theta_{s_{k}}}{dt}
-\end{equation}  
-and
-\begin{widetext}
-\begin{align}
-\begin{split}
-\frac{d\alpha}{dt} =& \frac{\cos\alpha}{\sin\iota}\Bigg(\frac{-1}{\|\boldsymbol{L}\|}\sum\limits_{k = 1}^{2} m^{2}_{k}\|\boldsymbol{{\chi}}_{k}\|\left[\frac{d\theta_{S_{k}}}{dt}\cos\theta_{s_{k}}\sin\phi_{s_{k}} + \frac{d\phi_{S_{k}}}{dt}\sin\theta_{s_{k}}\cos\phi_{s_{k}}\right]\Bigg) \\
-- & \frac{\sin\alpha}{\sin\iota}\Bigg(\frac{-1}{\|\boldsymbol{L}\|}\sum\limits_{k = 1}^{2} m^{2}_{k}\|\boldsymbol{{\chi}}_{k}\|\left[\frac{d\theta_{S_{k}}}{dt}\cos\theta_{s_{k}}\cos\phi_{s_{k}} - \frac{d\phi_{S_{k}}}{dt}\sin\theta_{s_{k}}\sin\phi_{s_{k}}\right]\Bigg)
-\end{split}
+\begin{align}\label{eq:FbydE}
+\frac{dx}{dt} & = \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
+    + x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
+    &\quad + x^2 \left[ \frac{34103}{18144} + \frac{13661 \, \eta}{2016} + \frac{59 \, \eta^2}{18} + \left( \frac{719}{96}- 30 \eta\right)(\chiadL)^2 +\chiaSqr \left(-\frac{233}{96}+10 \eta \right) \right.
+\nonumber\\
+   &\qquad+\frac{719 \chiadL \, \chisdL \delta
+   }{48}-\frac{233 \chisDchia \delta}{48} \left. + \, (\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
+   }{24}+\frac{233}{96}\right) \right] \nonumber \\ 
+    &\quad + x^{5/2} \left[-\frac{4159 \, \pi}{672}-\frac{189 \, \pi \, \eta}{8} + \chiadL \delta  \left(-\frac{31319}{1008} + \frac{1159 \, \eta}{24}\right)+\chisdL \left(-\frac{31319}{1008} + \frac{22975 \, \eta}{252} -\frac{79 \, \eta^2}{3}\right) \right]  \nonumber\\
+    &\quad \left.
++ x^3 \left[\frac{16447322263}{139708800} + \frac{16 \, \pi^2}{3} -\frac{1712 \, \gamma_{E}}{105} -\frac{56198689 \, \eta}{217728} + \frac{451 \, \pi^2 \, \eta}{48} + \frac{541 \, \eta^2}{896} -\frac{5605 \, \eta^{3}}{2592} -\frac{80 \,\pi \delta\chiadL }{3} \right.\right.
+\nonumber \\
+    &\qquad \left.\left.
++ \left(\frac{575}{448} + \frac{565 \, \delta^2}{9}-\frac{65815 \, \eta}{4032} + \frac{89 \, \eta^2}{2}\right)(\chiadL)^2 + \left(-\frac{145}{448}+\frac{21985 \,\eta}{4032} -\frac{89 \, \eta^2}{6}\right)\chiaSqr + \right(-\frac{80 \, \pi}{3} + \frac{40 \, \pi \, \eta}{3} \right.
+\nonumber \\
+    &\qquad  \left.+\delta\left(\frac{258295}{2016} -\frac{9203 \, \eta}{96}\right)\chiadL\left)\chisdL + \left(\frac{258295}{4032} -\frac{48773 \, \eta}{576} + \frac{3041 \, \eta^2}{144}\right)(\chisdL)^2 + \delta\left(-\frac{145}{224} + \frac{2143 \, \eta}{288}\right)\chisDchia \right.\right.\nonumber \\
+    &\qquad\left.\left.+\left(-\frac{145}{448} + \frac{1891 \, \eta}{576} -\frac{7 \, \eta^2}{144}\right)\chisSqr - \frac{3424 \, \log[2]}{105} - \frac{1712 \log[x]}{210}\right]\right\}
 \end{align}
@@
 
- The angles corresponding to spin vectors($S_{1}$ and $S_{2}$)can also be written as:
- \begin{eqnarray}
- \frac{d\theta_{s}}{dt} = \frac{1}{\|\boldsymbol{\chi}\|\cos\theta_{s}}\left[
- \cos\phi_{s}(\Omega_{y}\chi_{z}-\Omega_{z}\chi_{y}) + \sin\phi_{s}(\Omega_{z}\chi_{x}-\Omega_{x}\chi_{z}) \right], \nonumber \\ 
-  \frac{d\phi_{s}}{dt} = \frac{1}{\|\boldsymbol{\chi}\|\sin\theta_{s}}\left[
-  \cos\phi_{s}(\Omega_{z}\chi_{x}-\Omega_{x}\chi_{z}) - \sin\phi_{s}(\Omega_{y}\chi_{z}-\Omega_{z}\chi_{y}) \right], \nonumber \\ 
-i = 1,2 \end{eqnarray}
-where the expression for $\Omega$ are given in \ref{eq:Omega}
+\subsection{RESULTS: MEMORY CONTRIBUTION TO THE POST-NEWTONIAN WAVEFORM OF QUASI-CIRCULAR, INSPIRALLING BINARIES}
+The memory contribution to the spin-weighted spherical-harmonic modes of the polarization waveform [ref eqn]. These quantities are related via 
+\begin{equation}
+	h_{l0}^{(mem)} = \frac{\alpha}{\sqrt{2} \, R} U_{l0}^{(mem)} = 8 \sqrt{\frac{\pi}{5}}\frac{\eta M x}{R} 
+\end{equation}
+where we have followed the notation of Sec. 9 of Ref [add referece]. The notational parameter $\alpha$ accounts for the two commonly used choices for the polarization triad 
+\begin{equation}
+	content...
+\end{equation}
+(\ashok{Need to add}).\\
+The results of polarization modes in terms of $\hat{H}_{l0}$ are:
+\begin{widetext}
+\begin{align}\label{eq:H20}
+\hat{H}_{20} &= \alpha \frac{5}{15\sqrt{6}}\left\{ 1 + x\left(-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right)+ x^{3/2}\left(\left(-\frac{7}{1440} -\frac{27 \, \delta}{10} + \frac{7 \,\eta}{360}\right)\chiadL+\left(-\frac{27}{10} -\frac{7 \, \delta^3}{1440} + \frac{22 \, \eta}{15}\right)\chisdL\right)\right.
+\nonumber \\
+& \quad \left.
+x^2\left(-\frac{151877213}{67060224} -\frac{123815 \, \eta}{44352} +\frac{205 \, \eta^2}{352} -\left(\frac{1459}{864} -\frac{1441 \, \eta}{216}\right)\chiaSqr - \left(\frac{27 \, \delta}{8} +\frac{\delta^3}{432}\right)\chisDchia - \left(\frac{27}{16}+\frac{\delta^4}{864} -\frac{\eta}{12}\right)\chisSqr\right)\right.
+\nonumber \\
+& \quad \left.
+x^{5/2}\left(-\frac{253 \, \pi}{336}+ \frac{253 \, \pi \, \eta}{84} -\left(-\frac{631}{96768} + \frac{2729857 \, \delta}{84672} -\frac{146597 \, \pi \, \delta}{1764}+\frac{5 \, \eta}{378}+\frac{3007 \, \delta \, \eta}{1008} -4\pi \, \delta \, \eta +\frac{311 \, \eta^2}{6048}\right)\chiadL\right.\right.
+\nonumber\\
+& \qquad\left.
+-\left(\frac{2729857}{84672}-\frac{146597 \, \pi}{1764}-\frac{631 \, \delta^3 }{96798} -\frac{44111 \, \eta}{1764}+\frac{4852 \, \pi \, \eta}{63}-\frac{311 \, \delta^3 \, \eta}{24192}-\frac{1121 \, \eta^2}{252} +\frac{68 \, \pi \, \eta^2}{7}\right)\chisdL\right) \nonumber \\
+& \quad  x^3\left[-\frac{4397711103307}{532580106240} + \left( \frac{700464542023}{13948526592} - \frac{205 \, \pi^2}{96}\right)\eta + \frac{69527951 \, \eta^2}{166053888} + \frac{1321981 \, \eta^3}{5930496} -  \left(-\frac{160867}{13824} -\frac{791 \, \delta}{27648} + \right.\right.
+\nonumber \\
+& \qquad \left.
+\frac{7345 \, \delta^2}{576} + \left(\frac{3953767}{96768}+\frac{791 \, \delta}{6912}\right)\eta + \frac{22885 \, \eta^2}{1152}\right)\chiaSqr -\left(-\frac{275 \, \pi}{12} + \frac{7 \, \pi \, \delta^3}{768} + 17 \, \pi \, \eta \right)\chisdL + \left(\frac{47869}{43008}  \right.\nonumber\\
+& \qquad \left.\left.
+- \frac{791 \, \delta^3}{27648} + \frac{743 \, \delta^4}{387072} + \left(-\frac{410453}{16128} + \frac{133 \, \delta^3}{6912} + \frac{11 \, \delta^4}{4608}\right)\eta + \frac{10229 \, \eta^2}{1152}\right)\chisSqr - \left( \frac{7 \, \pi}{768}-\frac{275 \, \pi \, \delta}{12}-\frac{7 \, \pi \, \eta}{192}\right.\right.
+\nonumber \\
+&\qquad \left.\left.\left.
++ \left(-\frac{791}{27648} + \frac{47869 \, \delta}{21504} + \frac{743 \, \delta^3}{193536} - \frac{791 \, \delta^4}{27648} + \left(\frac{77}{576} - \frac{71749 \, \delta}{2304} + \frac{11 \, \delta^3}{2304}\right)\eta - \frac{133 \, \eta^2}{1728}\right)\chisdL\right)\chisdL
+\right]
+\right\}
+\end{align}
 
-\section{Computing the GW polarizations in the radiation frame}
+\end{widetext}
 
-In the absence of precession, the phase evolution of the (dominant harmonic) GWs is twice the orbital phase $\varphi(t)$.
-But the precession of the orbital plane introduces additional modulation in the orbital phase.
-If we define $\Phi(t)$ as the orbital phase with respect to the line of ascending nodes (the point at which the orbit crosses the $x-y$ plane from below) then the the phase evolution of the (dominant harmonic) GWs is given by $2\Phi(t)$, where $\Phi(t)$ is given by
-%
+\section{Memory calculation using Numerical relativity data}
+Numerical relativity simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by
 \begin{equation}
-\label{eq:dPhiByDt}
-\frac{d\Phi}{dt} = \omega - \frac{d\alpha}{dt} \cos i, 
-\end{equation}
-%
-where $\alpha$ and $i$ are the angles describing the evolution of the orbital angular momentum vector $\LNh$ in the Finn-Chernoff coordinate system:
-%
+	\psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_{\times}
+\end{equation}\
+Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$.
+It has long been noted that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift.
+The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation.
+This nonlinearity is potentially caused by the fact that $\psi_4$ is typically extracted at a finite distance from the gravitating source.
+The strain $h$ is only related to $\psi_4$ at an infinite distance and also strictly valid in a particular gauge.
+As a result in numerical simulation, the finite distance calculation introduces a systematic error.
+In more recent simulations [added ref], which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
+\par It has been argued in [add ref] that an important source of unphysical non-linear drift in numerical computation of gravitational wave strain lies in the transformation of the measured data to the strain $h$  which generally involves an integration in time.
+The output of the numerical simulation is a discretely sampled time series of finite duration, incorporating some component of unresolved frequencies due to numerical error.
+This can lead to an uncontrollable non-linear drift if the integration is performed in the time domain.
+\par The memory contribution to $h_{+}$ is given by
+
 \begin{equation}
-\alpha \equiv \arctan (\hat{L}_{Ny}/\hat{L}_{Nx}), ~~~ i \equiv \arccos (\hat{L}_{Nz}). 
-\end{equation}
-%
-Time domain GW polarizations in the radiation frame can be written as 
-%
-\begin{eqnarray}
-\label{eq:polzns_timedom}
-h_{+,\times}(t) & = & \frac{-m \eta v^2}{d} ~\Bigg\{ C_{+,\times}\cos 2[\Phi(t)+\Phi_{0}] \nonumber \\ 
- & + & S_{+,\times}\sin 2[\Phi(t)+\Phi_{0}] \Bigg\}, \nonumber \\ 
-\end{eqnarray} 
-% 
-which can be rewritten as 
-%
-\begin{eqnarray}
-h_{+,\times}(t) = & A_{+,\times}(t) \, \cos \Psi_{+,\times}(t),
-\end{eqnarray} 
-%
-where 
-\begin{eqnarray}
-\label{eq:polz_timedom_ampphase}
-A_{+,\times}(t) & = & \frac{-m \eta v^2}{d} \sqrt{C_{+,\times}^2 + S_{+,\times}^2}, \nonumber \\ 
-\Psi_{+,\times}(t) & = & 2 [\Phi(t) + \Phi_0] - \arctan \left(\frac{S_{+,\times}}{C_{+,\times}} \right),
-\end{eqnarray}
-where 
-\begin{eqnarray}
-C_+(t) & = & 2 \cos^2\alpha (\cos^2\Theta \cos^2 i+1)
-	 -  \sin^2\alpha (\cos 2\Theta+\cos 2i+2) \nonumber \\ 
-	& + & \cos \alpha \sin 2\Theta \sin 2i+2 \sin^2 \Theta \sin^2 i , \nonumber \\ 
-S_+(t) & = & -2 \sin\alpha (\cos\alpha (\cos 2\Theta+3) \cos i + 2 \sin\Theta \cos\Theta \sin i), \nonumber \\ 
-C_\times(t) & = & \sin \alpha (2 \cos \alpha \cos \Theta (\cos 2i+3)+4 \sin \Theta \sin i \cos i), \nonumber \\ 
-S_\times(t) & = & 4 \cos 2\alpha \cos \Theta \cos i+ 4 \cos \alpha \sin \Theta \sin i.
-\end{eqnarray}
-
-\section{Computation of the Fourier transform}
-
-In this section, we compute the Fourier transform of the time domain GW polarizations given in Eq.~(\ref{eq:polz_timedom_ampphase}).
-
-\subsection{Stationary phase approximation}
-
-We compute the Fourier transform of Eq.~(\ref{eq:polz_timedom_ampphase}) using stationary phase approximation (SPA).
-According to the SPA, the power in a Fourier frequency $f$ should mostly come from the saddle point $t_f$ given by the relation
-%
-\begin{equation}
- f = F_{+,\times}(t_f).
- \label{eq:saddle_point}
-\end{equation}
-%
-Above $F_{+,\times}(t_f)$ is the instantaneous GW frequency of the $+,\times$ polarization in the radiation frame. That is, 
-\begin{equation}
-\label{eq:saddle_point}
-2 \pi F_{+,\times}(t) =  \frac{d\Psi_{+,\times}(t)}{dt} = 2 \omega_\mathrm{orb}(t) - \xi_{+,\times}(t), 
-\end{equation}
-where 
-\begin{equation}
-\xi_{+,\times}(t) = 2 \frac{d\alpha}{dt} \cos i - \frac{d}{dt} \arctan\left(\frac{S_{+,\times}}{C_{+,\times}}\right).
-\end{equation}
-% 
-Then the Fourier transform can be written as as 
-%
-\begin{equation}
-\tilde{h}_{+,\times}(f) = \mathcal{A}_{+,\times}(t_f) ~ e^{i[\Psi_{f+,\times}(t_f)-\pi/4]},
-\end{equation}
-where 
-\begin{equation}
-\label{eq:fourier_ampl}
-\mathcal{A}_{+,\times}(t_f) = \frac{A_{+,\times}(t_f)}{2 \sqrt{\dot{F}_{+,\times}(t_f)}},
-\end{equation}
-and 
-\begin{equation}
-\label{eq:fourier_phase}
-\Psi_{f\pc}(t_f) = 2\pi f t_{f}- 2 \varphi_\mathrm{orb}(t_f) + \int_{t_0}^{t_f} \xi_\pc(t) \, dt 
+	h_{+}^{(mem)} \approx \frac{R}{192 \, \pi}s_{\Theta}^{2}\left(17 +c_{\Theta}^{2} \right)\int_{-\infty}^{T_R}|\dot{h}_{22}|^2 dt
 \end{equation}
 
-The time-derivative of the instantaneous frequency appearing in Eq.~(\ref{eq:fourier_ampl}) can be written as 
-%
-\begin{eqnarray}
-\frac{dF_\pc(t_f)}{dt} & = & \frac{dF_\pc}{dv} \frac{dv}{dt} \nonumber \\
- & = & -\frac{3v_f^2}{\pi m^2} \frac{\mathcal{F}(v_f)}{E'(v_f)}\left[1-\frac{m}{6v_f^2}\frac{d\xi_{+,\times}(t_f)}{dv} \right].
-\end{eqnarray}
-%
-Using the binomial expansion of $(1-x)^{-1/2} \simeq (1+x/2)$, 
-%
-\begin{eqnarray}
-\frac{1}{\sqrt{\dot{F}_{+,\times}(t_f)}} \simeq \sqrt{\frac{\pi}{3}} \frac{m}{v_f} \left[\frac{-E'(v_f)}{\mathcal{F}(v_f)} \right]^{1/2} \left[1+\frac{m}{12v_f^2} \frac{d\xi_{+,\times}(t_f)}{dv}\right]
-\end{eqnarray}
-%
-Plugging this in Eq.~(\ref{eq:fourier_ampl}),
-%
-\begin{eqnarray}
-\mathcal{A}_{+,\times}(t_f) & = &  \frac{m^2\eta v_f}{2d} \sqrt{\frac{\pi}{3}} \left[C_{+,\times}(t_f)^2 + S_{+,\times}(t_f)^2 \right]^{1/2} \nonumber \\ 
-& & \left[\frac{-E'(v_f)}{\mathcal{F}(v_f)} \right]^{1/2} \left[1+ \frac{m}{12 v_f^2} \frac{d\xi_{+,\times}}{dv}\right]
-\label{eq:spa_amp}
-\end{eqnarray}
-%
-\subsection{Computing the saddle point}
+The above equation requires a time integration of absolute value of $\dot{h}_{22}$, but this is not typically the quantity which is directly computed in numerical simulation.
+The output of numerical simulation is usually in the form of component of curvature tensor components, or Zerilli-Moncrief-type variables defined relative to a background.
+\subsection{Evaluating Bondi News from numerical data}
+The Bondi News is determined by
+\begin{equation}
+	\mathcal{N} = \int_{-\infty}^{t}dt' \psi_4
+\end{equation}
 
-For each Fourier frequency $f$, compute the corresponding saddle point $v_f$ (value of $v$ at the saddle point) by solving Eq.(\ref{eq:saddle_point}), which can be rewritten as 
-%
+\subsection{Integration of finite length signals in frequency domain}
+
+Consider the Fourier transform, $\mathcal{F}$, applied to an absolutely integrable function $f(t)$,
 \begin{equation}
-v_f = v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f)}{2\pi f} \right]^{1/3} \simeq  v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f)}{6\pi f} \right]
-\label{eq:saddle_point2}
+	\tilde{f}(\omega) = \mathcal{F}[f]=\int_{-\infty}^{\infty}\e^{-\iota \omega t}f(t)dt.	
 \end{equation} 
-% 
-where $v_f^0 = (\pi m f)^{1/3}$. Below, we describe the different approximations used in solving this equation.
+The Fourier transform of the time integral is given by
+\begin{equation}
+	\mathcal{F}\left[\int_{-\infty}^{t}dt' f(t')\right] |_\omega = -\iota \frac{\tilde{f}(\omega)}{\omega}
+\end{equation}
 
-\paragraph{\textbf{Approximation A}}: We neglect the correction to the saddle point due to the modulational effect of precession, by neglecting the contribution from $\xi_\pc$ in Eq.~(\ref{eq:saddle_point2}).
-We take $v_f^0$ as the saddle point corresponding to the Fourier frequency $f$.
+\subsubsection{Fixed frequency integration ($\mathit{FFI}$)}
 
-Comparison of the amplitude and phase of the this version of SPA with the FFT is given in Figures~\ref{fig:fft_spa_comparison_approx_1_nonprec} and \ref{fig:fft_spa_comparison_approx_1_prec}.
-In each plot, the legend SPA$_1$ refers to the approximation in which the correction terms $\frac{m}{12 v_f^2} \frac{d\xi_{+,\times}}{dv}$ in Eq.(\ref{eq:spa_amp}) and $\int_{t_0}^{t_f} \xi_\pc(t) \, dt$ in Eq.(\ref{eq:fourier_phase}) are neglected, while SPA$_2$ refers to the version computed using the complete terms in  Eq.(\ref{eq:spa_amp}) and Eq.(\ref{eq:fourier_phase}).
+We evaluate the integral using the following mathod [add ref]
 
-\paragraph{\textbf{Approximation B}}: A slightly more accurate approximation of the saddle point is given by 
-%
-\begin{equation}
-v_f \simeq  v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f^0)}{6\pi f} \right],
-\label{eq:saddle_point2}
-\end{equation}
-%
-where the correction term is computed at $v_f = v_f^0$. This refined approximation is being compared against the FFT.
+\[
+\tilde{F} =
+\begin{cases}
+-\mathit{i} \frac{\tilde{f}(\omega)}{\omega_0} & \text{ $\omega\leq\omega_0$} \\
+-\mathit{i} \frac{\tilde{f}(\omega)}{\omega} & \text{ $\omega>\omega_0$}
 
+\end{cases}
+\]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[tb]
-\begin{center}
-\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_1.00_m_20.00_spin1[0.00,0.00,0.98]_spin2[0.00,0.00,0.98]_iota1.57}.pdf}
-\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_4.00_m_20.00_spin1[0.00,0.00,0.98]_spin2[0.00,0.00,0.98]_iota1.57}.pdf}
-\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_10.00_m_20.00_spin1[0.00,0.00,0.98]_spin2[0.00,0.00,0.98]_iota1.57}.pdf}
-\caption{Comparison of different SPA approximations with numerical FFTs in the case of some non-precessing-spin binaries.
-Left and middle panels show the amplitude and phase in the Fourier domain, while the right panels show the difference between the FFT and SPA phase.
-Parameters of the waveforms are shown in the titles.}
-\label{fig:fft_spa_comparison_approx_1_nonprec}
-\end{center}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=4.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
+\caption{NonSpinning case}
+\label{fig:q7}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[tb]
-\begin{center}
-\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_1.00_m_20.00_spin1[0.98,0.00,0.00]_spin2[0.98,0.00,0.00]_iota1.57}.pdf}
-\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_4.00_m_20.00_spin1[0.98,0.00,0.00]_spin2[0.98,0.00,0.00]_iota1.57}.pdf}
-\includegraphics[width=5.0in]{{../plots/spa/SPA_SpinAndOrbitEv_SpinTaylorT5_q_10.00_m_20.00_spin1[0.98,0.00,0.00]_spin2[0.98,0.00,0.00]_iota1.57}.pdf}
-\caption{Comparison of different SPA approximations with numerical FFTs in the case of some highly precessing binaries.
-Left and middle panels show the amplitude and phase in the Fourier domain, while the right panels show the difference between the FFT and SPA phase.
-Parameters of the waveforms are shown in the titles.}
-\label{fig:fft_spa_comparison_approx_1_prec}
-\end{center}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=4.0in]{../plots/MemoryPlot_alignedSpin/0p95.pdf}
+\caption{NonSpinning case}
+\label{fig:0p95}
+\end{figure}
 
@@
 
-\section{Ingredients for the Fourier domain waveforms}
-The orbital phase is given by 
-\begin{widetext}
-\begin{align}
-\begin{split}
-\varphi_\mathrm{orb}(f) = &\frac{3}{256\eta v^{5}}\Bigg\lbrace 1 + v^{2}\Bigg\lbrack\frac{55\eta}{9}+\frac{3715}{756}\Bigg\rbrack + v^{3}\Bigg\lbrack\frac{113\delta\boldsymbol{\chi_{a}.L}}{3} +\frac{113\boldsymbol{\chi_{s}.L}}{3} - \frac{76\eta\boldsymbol{\chi_{s}.L}}{3} -16\pi\Bigg\rbrack + \\& v^{4}\Bigg\lbrack\frac{15293365}{508032}+\frac{27145 \eta }{504}+\frac{3085 \eta ^2}{72}-\frac{3595 \boldsymbol{\chi_{a}.L}^2}{48}+300 \eta  \boldsymbol{\chi_{a}.L}^2+\frac{1165 \chi_{a}^{2}}{48}- 100 \eta  \chi_{a}^{2}-\\& \frac{3595 \delta  \boldsymbol{\chi_{a}.L} \boldsymbol{\chi_{s}.L}}{24}-\frac{3595 \boldsymbol{\chi_{s}.L}^2}{48}-\frac{5 \eta  \boldsymbol{\chi_{s}.L}^2}{12}+\frac{1165 \delta  \chi_{s}.\chi_{a}}{24}+\frac{1165 \chi_{s}^{2}}{48}+\frac{35 \eta \chi _{s}^{2}}{12}\Bigg\rbrack + \\& v^{5}\Bigg\lbrack
--\frac{5 \delta  (146597+7056 \eta ) \boldsymbol{\chi_{a}.L}}{2268}-\frac{5 \left(3 \pi  (-7729+1092 \eta )+\left(146597-135856 \eta -17136 \eta ^2\right) \boldsymbol{\chi_{s}.L}\right)}{2268}\Bigg\rbrack \\&(3 \text{Log}[u]+1) +
-v^{6}\Bigg\lbrack \frac{11583231236531}{4694215680}-\frac{15737765635 \eta}{3048192}+\frac{76055 \eta^2}{1728}-\frac{127825 \eta^3}{1296}-\frac{640 \pi ^2}{3}+\\&\frac{2255 \eta \pi ^2}{12}- \frac{6848 \text{$\gamma $E}}{21}-\frac{6848 \text{Log}[4]}{21}-\frac{6848 \text{Log}[u]}{21}\Bigg\rbrack + v^{7}\Bigg\lbrack\frac{77096675 \pi }{254016}+\frac{378515 \eta \pi }{1512}-\frac{74045 \eta^2 \pi }{756}\Bigg\rbrack
-\end{split}
-\label{eqn:fphase}
-\end{align}
-\end{widetext}
-
 \bibliography{Spin}

### 2018-12-23 notes/Paper.tex
updated latex

@@
 \section{Introduction}
-Gravitational wave memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it.
+
+Gravitational wave (GW) memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it.
+The are two kinds of gravitational wave memory: linear and nonlinear.
+Linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments.
+At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
+\begin{equation}
+	\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
+\end{equation} 
+where $I_{ij}$ is the source mass-quadrupole moment, R is the distance source and $\Delta$ describes the net change in the quantity from early to late times.
+Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion [add ref], asymmetric mass loss due to neutrino emission and gamma-ray-burst jets [add ref] \\
+
+A general formula for the linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by Thorn [add ref]
+
+\begin{equation}
+\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
+\end{equation}
+
+where the masses are unbound in their initial and final states(or both).
+Here $\Delta$ refers to take difference between the final and initial values of the summations, and $\boldsymbol{N}$ is the unit vector that points from the source to the observer.
+\\
  Nonlinear memory effect also called Christodoulou memory, arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave.
@@
 \begin{equation}
-	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'.N})}d\Omega'\Bigg]^{TT}
+	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'}\cdot\boldsymbol{N})}d\Omega'\Bigg]^{TT}
 \end{equation}
-where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy.
+where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy.\\
+The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity.
+It has a clear physical interpretation : It arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments.
+The nonlinear memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order.
+This memory effect causes a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral.
+The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence but, most contribution to this nonlinear memory come from the merger phase where the PN model do not work.
+In this case the memory must be extracted from Numerical relativity simulation.
+However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes.
+Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.\\
 
-\section{POST-NEWTONIAN WAVEFORM POLARIZATION AND MODE DECOMPOSITION}
-Following Kidder [add ref] we introduce a asymptotically-flat radiative coordinate system described by spherical coordinates $(T, R, \Theta, \Phi)$ and their orthogonal basis vectors $(\vec{e}_T, \vec{e}_R, \vec{e}_{\Theta}, \vec{e}_{\Phi})$.
-The coordinate origin is center-of-mass of the source.
-The retarded time in radiative coordinates is $T_R = T - R$
-\\
+The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata [].
+However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})
+
+In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.In Sec.
+II, we describe the calculation required to compute memory contribution to the post-Newtonian waveform of quasi-circular, inspiralling binaries.
+In Sec.
+III, we describe the method used extract memory contribution from NR simulations.
+Sec.
+IV we give the results which looks at the effect of spin on total nonlinear memory.
+We use PN expressions for memory for early time and for late time we use memory extracted from NR simulations.
+
+\section{MEMORY CONTRIBUTION TO POST-NEWTONIAN WAVEFORM POLARIZATION}
+
 The mode decomposition of gravitational wave polarization is given by
@@
 \end{equation}
-where
+where $_{-2}Y^{lm}$ are spin-weighted spherical harmonics, the angles $(\Theta, \Phi)$ indicates the direction from the source to the observer.
+In a multipolar expansion of the GW field, the modes $h^{lm}$ are related to the radiative mass $U^{lm}$ and the current $V^{lm}$ multipole via
 \begin{equation}
-	h^{lm}= \frac{G}{\sqrt{2}\, R \, c^{l+2}}\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
+	h^{lm}= \frac{G}{\sqrt{2}\, R }\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
 \end{equation}
-
-Here $U^{lm}(T_R)$ and $V^{lm}(T_R)$ are "scalar" mass and current multipoles respectively. The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
+Here $R$ is the distance from source to observer, $T_R$ is retarded time. We assume $c\,=\,G\,=\,1$ for all out calculations.
+ The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
 \begin{equation}
@@
 \end{align}
-where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$.
+where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$. \\
+When the GW field is decomposed into modes as in [ref eqn], the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders.
+\begin{equation}
+	U^{(mem)}_{lm}=32\pi\frac{(l-2)!}{2(l+2)!}\int_{-\infty}^{T_R}dt\int d\Omega\frac{dE_{gw}}{dtd\Omega}(\Omega)Y^*_{lm}(\Omega)
+\end{equation}
+The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by [add ref]
+\begin{equation}
+	\frac{dE_{gw}}{dtd\Omega} = R^2 T^{gw}_{00} = \frac{R^2}{32\pi}\langle\dot{h}^{TT}_{jk}\dot{h}^{TT}_{jk}\rangle =\frac{R^2}{16\pi}\langle\dot{h}^2_{+}+ \dot{h}^2_{\times}\rangle
+\end{equation}
+where the angled bracket mean to average over several wavelengths. The energy flux and be written in terms of $h_{lm}$ modes using equation 2.1 in the following way:
+\begin{widetext}
+\begin{equation}
+		\frac{dE_{gw}}{dtd\Omega}=\frac{R^2}{16\pi}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}\langle\dot{h}_{l'm'}\dot{h}^*_{l''m''}\rangle _{-2}Y^{l'm'}(\theta,\phi)_{-2}Y^{l''m''*}(\theta, \phi)
+\end{equation}
+\end{widetext}
 
-\section{MEMORY CONTRIBUTION TO THE RADIATIVE-MASS MULTIPOLES}
-The GW polarization can be expanded on the basis of spin-weighted spherical harmonics.
-\begin{equation}
-	\sqrt{2}R(h_{+} - ih_{\times}) = \sum_{l=2}^{\infty}\sum_{m=-l}^{l}(U_{lm} - \mathit{i}V_{lm})_2Y^{lm}(\Theta, \Phi)
-\end{equation}
-Here $(R,\Theta,\Phi)$ are the spherical coordinates pointing from the source to the observer. The coefficient $U_{lm}$ and $V_{lm}$ are mass and current multipole moments. \\
-The memory contribution to the waveform polarization can be evaluated in the the following ways as described in [add referece]
-
-\subsection{Computing angular integral}
-
-\begin{equation}
-	\int d\Omega n_{L},
-\end{equation}
+The memory contribution to mass multipole moment can now be evaluated using equation 2.6 and 2.8. The angular part of the integral has the following form
 
@@
 \end{equation}
+
+The above integral can be evaluated as done in ref [add reference]. The result is show in appendix. \\
+The time derivative of the memory mass-multipole moment is defines as [add ref] 
+ $U^{(mem)(1)}_{lm}\equiv\frac{dU^{(mem)}_{lm}}{dT_R}$, which when combined using equation 2.6 and 2.8, can be written as
 
@@
 
-\subsection{Time derivatives of the memory multipole moments}
-
+where $G^{2-20}_{l'm'lm'-m''-m}$ is the angular integral of equation 2.9 and is given in Appendix [add].
+The above equation can be used to compute the first time derivative of mass-multipole moment by substituting the GW modes $h_{lm}$.
+When evaluating equation 2.10, we are only interested in $m=0$ modes.
+The $m\neq=0$ terms yield oscillatory contribution to the waveform polarizations that enter at higher PN orders then the nonoscillatory, $m=0$ terms.
+The non-vanishing, non-oscillatory modes are $U^{(mem)}_{l0}$ with l-even.
+These $U^{(mem)}_{l0}$ consist of time-integral of polynomials in $x\equiv(M\omega)^{2/3}$ (where $\omega$ is the orbital angular frequency), with each terms of the form
 \begin{equation}
@@
 \end{equation}
-
-\subsection{Computing time integral over the past history of the source}
-sesadsd
-
-% 
+After performing the change of variable indicated in above equation 2.12, evaluation of the resulting integrals requires a model for the frequency evolution of the binary.
+The adiabatic evolution of the frequency (or $x$) is easily derived from energy balance ($\mathcal{L}_{GW}=-\dot{E}$).
+and the relation $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$.
+The 3.5PN orbital energy is given by equation C in ref[add aruns paper ref]
+\begin{widetext}
+	\begin{align}\label{eq:E}
+	E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right.
+\nonumber \\
+	& \quad \left.
+x^2\left[-\frac{27}{8} + \frac{19 \, \eta}{8}-\frac{\eta^2}{24} + \eta^2 \left\{\left(\chisSqr-\chiaSqr\right) -3\left[\left(\chisdL\right)^2 - \left(\chiadL\right)^2\right]\right\}\right.\right.
+\nonumber \\
+	& \qquad \left(\frac{1}{2}-\eta\right)\left\{\chisSqr+\chiaSqr-3\left[\left(\chisdL\right)^2+\left(\chiadL\right)^2\right]\right\}+\delta\left\{\chisDchia-3\left[\left(\chisdL\chisdL\right)\right]\right\} \nonumber\\
+	& \quad \left.\left.x^{5/2}\left[\left(8-\frac{121\,\eta}{9} +  \frac{2\,\eta^2}{9}\right)\chisdL + \left(8-\frac{31\,\eta}{9}\right)\delta\chiadL\right] + x^3\left[-\frac{675}{64}+\left(\frac{34445}{576}-\frac{205\,\pi^2}{96}\right)\eta-\frac{155\,\eta^2}{96}-\frac{35\,\eta^3}{5184}\right]
+	\right]
+	\right\}
+	\end{align}
+\end{widetext}
+and the 3PN GW luminosity is given by Eq C7 to C12 in [ref]
+\begin{widetext}
+	\begin{align}\label{eq:F}
+\mathcal{L} &=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right.\right]\nonumber\\
+&\quad \left.\left.
+x^2\left[-\frac{44711}{9072}+\frac{9271\eta}{504}+\frac{65\, \eta^2}{18} + \left(\frac{287}{96} +\frac{\eta}{24}\right)\left(\chisdL\right)^2-\left(\frac{89}{96}+\frac{7\,\eta}{24}\right)\chisSqr\right.\right.\right.\nonumber\\
+&\qquad\left.\left.\left.
++ \left(\frac{287}{96}-12\eta\right)\left(\chiadL\right)^2 + \left(-\frac{89}{96}+4\eta\right)\chiaSqr + \frac{287}{48}\delta\left(\chisdL\right)\left(\chiadL\right) -\frac{89}{48}\delta\left(\chisDchia\right)
+\right]\right.\right.\nonumber\\
+&\quad \left.x^{5/2}\left[\left(-\frac{8191}{672}-\frac{583\,\eta}{24}\right)\pi+\left\{\left(-\frac{59}{16}+\frac{227\, \eta}{9}-\frac{157\, \eta^2}{9}\right)\chisdL + \left(-\frac{59}{16} + \frac{701\,\eta}{36}\delta\chiadL\right)\right\}\right.\right]\nonumber\\
+& \left.
+\quad x^3\left[\frac{6643739519}{69854400}+\frac{16\,\pi^2}{3}-\frac{1712\,\gamma_E}{105}-\frac{856}{105}\log(16x)+\left(-\frac{134543}{7776}+\frac{41\,\pi^2}{48}\right)\eta -\frac{94403\,\eta^2}{3024}-\frac{775\,\eta^3}{324}
+\right]
+\right\}
+	\end{align}
+\end{widetext}
+Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, we get the following result 
 \begin{widetext}
@@
 \end{widetext}
+The above equation can be used to evaluate the integral in Eq. 2.12
 
@@
 \end{align}
-
 \end{widetext}
+We can see that all contribution to $m=0$, even-$l$ modes arise solely from the Christodoulou memory of $U_{lm}$
 

### 2018-12-25 notes/Paper.tex
updated paper

@@
 However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes.
-Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.\\
+Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.
+These $m=0$ modes tend to be much smaller in magnitude and depends more sensitively on	initial conditions than other modes.
+For detailed discussion on this see ref [add ref] \\
+The NR simulations mostly easily compute  the curvature scalar $\Psi_4$ and decompose its values at large $R$ into a sum over spin-weighted spherical harmonic modes:
+\begin{equation}
+	\Psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_\times = \sum_{l=2}^{\infty}\sum_{m=-l}^{l}\psi_{lm}(t, R) _{-2}Y_{lm}(\Theta, \Phi)
+\end{equation} 
+In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$.
+There are other difficulties in calculating memory from numerical relativity data.Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified.
+Choosing these integration constant to be zero [add ref] causes and artificial memory and a slope in the $h_{lm}$ modes.
+This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.\\
+Numerical relativity simulations starts at a finite radius, so although it contains a most part of gravitational wave memory, PN results should be used for early inspiral phase.
+The significance of memory accumulated during the inspiral phase can be seen by looking at the largest memory mode $h_{20}$.
+Consider the leading-(Newtonian)-order piece of the largest memory mode $h_{20}$:
+\begin{equation}
+	h^{NR}_{20}\left(T_R\right) = \frac{1}{\sqrt{2}R}\int_{T_0}^{T_R}U^{(mem)(1)}_{20}(t)dt
+\end{equation}
+where $U^{(mem)(1)}_{20}$ is given by the leading order piece of Eq.
+(3.13a) in ref [add ref], and the $T_0$ is the starting time of the simulation.
+For a realistic, quasicircular binary that inspirals from a large initial separation, $T_0 \rightarrow -\infty$.
+The NR simulations of binaries that starts at a finite initial separation would underestimate the size of the memory.
+For a simulation that starts with an initial separation of $10M$, the error in the memory at a harmonic coordinate separation of $5M$ is $50\%$.
 
-The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata [].
+\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata [].
 However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})
@@
 	\begin{align}\label{eq:E}
-	E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right.
-\nonumber \\
-	& \quad \left.
-x^2\left[-\frac{27}{8} + \frac{19 \, \eta}{8}-\frac{\eta^2}{24} + \eta^2 \left\{\left(\chisSqr-\chiaSqr\right) -3\left[\left(\chisdL\right)^2 - \left(\chiadL\right)^2\right]\right\}\right.\right.
-\nonumber \\
-	& \qquad \left(\frac{1}{2}-\eta\right)\left\{\chisSqr+\chiaSqr-3\left[\left(\chisdL\right)^2+\left(\chiadL\right)^2\right]\right\}+\delta\left\{\chisDchia-3\left[\left(\chisdL\chisdL\right)\right]\right\} \nonumber\\
-	& \quad \left.\left.x^{5/2}\left[\left(8-\frac{121\,\eta}{9} +  \frac{2\,\eta^2}{9}\right)\chisdL + \left(8-\frac{31\,\eta}{9}\right)\delta\chiadL\right] + x^3\left[-\frac{675}{64}+\left(\frac{34445}{576}-\frac{205\,\pi^2}{96}\right)\eta-\frac{155\,\eta^2}{96}-\frac{35\,\eta^3}{5184}\right]
-	\right]
-	\right\}
+	E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right\} \nonumber \\
 	\end{align}
-\end{widetext}
+
 and the 3PN GW luminosity is given by Eq C7 to C12 in [ref]
-\begin{widetext}
+
 \begin{align}\label{eq:F}
 \mathcal{L} &=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right.\right]\nonumber\\
-&\quad \left.\left.
-x^2\left[-\frac{44711}{9072}+\frac{9271\eta}{504}+\frac{65\, \eta^2}{18} + \left(\frac{287}{96} +\frac{\eta}{24}\right)\left(\chisdL\right)^2-\left(\frac{89}{96}+\frac{7\,\eta}{24}\right)\chisSqr\right.\right.\right.\nonumber\\
-&\qquad\left.\left.\left.
-+ \left(\frac{287}{96}-12\eta\right)\left(\chiadL\right)^2 + \left(-\frac{89}{96}+4\eta\right)\chiaSqr + \frac{287}{48}\delta\left(\chisdL\right)\left(\chiadL\right) -\frac{89}{48}\delta\left(\chisDchia\right)
-\right]\right.\right.\nonumber\\
-&\quad \left.x^{5/2}\left[\left(-\frac{8191}{672}-\frac{583\,\eta}{24}\right)\pi+\left\{\left(-\frac{59}{16}+\frac{227\, \eta}{9}-\frac{157\, \eta^2}{9}\right)\chisdL + \left(-\frac{59}{16} + \frac{701\,\eta}{36}\delta\chiadL\right)\right\}\right.\right]\nonumber\\
-& \left.
-\quad x^3\left[\frac{6643739519}{69854400}+\frac{16\,\pi^2}{3}-\frac{1712\,\gamma_E}{105}-\frac{856}{105}\log(16x)+\left(-\frac{134543}{7776}+\frac{41\,\pi^2}{48}\right)\eta -\frac{94403\,\eta^2}{3024}-\frac{775\,\eta^3}{324}
-\right]
-\right\}
 \end{align}
-\end{widetext}
+
 Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, we get the following result
-\begin{widetext}
+
 \begin{align}\label{eq:FbydE}
@@
 + x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
-	&\quad + x^2 \left[ \frac{34103}{18144} + \frac{13661 \, \eta}{2016} + \frac{59 \, \eta^2}{18} + \left( \frac{719}{96}- 30 \eta\right)(\chiadL)^2 +\chiaSqr \left(-\frac{233}{96}+10 \eta \right) \right.
-\nonumber\\
-	&\qquad+\frac{719 \chiadL \, \chisdL \delta
-	}{48}-\frac{233 \chisDchia \delta}{48} \left. + \, (\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
-}{24}+\frac{233}{96}\right) \right] \nonumber \\ 
-&\quad + x^{5/2} \left[-\frac{4159 \, \pi}{672}-\frac{189 \, \pi \, \eta}{8} + \chiadL \delta  \left(-\frac{31319}{1008} + \frac{1159 \, \eta}{24}\right)+\chisdL \left(-\frac{31319}{1008} + \frac{22975 \, \eta}{252} -\frac{79 \, \eta^2}{3}\right) \right]  \nonumber\\
-&\quad \left.
-+ x^3 \left[\frac{16447322263}{139708800} + \frac{16 \, \pi^2}{3} -\frac{1712 \, \gamma_{E}}{105} -\frac{56198689 \, \eta}{217728} + \frac{451 \, \pi^2 \, \eta}{48} + \frac{541 \, \eta^2}{896} -\frac{5605 \, \eta^{3}}{2592} -\frac{80 \,\pi \delta\chiadL }{3} \right.\right.
-\nonumber \\
-&\qquad \left.\left.
-+ \left(\frac{575}{448} + \frac{565 \, \delta^2}{9}-\frac{65815 \, \eta}{4032} + \frac{89 \, \eta^2}{2}\right)(\chiadL)^2 + \left(-\frac{145}{448}+\frac{21985 \,\eta}{4032} -\frac{89 \, \eta^2}{6}\right)\chiaSqr + \right(-\frac{80 \, \pi}{3} + \frac{40 \, \pi \, \eta}{3} \right.
-\nonumber \\
-&\qquad  \left.+\delta\left(\frac{258295}{2016} -\frac{9203 \, \eta}{96}\right)\chiadL\left)\chisdL + \left(\frac{258295}{4032} -\frac{48773 \, \eta}{576} + \frac{3041 \, \eta^2}{144}\right)(\chisdL)^2 + \delta\left(-\frac{145}{224} + \frac{2143 \, \eta}{288}\right)\chisDchia \right.\right.\nonumber \\
-&\qquad\left.\left.+\left(-\frac{145}{448} + \frac{1891 \, \eta}{576} -\frac{7 \, \eta^2}{144}\right)\chisSqr - \frac{3424 \, \log[2]}{105} - \frac{1712 \log[x]}{210}\right]\right\}
 \end{align}
@@
 \begin{align}\label{eq:H20}
-\hat{H}_{20} &= \alpha \frac{5}{15\sqrt{6}}\left\{ 1 + x\left(-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right)+ x^{3/2}\left(\left(-\frac{7}{1440} -\frac{27 \, \delta}{10} + \frac{7 \,\eta}{360}\right)\chiadL+\left(-\frac{27}{10} -\frac{7 \, \delta^3}{1440} + \frac{22 \, \eta}{15}\right)\chisdL\right)\right.
+\hat{H}_{20} &= \alpha \frac{5}{14\sqrt{6}}\left\{ 1 + x\left[-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right]+ x^{3/2}\left[\left(\frac{7}{480} +\frac{27 \, \delta}{10} - \frac{7 \,\eta}{120}\right)\chiadL+\left(\frac{27}{10} +\frac{7 \, \delta^3}{480} - \frac{2 \, \eta}{15}\right)\chisdL\right]\right.
 \nonumber \\
 & \quad \left.
-x^2\left(-\frac{151877213}{67060224} -\frac{123815 \, \eta}{44352} +\frac{205 \, \eta^2}{352} -\left(\frac{1459}{864} -\frac{1441 \, \eta}{216}\right)\chiaSqr - \left(\frac{27 \, \delta}{8} +\frac{\delta^3}{432}\right)\chisDchia - \left(\frac{27}{16}+\frac{\delta^4}{864} -\frac{\eta}{12}\right)\chisSqr\right)\right.
+x^2\left[-\frac{151877213}{67060224} -\frac{123815 \, \eta}{44352} +\frac{205 \, \eta^2}{352} +\left(-\frac{163}{96} +\frac{161 \, \eta}{24}\right)\chiaSqr + \left(-\frac{27 \, \delta}{8} -\frac{\delta^3}{48}\right)\chisDchia + \left(-\frac{27}{16}-\frac{\delta^4}{96} -\frac{\eta}{12}\right)\chisSqr\right]\right.
 \nonumber \\
 & \quad \left.
-x^{5/2}\left(-\frac{253 \, \pi}{336}+ \frac{253 \, \pi \, \eta}{84} -\left(-\frac{631}{96768} + \frac{2729857 \, \delta}{84672} -\frac{146597 \, \pi \, \delta}{1764}+\frac{5 \, \eta}{378}+\frac{3007 \, \delta \, \eta}{1008} -4\pi \, \delta \, \eta +\frac{311 \, \eta^2}{6048}\right)\chiadL\right.\right.
+x^{5/2}\left[-\frac{253 \, \pi}{336}+ \frac{253 \, \pi \, \eta}{84} +\left\{\frac{631}{32256} + \frac{788471\,\delta}{84672} +\left(-\frac{5}{126}-\frac{991\,\delta}{1008}\right)\eta-\frac{311\,\eta^2}{2016}\right\}\chiadL\right.\right.
 \nonumber\\
-& \qquad\left.
--\left(\frac{2729857}{84672}-\frac{146597 \, \pi}{1764}-\frac{631 \, \delta^3 }{96798} -\frac{44111 \, \eta}{1764}+\frac{4852 \, \pi \, \eta}{63}-\frac{311 \, \delta^3 \, \eta}{24192}-\frac{1121 \, \eta^2}{252} +\frac{68 \, \pi \, \eta^2}{7}\right)\chisdL\right) \nonumber \\
-& \quad  x^3\left[-\frac{4397711103307}{532580106240} + \left( \frac{700464542023}{13948526592} - \frac{205 \, \pi^2}{96}\right)\eta + \frac{69527951 \, \eta^2}{166053888} + \frac{1321981 \, \eta^3}{5930496} -  \left(-\frac{160867}{13824} -\frac{791 \, \delta}{27648} + \right.\right.
+& \qquad\left. +\left\{\frac{788471}{84672}-\frac{631\, \delta^3}{32256} -\left(-\frac{36509}{2646} + \frac{311\,\delta^3}{8064}\right)\eta+\frac{853\,\eta^2}{252}\right\}\chisdL\right] \nonumber \\
+& \quad  x^3\left[-\frac{4397711103307}{532580106240} + \left( \frac{700464542023}{13948526592} - \frac{205 \, \pi^2}{96}\right)\eta + \frac{69527951 \, \eta^2}{166053888} + \frac{1321981 \, \eta^3}{5930496} -  \left(-\frac{7\,\pi}{256} + \frac{275\,\pi\,\delta}{12} \right.\right.
 \nonumber \\
 & \qquad \left.
-\frac{7345 \, \delta^2}{576} + \left(\frac{3953767}{96768}+\frac{791 \, \delta}{6912}\right)\eta + \frac{22885 \, \eta^2}{1152}\right)\chiaSqr -\left(-\frac{275 \, \pi}{12} + \frac{7 \, \pi \, \delta^3}{768} + 17 \, \pi \, \eta \right)\chisdL + \left(\frac{47869}{43008}  \right.\nonumber\\
-& \qquad \left.\left.
-- \frac{791 \, \delta^3}{27648} + \frac{743 \, \delta^4}{387072} + \left(-\frac{410453}{16128} + \frac{133 \, \delta^3}{6912} + \frac{11 \, \delta^4}{4608}\right)\eta + \frac{10229 \, \eta^2}{1152}\right)\chisSqr - \left( \frac{7 \, \pi}{768}-\frac{275 \, \pi \, \delta}{12}-\frac{7 \, \pi \, \eta}{192}\right.\right.
-\nonumber \\
-&\qquad \left.\left.\left.
-+ \left(-\frac{791}{27648} + \frac{47869 \, \delta}{21504} + \frac{743 \, \delta^3}{193536} - \frac{791 \, \delta^4}{27648} + \left(\frac{77}{576} - \frac{71749 \, \delta}{2304} + \frac{11 \, \delta^3}{2304}\right)\eta - \frac{133 \, \eta^2}{1728}\right)\chisdL\right)\chisdL
-\right]
-\right\}
+\frac{7 \, \pi\,\delta}{256}\right)\chiadL + \left\{-\frac{42787}{32256} + \frac{791\,\delta}{9216} +\frac{113\,\delta^2}{576} + \left(\frac{354035}{32256} -\frac{791\,\delta}{2304}\right)\eta - \frac{2533\,\eta^2}{128}\right\}\chiaSqr\nonumber\\
+&\qquad \left.\left.
+\left\{\frac{275\,\pi}{12} -\frac{7\,\pi\,\delta^3}{256}-\frac{56\,\pi\eta}{3} + \left(\frac{791}{9216} -\frac{47869\,\delta}{21504}-\frac{743\,\delta^3}{21504} + \frac{791\,\delta^4}{9216} + \left(-\frac{77}{192}+\frac{87269\, \delta}{2304}-\frac{11\, \delta^3}{256}\right)\eta + \frac{133\,\eta^2}{576}\right)\chiadL\right\}\chisdL\right.\right.\nonumber\\
+& \qquad \left.\left.\left\{-\frac{47869}{43008} + \frac{791\,\delta^3}{9216} -\frac{743\,\delta^4}{43008} + \left(\frac{57677}{1792} -\frac{133\,\delta^3}{2304} -\frac{11\,\delta^4}{512}\right)\eta - \frac{15029\,\eta^2}{1152}\right\}\chisSqr
+\right]\right\}
+\end{align}
+
+\begin{align}\label{eq:H40}
+	\hat{H}_{40} &=\alpha\frac{1}{504\sqrt{2}}\left\{1 + x\left[-\frac{180101}{29568} + \frac{27227\,\eta}{1056}\right] + x^{3/2}\left[\left(\frac{7}{40} + \frac{27\,\delta}{10} -\frac{7\,\eta}{10}\right)\chiadL+\left(\frac{27}{10}+\frac{7\,\delta^3}{40} +\frac{46\,\eta}{5}\right)\chisdL\right]\right.\nonumber\\
+	&\qquad x^{5/2}\left[-\frac{13565\,\pi}{1232} + \frac{13565\,\pi\,\eta}{308} + \left\{\frac{379}{2688}-\frac{1457321\,\delta}{68992}+\left(-\frac{59}{336} + \frac{3094367\,\delta}{22176}\right)\eta -\frac{87\,\eta^2}{56}\right\}\chiadL\right.\nonumber\\
+	&\qquad\qquad \left.
+\left\{-\frac{1457321}{68992} + \frac{379\,\delta^3}{2688} + \left(\frac{5892749}{38808} + \frac{87\,\delta^3}{224}\eta-\frac{469813\,\eta^2}{5544}\right)\right\}\chisdL\right]\nonumber\\
+	&\quad\qquad x^2\left[\frac{2201411267}{158505984} -\frac{34829479\,\eta}{432432} + \frac{844951\,\eta^2}{27456} + \left(-\frac{85}{48} + 7\,\eta\right)\chiaSqr + \left(-\frac{27\, \delta}{8} -\frac{\delta^3}{6}\right)\chisDchia + \left(-\frac{27}{16} -\frac{\delta^4}{12} + \frac{\eta}{12}\right)\chisSqr\right].\nonumber\\
+	&\quad\qquad x^3\left[\frac{167644780815461}{8592292380672}+\left(-\frac{11327099812895}{306867585024} -\frac{205\,\pi^2}{96}\right)\eta -\frac{137763118975\,\eta^2}{1217728512} + \frac{2435519695\,\eta^3}{130470912}\right.\nonumber\\
+	&\qquad\qquad  + \left(-\frac{21\,\pi}{64} + \frac{275\,\pi\,\delta}{12} + \frac{21\,\pi\,\eta}{16}\right)\chiadL + \left(\frac{32394371}{2838528} + \frac{791\,\delta}{768} + \frac{113\,\eta^2}{576} + \left(-\frac{71832779}{709632} -\frac{791\,\delta}{192}\right)\eta + \frac{59317\,\eta^2}{264}\right)\chiaSqr .\nonumber\\
+	&\qquad\qquad + \left\{\frac{275\,\pi}{12} - \frac{21\,\pi\,\delta^3}{64} -\frac{91\,\pi\,\eta}{3} + \left(\frac{791}{768} + \frac{11114513\,\delta}{473088} -\frac{743\,\delta^3}{2688} + \frac{791\,\delta^4}{768} + \left(-\frac{77}{16} -\frac{649327\,\delta}{16896} -\frac{11\,\delta^3}{32}\right)\eta  \right.\right.\nonumber\\
+	&\qquad\qquad \left.\left.\left.
+\left.+ \frac{133\,\eta^2}{48}\right)\chiadL\right\}\chisdL + \left(\frac{11114513}{946176} + \frac{791\,\delta^3}{768} -\frac{743\,\delta^4}{5376} + \left(\frac{2007451}{118272} -\frac{133\,\delta^3}{192} -\frac{11\,\delta^4}{64}\right)\eta - \frac{82411\,\eta^2}{2816}\right)\chisdL\right]\nonumber
+	\right\}\\
+\end{align}
+
+\begin{align}\label{eq:H60}
+	\hat{H}_{60} &=-\alpha\frac{4195}{1419264\sqrt{273}}x\left\{1-\frac{3612\,\eta}{839}+ x\left[-\frac{45661561}{6342840} + \frac{101414\,\eta}{2517} - \frac{48118\,\eta^2}{839}\right]\right.
++ x^{3/2} \left[\frac{1248\,\pi}{839} -\frac{4992\,\pi\,\eta}{839}\right.\nonumber\\
+	&\qquad + \left(-\frac{44}{5873} + \frac{14129\,\delta}{2517} + \left(\frac{264}{5873} - \frac{140148\,\delta}{5873}\right)\eta -\frac{352\,\eta^2}{5873}\right)\chiadL + \left(\frac{14129}{2517} -\frac{44\,\delta^3}{5873}\right.\nonumber\\
+	&\quad\qquad + \left.\left.\left(-\frac{79120}{2517} + \frac{88\,\delta^3}{5873}\right)\eta + \frac{161136\,\eta^2}{5873}\right)\chisdL\right] + x^{5/2}\left[\frac{66616409\,\pi}{2718360} -\frac{874693\,\pi\,\eta}{7551} + \frac{239134\,\pi\,\eta^2}{2517}\right.
++ \nonumber\\
+	&\qquad\qquad \left(-\frac{8173}{634248} -\frac{1240441999\,\delta}{57085560} + \left(\frac{6479}{105714} + \frac{15116383\,\delta}{105714}\right)\eta + \left(-\frac{1100}{158571}-\frac{2406356\,\delta}{7551}\right)\eta^2-\frac{968\,\eta^3}{7551}\right)\chiadL\nonumber\\
+	&\qquad\qquad \left.
+\left(-\frac{1240441999}{57085560} -\frac{8173\,\delta^3}{634284} + \left(\frac{2016684797}{14271390} + \frac{3091\,\delta^3}{317142}\right)\eta + \left(-\frac{19243010}{52857} + \frac{242\,\delta^3}{7551}\right)\eta^2 + \frac{1891144\,\eta^3}{7551}\right)\chisdL\right]\nonumber\\
+	&\qquad x^2\left[\frac{33133408041593}{1594133291520}-\frac{912563795039\,\eta}{25102880} + \frac{8054229309\,\eta^2}{25102880}-\frac{1363643519\,\eta^3}{18827160} + \left(-\frac{81}{32} + \frac{140263\,\eta}{6712}-\frac{36120\,\eta^3}{839}\right)\chiaSqr\right.
+\nonumber\\
+	& \quad\qquad \left.+ \left(-\frac{81\,\delta}{16} + \frac{73143\,\delta\,\eta}{3356}\right)\chisDchia + \left(-\frac{81}{32} + \frac{36991\,\eta}{3356} - \frac{7047\,\eta^2}{1678}\right)\chisSqr\right]
+\end{align}
+
+\begin{align}\label{eq:H80}
+	\hat{H}_{80} &=\alpha\frac{75601}{213497856\sqrt{119}}x^2\left[1-\frac{452070\,\eta}{75601} + \frac{733320\,\eta^2}{75601} + x\left(-\frac{265354303}{33869248} + \frac{18177343651\,\eta}{321757856} - \frac{5057483259\,\eta^2}{40219732} + \frac{1828935981\,\eta^3}{20109866}\right)\right]\nonumber\\
+\end{align}
+\begin{align}\label{eq:H80}
+\hat{H}_{100} &=-\alpha\frac{525221\,x^3}{6452379648\sqrt{154}}\left(1 - \frac{79841784\,\eta}{9979199} + \frac{198570240\,\eta^2}{9979199} -\frac{172307520\,\eta^3}{9979199}\right)
 \end{align}
 \end{widetext}
-We can see that all contribution to $m=0$, even-$l$ modes arise solely from the Christodoulou memory of $U_{lm}$
+We can see that all contribution to $m=0$, even-$l$ modes arise solely from the Christodoulou memory piece of $U_{lm}$.
+There are no hereditary-memory contribution to the radiative current multipole moments $V_{lm}$\\
+We can combine the above with Eq 2.16 and explicitly compute the memory contribution to the + waveform polarization.This can be done in the following way
+\begin{equation}
+	h_{+,\times} = \frac{2\eta Mx}{R}H_{+,\times} + O\left(\frac{1}{R^2}\right)
+\end{equation}
 
@@
 This can lead to an uncontrollable non-linear drift if the integration is performed in the time domain.
+The unfiltered GW waveform is shown in fig [add fig], which clearly shows an oscillating waveform which settles to a nonzero value.
+The method described in ref[add ref] gives us a way to get rid of this unphysical behaviour
 \par The memory contribution to $h_{+}$ is given by
@@
 \appendix 
+\begin{widetext}
+
+\section{Angular Integral Of The Triple Product Of Spin-Weighted Spherical Harmonics}
+The following integral appears in the evaluation of nonlinear memory contribution to the radiative mass-multipole moments $U_{lm}$. Equation 2.4 can be rewritten as
+\begin{equation}
+	d^{\,l_j}_{m_js_j}(\Theta)=\sum_{k_j=k_i(j)}^{k_f(j)}g_j(k_j)\left(\sin\frac{\Theta}{2}\right)^{p_j}\left(\cos\frac{\Theta}{2}\right)^{2l_j-p_j}
+\end{equation}
+where
+\begin{equation}
+	g_j(k_j)=\frac{(-1)^k_j[(l_j + m_j)!(l_j-m_j)!(l_j+s_j)!(l_j-s_j)!]^{1/2}}{k_j!(l_j+m_j-k_j)!(l_j-s_j-k_j)!(s_j-m_j+k_j)!},
+\end{equation}
+$p_j=2k_j+s_j-m_j$, $k_i=\max(0, m-s)$ and $k_f=\min(l+m, l-s)$. The index $j=1,2,3$ serves only the three harmonics. The required integral can be written as:
+
+\begin{align}
+	G^{s_1 s_2 s_3}_{l_1 l_2 l_3 m_1 m_2 m_3}&\equiv\int\, _{-s1}Y^{l_1m_1}(\Theta, \Phi)_{-s2}Y^{l_2m_2}(\Theta, \Phi)_{-s3}Y^{l_3m_3}(\Theta, \Phi)d\Omega \nonumber \\
+	&=(-1)^{s_1+s_2+s_3}\frac{[(2l_1+1)(2l_2+1)(2l_3+1)]^{1/2}}{(4\pi)^{3/2}}\int_{0}^{2\pi}e^{\mathit{i}(m_1+m_2+m_3)\Phi}d\Phi\int_{0}^{\pi}d^{l_1}_{m_1s_1}d^{l_2}_{m_2s_2}d^{l_3}_{m_3s_3}\sin(\Theta)d\Theta
+\end{align}
+The $\Phi$ integral is 
+\begin{equation}
+	\int_{0}^{2\pi}e^{\mathit{i}(m_1+m_2+m_3)\Phi}d\Phi = 2\pi\delta^{-m_1}_{m_2+m_3}
+\end{equation}
+The $\Phi$ integral can be written as 
+\begin{equation}
+	\int_{0}^{\pi}d^{l_1}_{m_1s_1}d^{l_2}_{m_2s_2}d^{l_3}_{m_3s_3}\sin\Theta d\Theta = 2\sum_{k_1k_2k_3}g_1(k_1)g_2(k_2)g_3(k_3)\int_{0}^{\pi}\left(\sin\frac{\Theta}{2}\right)^{2a-1}\left(\cos\frac{\Theta}{2}\right)^{2b-1}d\Theta
+\end{equation}
+where $a=1+(p_1+p_2+p_3)/2$ and $b=1+l_1+l_2+l_3-(p_1+p_2+p_3)/2$. The $\Phi$ integral is expression in terms of the Beta or Gamma function and can be found in standard tables.
+\begin{equation}
+	\int_{0}^{\pi}\left(\sin\frac{\Theta}{2}\right)^{2a-1}\left(\cos\frac{\Theta}{2}\right)^{2b-1}d\Theta=\frac{\Gamma(a)\Gamma(b)}{\Gamma(a+b)}=B(a,b)
+\end{equation}
+
+\section{Calculation Of The $\dot{h}_{lm}$ Modes}
+The $h_{lm}$ modes are written as 
+\begin{equation}
+	h_{lm}=8\sqrt{\frac{\pi}{5}}\frac{\eta\,M}{R}x\hat{H}_{lm}(x)e^{\mathit{i}m\psi}
+\end{equation}
+where the $\hat{H}_{lm}$ given in ref[add ref] for the case of precessing binaries.
+In this paper we are interested in nonprecessing case, so we rewrite the expressions such that spin are aligned with the total orbital angular momentum.
+The phase variable $\psi$ is related to the orbital phase $\phi$ via
+\begin{equation}
+	\psi=\varphi - 3x^{3/2}\left[1-\frac{\eta}{2}x\right]\ln\left(\frac{x}{x_0}\right)
+\end{equation}
+where $\ln x_0 = \frac{11}{18} - \frac{2}{3}\gamma_E -\frac{4}{3}\ln2 + \frac{2}{3}\ln\left(\frac{M}{r_0}\right)$ is related to the arbitrary constant $r_0$ appearing in the coordinate transformation of Eq.
+(2.20) or ref[add ref].
+The orbital phase can be expressed a function of x for nonprecessing binaries, to 3.5PN order.
+\begin{align}
+\varphi(x)& = -\frac{1}{32\,\eta\,x^{5/2}}\left\{1+x\left[\frac{55\eta}{12}+\frac{3715}{1008}\right]+x^{3/2}\left[\frac{565}{24}\left(\left(1-\frac{76\eta}{133}\right)\chisdL + \delta\chiadL\right)-10\pi\right]\right.\nonumber\\
+&\quad + x^2\left[\left(\chiadL\right)^2\left(150\eta-\frac{3595}{96}\right)-\frac{3595\chiadL\chisdL\delta}{48} + \chiaSqr\left(\frac{1165}{96}- 50\eta\right) + \frac{1165\chisDchia\delta}{\delta}\right]\nonumber\\
+&\qquad + \left.\left(\chisdL\right)^2\left(-\frac{5\eta}{24} - \frac{3595}{96}\right) + \chisSqr\left(\frac{35\eta}{24}+\frac{1165}{96}\right) + \frac{3085\eta^2}{144} + \frac{27145\eta}{1008} + \frac{15293365}{10160664}\right]\nonumber\\
+&\quad +  \frac{x^{5/2}}{2}\left[\left(\chiadL\left(-\frac{35\delta\eta}{2} - \frac{732985\delta}{2016}\right) + \chisdL\left(\frac{85\eta^2}{2} + \frac{6065\eta}{18} - \frac{732985}{2016}\right) -\frac{65\pi\eta}{8} + \frac{38645\pi}{672}\right)\ln(x)\right]\nonumber\\
+&\quad x^3\left[-\frac{127825\eta^3}{5184} + \frac{76055\eta^2}{6912} + \frac{2255\pi^2\eta}{48} - \frac{15737765635\eta}{12192768} - \frac{1712\gamma_E}{21} -\frac{160\pi^2}{3} + \frac{12348611916451}{18776862720} - \frac{1712\ln(16x)}{42}\right]\nonumber\\
+&\quad \left.x^{7/2}\left[-\frac{74045\pi\eta^2}{6048} + \frac{378515\pi\eta}{12096} + \frac{77096675\pi}{2032128}\right]\right\}
+\end{align} 
+where $\varphi_0$ is a certain reference phase.
+
+\section{Complete 3.5 PN expressions}
+
+\begin{align}\label{eq:Ef}
+E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right.
+\nonumber \\
+& \quad \left.
+x^2\left[-\frac{27}{8} + \frac{19 \, \eta}{8}-\frac{\eta^2}{24} + \eta^2 \left\{\left(\chisSqr-\chiaSqr\right) -3\left[\left(\chisdL\right)^2 - \left(\chiadL\right)^2\right]\right\}\right.\right.
+\nonumber \\
+& \qquad \left(\frac{1}{2}-\eta\right)\left\{\chisSqr+\chiaSqr-3\left[\left(\chisdL\right)^2+\left(\chiadL\right)^2\right]\right\}+\delta\left\{\chisDchia-3\left[\left(\chisdL\chisdL\right)\right]\right\} \nonumber\\
+& \quad \left.\left.x^{5/2}\left[\left(8-\frac{121\,\eta}{9} +  \frac{2\,\eta^2}{9}\right)\chisdL + \left(8-\frac{31\,\eta}{9}\right)\delta\chiadL\right] + x^3\left[-\frac{675}{64}+\left(\frac{34445}{576}-\frac{205\,\pi^2}{96}\right)\eta-\frac{155\,\eta^2}{96}-\frac{35\,\eta^3}{5184}\right]
+\right]
+\right\}
+\end{align}
+
+\begin{align}\label{eq:Ff}
+\mathcal{L}&=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right.\right]\nonumber\\
+&\quad \left.\left.
+x^2\left[-\frac{44711}{9072}+\frac{9271\eta}{504}+\frac{65\, \eta^2}{18} + \left(\frac{287}{96} +\frac{\eta}{24}\right)\left(\chisdL\right)^2-\left(\frac{89}{96}+\frac{7\,\eta}{24}\right)\chisSqr\right.\right.\right.\nonumber\\
+&\qquad\left.\left.\left.
++ \left(\frac{287}{96}-12\eta\right)\left(\chiadL\right)^2 + \left(-\frac{89}{96}+4\eta\right)\chiaSqr + \frac{287}{48}\delta\left(\chisdL\right)\left(\chiadL\right) -\frac{89}{48}\delta\left(\chisDchia\right)\right]\right.\right.\nonumber\\
+&\quad \left.x^{5/2}\left[\left(-\frac{8191}{672}-\frac{583\,\eta}{24}\right)\pi+\left\{\left(-\frac{59}{16}+\frac{227\, \eta}{9}-\frac{157\, \eta^2}{9}\right)\chisdL + \left(-\frac{59}{16} + \frac{701\,\eta}{36}\delta\chiadL\right)\right\}\right.\right]\nonumber\\&\quad \left.
+x^3\left[\frac{6643739519}{69854400}+\frac{16\,\pi^2}{3}-\frac{1712\,\gamma_E}{105}-\frac{856}{105}\log(16x)+\left(-\frac{134543}{7776}+\frac{41\,\pi^2}{48}\right)\eta -\frac{94403\,\eta^2}{3024}-\frac{775\,\eta^3}{324}
+\right]\right\}
+\end{align}
+
+\begin{align}\label{eq:FbydE}
+\frac{dx}{dt}&= \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
++ x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
+&\quad + x^2 \left[ \frac{34103}{18144} + \frac{13661 \, \eta}{2016} + \frac{59 \, \eta^2}{18} + \left( \frac{719}{96}- 30 \eta\right)(\chiadL)^2 +\chiaSqr \left(-\frac{233}{96}+10 \eta \right) \right.
+\nonumber\\
+&\qquad+\frac{719 \chiadL \, \chisdL \delta}{48}-\frac{233 \chisDchia \delta}{48} \left. + \, (\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
+}{24}+\frac{233}{96}\right) \right] \nonumber \\ 
+&\quad + x^{5/2} \left[-\frac{4159 \, \pi}{672}-\frac{189 \, \pi \, \eta}{8} + \chiadL \delta  \left(-\frac{31319}{1008} + \frac{1159 \, \eta}{24}\right)+\chisdL \left(-\frac{31319}{1008} + \frac{22975 \, \eta}{252} -\frac{79 \, \eta^2}{3}\right) \right]  \nonumber\\
+&\quad \left.
++ x^3 \left[\frac{16447322263}{139708800} + \frac{16 \, \pi^2}{3} -\frac{1712 \, \gamma_{E}}{105} -\frac{56198689 \, \eta}{217728} + \frac{451 \, \pi^2 \, \eta}{48} + \frac{541 \, \eta^2}{896} -\frac{5605 \, \eta^{3}}{2592} -\frac{80 \,\pi \delta\chiadL }{3} \right.\right.
+\nonumber \\
+&\qquad \left.\left.
++ \left(\frac{575}{448} + \frac{565 \, \delta^2}{9}-\frac{65815 \, \eta}{4032} + \frac{89 \, \eta^2}{2}\right)(\chiadL)^2 + \left(-\frac{145}{448}+\frac{21985 \,\eta}{4032} -\frac{89 \, \eta^2}{6}\right)\chiaSqr + \right(-\frac{80 \, \pi}{3} + \frac{40 \, \pi \, \eta}{3} \right.
+\nonumber \\
+&\qquad  \left.+\delta\left(\frac{258295}{2016} -\frac{9203 \, \eta}{96}\right)\chiadL\left)\chisdL + \left(\frac{258295}{4032} -\frac{48773 \, \eta}{576} + \frac{3041 \, \eta^2}{144}\right)(\chisdL)^2 + \delta\left(-\frac{145}{224} + \frac{2143 \, \eta}{288}\right)\chisDchia \right.\right.\nonumber \\
+&\qquad\left.\left.+\left(-\frac{145}{448} + \frac{1891 \, \eta}{576} -\frac{7 \, \eta^2}{144}\right)\chisSqr - \frac{3424 \, \log[2]}{105} - \frac{1712 \log[x]}{210}\right]\right\}
+\end{align}
 
 \bibliography{Spin}
-
+\end{widetext}
 \end{document}

### 2018-12-27 scripts/Make_memoryCalculationplots_AlgnedSpin_v3.py
updated scripts

@@ -4,7 +4,7 @@ import plotsettings
 from matplotlib.font_manager import FontProperties
 from scipy.integrate import odeint
 
-#memory expressions from Favata's paper
+#memory expressions from Favata's paper(our results when spin included)
 def H0mem(theta):
 	H0 = (1.0/96.0)*(np.sin(theta)**2)*(17.0 + (np.cos(theta)**2))
 	return H0
@@ -13,12 +13,19 @@ def H1mem(theta, eta):
 	H1 = (np.sin(theta)**2)*(-(354241.0/2064384.0)-(62059.0/1032192.0)*(np.cos(theta)**2)-(4195.0/688128.0)*(np.cos(theta)**4) + \
 		  ((15607.0/73728.0) + (9373.0/36864.0)*(np.cos(theta)**2) + (215.0/8192.0)*(np.cos(theta)**4))*eta)
 	return H1
+	
+def H1p5(theta, eta, chiZa, chiZs):
+	H1p5 = (np.sin(theta)**2)*(((3.0/1280.0) + (153.0/320.0)*np.sqrt(1.0 - 4*eta) - (3.0*eta/320.0) + (7.0/3840.0)*np.cos(theta)**2 + (9.0/320.0)*np.sqrt(1.0-4.0*eta)*np.cos(theta)**2\
+		-(7.0*eta/960.0)*np.cos(theta)**2)*chiZa + ((153.0/320.0) + (3.0*np.sqrt(1.0 - 4.0*eta)/1280.0) - (3.0*eta/80.0) - (3.0*np.sqrt(1.0-4.0*eta)/320.0)*eta \
+		+ (9.0/320.0)*np.cos(theta)**2  + (7.0*np.sqrt(1.0-4.0*eta)/3840.0)*np.cos(theta)**2 + (23.0/240.0)*eta*np.cos(theta)**2 -(7.0*np.sqrt(1.0-4.0*eta)/960.0)*eta\
+		*np.cos(theta)**2)*chiZs) 	
+	return H1p5
 
 def H2mem(theta, eta):
-	H2 = (np.sin(theta)**2)*(-(3968456539.0/9364045824.0) + (570408173.0/4682022912.0)*(np.cos(theta)**2) + \
[…]

### 2018-12-27 scripts/plot_h22mode.py

@@ -68,8 +68,8 @@ plt.plot(time,d2h22_real_by_dt2,'k' )
 plt.show()
 
 #Subtract linearfit
-linear_fit_coeff_real = np.polyfit(time, h22_real, 1)
-linear_fit_coeff_imag = np.polyfit(time, h22_imag, 1)
+linear_fit_coeff_real = np.polyfit(time, h22_real, 3)
+linear_fit_coeff_imag = np.polyfit(time, h22_imag, 3)
 
 linear_fit_real=linear_fit_coeff_real[0]*time + linear_fit_coeff_real[1]
 plt.plot(time, h22_real-linear_fit_real,'r--', label="linear fit sub from h22")

### 2018-12-27 scripts/plotsettings.py

@@ -7,11 +7,11 @@ rc('font', serif='times')
 rc('mathtext', default='sf')
 rc("lines", markeredgewidth=0.6)
 rc("lines", linewidth=2)
-rc('axes', labelsize=10) #24
+rc('axes', labelsize=18) #24
 rc("axes", linewidth=0.5) #2)
-rc('xtick', labelsize=8)
-rc('ytick', labelsize=8)
-rc('legend', fontsize=10) #16
+rc('xtick', labelsize=20)
+rc('ytick', labelsize=20)
+rc('legend', fontsize=16) #16
 rc('xtick.major', pad=8) #8)
 rc('ytick.major', pad=8) #8)
 rc('xtick.major', size=13) #8)

### 2019-01-07 notes/Paper.tex
updated Latex document

@@
 
-\title{Spin effect on Christodoulou memory for merging blackhole binaries}
+\title{Effect of Black Hole Spin on Christodoulou Memory}
 \author{Ashok Choudhary}\email{aschoudhary@mix.wvu.edu}
@@
 
-Gravitational wave (GW) memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it.
+Gravitational wave (GW) memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it~\cite{Abbott_2016nmj}.
 The are two kinds of gravitational wave memory: linear and nonlinear.
@@
 We can combine the above with Eq 2.16 and explicitly compute the memory contribution to the + waveform polarization.This can be done in the following way
+\begin{widetext}
 	\begin{equation}
@@
 \end{equation}
-
+\begin{equation}
+	H_{+,\,\times}= \sum_{n=0}^{\infty}x^{n/2}H_{+,\,\times}^{(n/2)}
+\end{equation}
+The memory contribution to $H_{+}^{n/2} $ are:
+\begin{subequations}
+\begin{align}
+	H_{+}^{(0,mem)}&=\alpha\frac{1}{96}s_{\Theta}^2\left(17 + c_{\Theta}^2\right),\\
+	H_{+}^{(0.5,mem)}&=0,\\
+	H_{+}^{(1,mem)}&=\alpha s_{\Theta}^2\left[-\frac{354241}{2064384} - \frac{62059}{1032192}c_{\Theta}^2 - \frac{4195}{688128}c_{\Theta}^4 + \left(\frac{15607}{73728} + \frac{9373}{36864}c_{\Theta}^2 +\frac{215}{8192}c_{\Theta}^4\right)\right],\\
+	H_{+}^{(1.5,mem)}&=\alpha s_{\Theta}^2\left\{\left[\frac{3}{1280}  + \frac{7}{3840}c_{\Theta}^2 +  \delta\left(\frac{153}{320} + \frac{9}{320}c_{\Theta}^2\right) +\eta \left(-\frac{3}{320}-\frac{7}{960}c_{\Theta}^2\right)\right]\chiadL\right.\nonumber\\
+	& \quad+\left.\left[\left(\frac{153}{320} + \frac{9}{320}c_{\Theta}^2\right) + \delta\left(\frac{3}{1280} + \frac{7}{3840}c_{\Theta}^2\right) + \eta\left(-\frac{3}{80} +\frac{23}{240}c_{\Theta}^2\right) + \delta\eta\left(-\frac{3}{320}-\frac{7}{960}\right)\right]\chisdL\right\}\\
+	H_{+}^{(2,mem)} &=\alpha s_{\Theta}^2\left\{-\frac{3968456539}{9364045824} + \frac{570408173}{4682022912}c_{\Theta}^2 + \frac{122166887}{3121348608}c_{\Theta}^4 + \frac{7560}{15925248}c_{\Theta}^6 \right.\nonumber\\
+	& \quad + \eta\left[-\frac{7169749}{18579456} - \frac{13220477}{18579456}c_{\Theta}^2 -\frac{1345405}{6193152}c_{\Theta}^4 - \frac{25115}{884736}c_{\Theta}^6\right] + \eta^2\left[\frac{10097}{147456}\right]
+\end{align}
+\end{subequations}
+\end{widetext}
 \section{Memory calculation using Numerical relativity data}
@@
 
-\begin{equation}
+\begin{equation}\label{3.2}
 h_{+}^{(mem)} \approx \frac{R}{192 \, \pi}s_{\Theta}^{2}\left(17 +c_{\Theta}^{2} \right)\int_{-\infty}^{T_R}|\dot{h}_{22}|^2 dt
@@
 The Bondi News is determined by
-\begin{equation}
+\begin{equation}\label{3.3}
 \mathcal{N} = \int_{-\infty}^{t}dt' \psi_4
 \end{equation}
-
+The gravitational wave memory contribution to $h_{+}$ polarization can be computed from this bondi news, by integrating the absolute value of bondi news as shown in equation \ref{hmem}.
+The memory contribution evaluated using Eq.\ref{hmem} will contain contribution from numerical noise, which is unphysical and must be removed from the data.
 \subsection{Integration of finite length signals in frequency domain}
-
+The unphysical part of mermory essential lies in low frequency. If we remove this low frequency part from data, we can get rid of memory contribution due to numerical noise. 
 Consider the Fourier transform, $\mathcal{F}$, applied to an absolutely integrable function $f(t)$,
@@
 \end{equation}
+The inverse fourier transform of above expression is given by
+\begin{equation}\label{3.6}
+	\int_{-\infty}^{t}dt'f(t')=\mathcal{F}^{-1}\left[-\mathit{i}\frac{\tilde{f}(\omega)}{\omega}\right] = -\frac{\mathit{i}}{2\pi}\int_{-\infty}^{\infty}\frac{1}{\omega}e^{\mathit{i\omega t}}\tilde{f}(\omega)d\omega
+\end{equation}
+In the frequency-domain, time integration becomes a simple division by the frequency.
+Above method is particularly susceptible to low-frequency error, under-resolved high-frequency modes can be aliased into low-frequency modes of the signal.
+We try to remove these unphysical frequencies by the method described below
+\subsubsection{Fixed frequency integration ($\mathit{FFI}$)}\label{FFI}
+The effect of the spurious low-frequency modes caused by spectral leakage or aliasing effects, can be significantly suppressed by use of signal filter.
+The time integral can be performed in frequency domain as shown in equation \ref{3.6}.
+The low frequencies which cause the unphysical nonlinear drift can be removed by choosing a cutoff frequency while taking a Fourier transform.
 
-\subsubsection{Fixed frequency integration ($\mathit{FFI}$)}
-
-We evaluate the integral using the following mathod [add ref]
-
-\[
-\tilde{F} =
+\[\tilde{F} =
 \begin{cases}
@@
 -\mathit{i} \frac{\tilde{f}(\omega)}{\omega} & \text{ $\omega>\omega_0$} 
+\end{cases}\]
 
-\end{cases}
-\]
+We use this method to integrate $\psi$ to obtain Bondi News as shown in equation \ref{3.3}.
+
+\subsection{Memory calculation from SXS Rumerical Relativity data}
+The publicly available SXS catalog provides data for Newman-Penrose scalar component $\psi_4$, decomposed into spin weighted spherical harmonics, extrapolated to infinite radius.
+We used this data and  performed a frequency domain time integration as described in equation \ref{3.6}, while removing the unphysical frequency by choosing a lower cutoff as show in section \ref{FFI}.
+The cutoff frequency in chosen as $\omega_0/10$ where $\omega_0$ is the instantaneous GW angular frequency.
+The first begin separating the complex Newman-Perose scalar $\psi_4$ data into phase and amplitude part.
+We compute instantaneous GW angular frequency from this by taking the first time derivative of phase corresponding to $\psi_4$.
+The instantaneous GW frequency often looks noisy at the beginning of the simulations due to numerical issues.
+We remove this noise data choose a lower and upper cutoff time.
+Once we find the upper and lower cutoff time, we can apply a frequency domain time integration for $\psi_4$ by choosing a cutoff frequency.
+After filtering the noise creating frequency we can use equation \ref{3.2} to compute the memory contribution to "plus" polarization of GW signal.
+The SXS catalog gives data for different parameter values like mass ratio, dimensionless spin parameter etc.
+The Christodoulou memory computed from Numerical Relativity data starts from zeros using equation \ref{3.2} due to finite simulation time.
+Chrostodoulou memory is hereditary, hence it's entire past should be included.
+This can be done by combining the post-Newtonian expression during early inspiral and numerical relativity results for merger part.
+The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{figure}
-
+\begin{figure}
+	\includegraphics[width=3.0in]{../plots/PNmemoryContributiontoHplus.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemoryContributionfunofTheta.pdf}
+	\caption{NonSpinning case}
+	\label{fig:0p95}
+\end{figure}
 \newpage
@@
 \right]
-\right\}
-\end{align}
-
-\begin{align}\label{eq:Ff}
-\mathcal{L}&=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right.\right]\nonumber\\
-&\quad \left.\left.
-x^2\left[-\frac{44711}{9072}+\frac{9271\eta}{504}+\frac{65\, \eta^2}{18} + \left(\frac{287}{96} +\frac{\eta}{24}\right)\left(\chisdL\right)^2-\left(\frac{89}{96}+\frac{7\,\eta}{24}\right)\chisSqr\right.\right.\right.\nonumber\\
-&\qquad\left.\left.\left.
-+ \left(\frac{287}{96}-12\eta\right)\left(\chiadL\right)^2 + \left(-\frac{89}{96}+4\eta\right)\chiaSqr + \frac{287}{48}\delta\left(\chisdL\right)\left(\chiadL\right) -\frac{89}{48}\delta\left(\chisDchia\right)\right]\right.\right.\nonumber\\
-&\quad \left.x^{5/2}\left[\left(-\frac{8191}{672}-\frac{583\,\eta}{24}\right)\pi+\left\{\left(-\frac{59}{16}+\frac{227\, \eta}{9}-\frac{157\, \eta^2}{9}\right)\chisdL + \left(-\frac{59}{16} + \frac{701\,\eta}{36}\delta\chiadL\right)\right\}\right.\right]\nonumber\\&\quad \left.
-x^3\left[\frac{6643739519}{69854400}+\frac{16\,\pi^2}{3}-\frac{1712\,\gamma_E}{105}-\frac{856}{105}\log(16x)+\left(-\frac{134543}{7776}+\frac{41\,\pi^2}{48}\right)\eta -\frac{94403\,\eta^2}{3024}-\frac{775\,\eta^3}{324}
-\right]\right\}
-\end{align}
-
-\begin{align}\label{eq:FbydE}
-\frac{dx}{dt}&= \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
+\right\}\nonumber\\
+\\
+\label{eq:Flux}
+\mathcal{L}&=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right]\right.\nonumber\\
+&\quad  x^2\left[-\frac{44711}{9072}+\frac{9271\eta}{504}+\frac{65\, \eta^2}{18} + \left(\frac{287}{96} +\frac{\eta}{24}\right)\left(\chisdL\right)^2-\left(\frac{89}{96}+\frac{7\,\eta}{24}\right)\chisSqr\right.\nonumber\\
+&\qquad\left.
++ \left(\frac{287}{96}-12\eta\right)\left(\chiadL\right)^2 + \left(-\frac{89}{96}+4\eta\right)\chiaSqr + \frac{287}{48}\delta\left(\chisdL\right)\left(\chiadL\right) -\frac{89}{48}\delta\left(\chisDchia\right)\right]\nonumber\\
+&\quad\left.x^{5/2}\left[\left(-\frac{8191}{672}-\frac{583\,\eta}{24}\right)\pi+\left\{\left(-\frac{59}{16}+\frac{227\, \eta}{9}-\frac{157\, \eta^2}{9}\right)\chisdL + \left(-\frac{59}{16} + \frac{701\,\eta}{36}\delta\chiadL\right)\right\}\right.\right]\nonumber\\&\quad \left.x^3\left[\frac{6643739519}{69854400}+\frac{16\,\pi^2}{3}-\frac{1712\,\gamma_E}{105}-\frac{856}{105}\log(16x)+\left(-\frac{134543}{7776}+\frac{41\,\pi^2}{48}\right)\eta -\frac{94403\,\eta^2}{3024}-\frac{775\,\eta^3}{324}
+\right]\right\}\nonumber\\
+\\
+\label{eq:dxdt}
+\frac{dx}{dt}&=\frac{64\, x^5  \, \eta}{5}\left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
 + x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\
@@
 \end{align}
+\end{widetext}
 
-\bibliography{Spin}
-\end{widetext}
+\section*{References}
+
+\bibliographystyle{apsrev4-1}
+\bibliography{references}
+
 \end{document}

### 2019-01-16 notes/Paper.tex
updated notes on paper

@@
 \hoffset -1 cm 
+\newcommand{\RomanNumeralCaps}[1]{\MakeUppercase{\romannumeral #1}}
 \newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
@@
 
-Gravitational wave (GW) memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it~\cite{Abbott_2016nmj}.
-The are two kinds of gravitational wave memory: linear and nonlinear.
-Linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments.
+Gravitational wave memory effect results in a permanent displacement of test masses after the gravitational wave has passed through an "ideal" GW detector.
+There are two kinds of gravitational wave memory effects: linear and nonlinear.
+The linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments.
 At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
@@
 \end{equation} 
-where $I_{ij}$ is the source mass-quadrupole moment, R is the distance source and $\Delta$ describes the net change in the quantity from early to late times.
-Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion [add ref], asymmetric mass loss due to neutrino emission and gamma-ray-burst jets [add ref] \\
+where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times.
+Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion [add ref], asymmetric mass loss due to neutrino emission and gamma-ray-burst jets, \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}
+\par The linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by the following general formula for by Thorn \cite{Thorne_Braginskii1987}
 
-A general formula for the linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by Thorn [add ref]
-
-\begin{equation}
+\begin{equation}\label{eq:1.2}
 \Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
@@
 \\
- Nonlinear memory effect also called Christodoulou memory, arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave.
+The other kind of memory is the nonlinear memory, which is also called Christodoulou memory.
+This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}.
+Christodoulou arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave.
 It can be understood in the following way.
 Consider the Einstein field equation in harmonic gauge.
-[add ref]
+\cite{Thorne1980}
 \begin{subequations}
-\begin{align}	
+\begin{align}\label{eq:3.1a}	
 \Box \bar{h}^{\alpha \beta} & = -16\pi (-g)(T^{\alpha\beta} + t_{LL}^{\alpha\beta})-\bar{h}^{\alpha\mu}_{,\nu}\bar{h}^{\beta\nu}_{,\mu}+\bar{h}^{\mu\nu}\bar{h}^{\alpha\beta}_{,\mu\nu}\\
@@
 is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator.
-The Landau-Lifshitz term is the Gravitational wave stress tensor[add ref].
+The Landau-Lifshitz term is the Gravitational wave stress tensor~\cite{Isaacson1968}.
 \begin{equation}
@@
 \end{equation}
-where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab,j}=\bar{h}^{TT}_{ab,j}$.
-When we apply standard Green's function to right hand side of equation (1.1a), we obtain the following correction term to GW field [add ref]
-\begin{equation}
+where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$.
+When we apply standard Green's function to right hand side of equation \ref{eq:3.1a}, we obtain the following correction term to GW field \cite{Wiseman1991}
+\begin{equation}\label{eq:1.6}
 	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'}\cdot\boldsymbol{N})}d\Omega'\Bigg]^{TT}
 \end{equation}
-where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy.\\
-The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity.
-It has a clear physical interpretation : It arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments.
+where $T_{R}$ is the retarded time.
+This equation shows that part of the distant GW is sourced by loss of GW energy.
+The nonlinear memory occurs in any system that radiates GWs.
+These systems are usually considered to have a vanishing linear memory.
+The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992}
+\par The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity, it arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments.
 The nonlinear memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order.
 This memory effect causes a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral.
-The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence but, most contribution to this nonlinear memory come from the merger phase where the PN model do not work.
-In this case the memory must be extracted from Numerical relativity simulation.
+The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work.
+In this case the memory must be extracted from Numerical relativity simulations.
 However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes.
@@
 These $m=0$ modes tend to be much smaller in magnitude and depends more sensitively on	initial conditions than other modes.
-For detailed discussion on this see ref [add ref] \\
-The NR simulations mostly easily compute  the curvature scalar $\Psi_4$ and decompose its values at large $R$ into a sum over spin-weighted spherical harmonic modes:
+For detailed discussion on this see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009}
+\par The NR simulations mostly easily compute  the curvature scalar $\Psi_4$ and decompose its values at large $R$ into a sum over spin-weighted spherical harmonic modes:
 \begin{equation}
@@
 \end{equation} 
-In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$.
-There are other difficulties in calculating memory from numerical relativity data.Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified.
-Choosing these integration constant to be zero [add ref] causes and artificial memory and a slope in the $h_{lm}$ modes.
-This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.\\
-Numerical relativity simulations starts at a finite radius, so although it contains a most part of gravitational wave memory, PN results should be used for early inspiral phase.
+In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$.
+There are many other difficulties too in calculating memory from numerical relativity data.
+Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified.
+Choosing these integration constant to be zero \cite{Reisswig_Pollney2011} causes an artificial memory and a slope in the $h_{lm}$ modes.
+This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.
+\par Numerical relativity simulations starts at a finite radius, so although it contains most part of gravitational wave memory, PN results should be used for early inspiral phase.
 The significance of memory accumulated during the inspiral phase can be seen by looking at the largest memory mode $h_{20}$.
@@
 where $U^{(mem)(1)}_{20}$ is given by the leading order piece of Eq.
-(3.13a) in ref [add ref], and the $T_0$ is the starting time of the simulation.
+(3.13a) in ref \cite{Favata2009}, and the $T_0$ is the starting time of the simulation.
 For a realistic, quasicircular binary that inspirals from a large initial separation, $T_0 \rightarrow -\infty$.
 The NR simulations of binaries that starts at a finite initial separation would underestimate the size of the memory.
-For a simulation that starts with an initial separation of $10M$, the error in the memory at a harmonic coordinate separation of $5M$ is $50\%$.
+For a simulation that starts with an initial separation of $10M$, the error in the memory at a harmonic coordinate separation of $5M$ is $50\%$.\cite{Favata2009}
 
-\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata [].
+\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata \cite{Favata2009}.
 However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})
-
-In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.In Sec.
+\par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.
+In Sec.
 II, we describe the calculation required to compute memory contribution to the post-Newtonian waveform of quasi-circular, inspiralling binaries.
 In Sec.
-III, we describe the method used extract memory contribution from NR simulations.
+III, we describe the method used to extract memory contribution from NR simulations.
 Sec.
@@
 The mode decomposition of gravitational wave polarization is given by
-\begin{equation}
+\begin{equation}\label{eq:2.1}
 	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
@@
 In a multipolar expansion of the GW field, the modes $h^{lm}$ are related to the radiative mass $U^{lm}$ and the current $V^{lm}$ multipole via
-\begin{equation}
+\begin{equation}\label{eq:2.2}
 	h^{lm}= \frac{G}{\sqrt{2}\, R }\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
@@
  The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
-\begin{equation}
+\begin{equation}\label{eq:2.3}
 	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
@@
 Here
-\begin{align}
+\begin{align}\label{eq:2.4}
 	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
@@
 where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$. \\
-When the GW field is decomposed into modes as in [ref eqn], the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders.
-\begin{equation}
+When the GW field is decomposed into spin-weighted spherical harmonic modes, the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders.
+For more detail about GW multipole expansion \cite{Thorne1980}.
+The correction to the radiative mass multipole moments is given by
+\begin{equation}\label{eq:2.6}
 	U^{(mem)}_{lm}=32\pi\frac{(l-2)!}{2(l+2)!}\int_{-\infty}^{T_R}dt\int d\Omega\frac{dE_{gw}}{dtd\Omega}(\Omega)Y^*_{lm}(\Omega)
 \end{equation}
-The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by [add ref]
-\begin{equation}
+The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by \cite{Favata2010}
+\begin{equation}\label{eq:2.7}
 	\frac{dE_{gw}}{dtd\Omega} = R^2 T^{gw}_{00} = \frac{R^2}{32\pi}\langle\dot{h}^{TT}_{jk}\dot{h}^{TT}_{jk}\rangle =\frac{R^2}{16\pi}\langle\dot{h}^2_{+}+ \dot{h}^2_{\times}\rangle
 \end{equation}
-where the angled bracket mean to average over several wavelengths. The energy flux and be written in terms of $h_{lm}$ modes using equation 2.1 in the following way:
+where the angled bracket mean to average over several wavelengths. The energy flux can be written in terms of $h_{lm}$ modes using equation \ref{eq:2.1} in the following way:
 \begin{widetext}
-\begin{equation}
+\begin{equation}\label{eq:2.8}
 		\frac{dE_{gw}}{dtd\Omega}=\frac{R^2}{16\pi}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}\langle\dot{h}_{l'm'}\dot{h}^*_{l''m''}\rangle _{-2}Y^{l'm'}(\theta,\phi)_{-2}Y^{l''m''*}(\theta, \phi)
@@
 
-The memory contribution to mass multipole moment can now be evaluated using equation 2.6 and 2.8. The angular part of the integral has the following form
+The memory contribution to the mass multipole moment can now be evaluated using equation \ref{eq:2.6} and \ref{eq:2.8}. The angular part of the integral has the following form
 
-\begin{equation}
+\begin{equation}\label{eq:2.9}
 	\int d\Omega _{-2}Y_{l'm'}(\theta,\phi)_{-2}Y_{l''m''}^{*}(\theta, \phi)Y_{lm}^{*}(\theta, \phi)
@@
 
-The above integral can be evaluated as done in ref [add reference]. The result is show in appendix. \\
-The time derivative of the memory mass-multipole moment is defines as [add ref] 
- $U^{(mem)(1)}_{lm}\equiv\frac{dU^{(mem)}_{lm}}{dT_R}$, which when combined using equation 2.6 and 2.8, can be written as
+The above integral can be evaluated as done in section III of  Favata\cite{Favata2009}.
+The results are also shown in appendix ~\ref{Appendix A}.
+The time derivative of the memory mass-multipole moment is defines as $U^{(mem)(1)}_{lm}\equiv\frac{dU^{(mem)}_{lm}}{dT_R}$ \cite{Favata2009}, which when combined using equation \ref{eq:2.6} and \ref{eq:2.8} and Appendix \ref{Appendix A}, can be written as
 
-\begin{align}
-	U_{lm}^{(mem)(1)}=&R^{2}\sqrt{\frac{2(l-2)!}{(l+2)!}}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}(-1)^{m+m'}\\
+\begin{align}\label{eq:2.10}
+	U_{lm}^{(mem)(1)}=&R^{2}\sqrt{\frac{2(l-2)!}{(l+2)!}}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}(-1)^{m+m'}\nonumber\\
 	&\times\Bigg\langle \dot{h}_{l'm'}\dot{h}_{l'm'}^{*}\Bigg\rangle G^{2-20}_{l'm'lm'-m''-m}
@@
 
-where $G^{2-20}_{l'm'lm'-m''-m}$ is the angular integral of equation 2.9 and is given in Appendix [add].
+where $G^{2-20}_{l'm'lm'-m''-m}$ is the angular integral of equation \ref{eq:2.9} and is given in Appendix \ref{Appendix A}.
 The above equation can be used to compute the first time derivative of mass-multipole moment by substituting the GW modes $h_{lm}$.
-When evaluating equation 2.10, we are only interested in $m=0$ modes.
-The $m\neq=0$ terms yield oscillatory contribution to the waveform polarizations that enter at higher PN orders then the nonoscillatory, $m=0$ terms.
+When evaluating equation \ref{eq:2.10}, we are only interested in $m=0$ modes.
+The $m\neq0$ terms yield oscillatory contribution to the waveform polarizations that enter at higher PN orders then the nonoscillatory, $m=0$ terms.
 The non-vanishing, non-oscillatory modes are $U^{(mem)}_{l0}$ with l-even.
 These $U^{(mem)}_{l0}$ consist of time-integral of polynomials in $x\equiv(M\omega)^{2/3}$ (where $\omega$ is the orbital angular frequency), with each terms of the form
-\begin{equation}
+\begin{equation}\label{eq:2.11}
 \int_{-\infty}^{T_{R}} x^{n}dt=\int_{-\infty}^{T_{R}}\frac{x^{n}}{\dot{x}}dx
 \end{equation}
-After performing the change of variable indicated in above equation 2.12, evaluation of the resulting integrals requires a model for the frequency evolution of the binary.
+After performing the change of variable indicated in above equation \ref{eq:2.11}, evaluation of the resulting integrals requires a model for the frequency evolution of the binary.
 The adiabatic evolution of the frequency (or $x$) is easily derived from energy balance ($\mathcal{L}_{GW}=-\dot{E}$).
 and the relation $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$.
-The 3.5PN orbital energy is given by equation C in ref[add aruns paper ref]
+The 3.5PN orbital energy and flux are given in Appendix C of reference \cite{Arun2009}.
+Here we who only upto 1.5 PN terms, the complete expression are show in Appendix \ref{Appendix:C1}.
+The 1.5 PN expression for energy is given by
 \begin{widetext}
@@
 
-and the 3PN GW luminosity is given by Eq C7 to C12 in [ref]
+and the 1.5 PN GW luminosity is given by
 
@@
 
-Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, we get the following result
+Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, and expanding upto 1.5 PN gives the following expression.
 
-\begin{align}\label{eq:FbydE}
+\begin{align}\label{eq:2.14}
 \frac{dx}{dt} & = \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
@@
 \end{widetext}
-The above equation can be used to evaluate the integral in Eq. 2.12
+The results upto 3.5 PN are shows in Appendix \ref{Appendix:C}.
+The equation \ref{eq:2.14} can be used change the integration with respect to time into integration with respect to $x$ as shown in equation in \ref{eq:2.11}.
+The expressions for $h_{lm}$ whose first time derivatives appear equation \ref{eq:2.10} are directly taken from reference \cite{Arun2009}.
+The PN expression for $h_{lm}$ there are computed for precessing binaries(Equations 4.17 in ref \cite{Arun2009}), but here we are only interested in aligned spin case.
+The structure of modes for non-precessing binaries can be obtained by aligning the orbital angular momentum with the total angular momentum($\iota = 0, \alpha = \pi$).
+These results for a special case of non-precessing binaries is shown in Appendix \ref{Appendix:D}.
+We show the $h_{lm}$ expressions for positive $m$ values, but the results for negative $m$ values can be obtained by the relation, $h_{l, -m} = (-1)^l \bar{h}_{lm}$.
+This results is not in general true for precessing binaries, but it continues to hold true of non-pressing case.
 
+\begin{widetext}
 \subsection{RESULTS: MEMORY CONTRIBUTION TO THE POST-NEWTONIAN WAVEFORM OF QUASI-CIRCULAR, INSPIRALLING BINARIES}
-The memory contribution to the spin-weighted spherical-harmonic modes of the polarization waveform [ref eqn]. These quantities are related via 
+The memory contribution to the spin-weighted spherical-harmonic modes of the polarization waveform [Eq. \ref{eq:2.2}]. These quantities are related via 
 \begin{equation}
@@
 \end{equation}
-where we have followed the notation of Sec. 9 of Ref [add referece]. The notational parameter $\alpha$ accounts for the two commonly used choices for the polarization triad 
+where we have followed the notation of Sec. 9 of Ref. \cite{Blanchet_Iyer2008}. The notational parameter $\alpha$ accounts for the two commonly used choices for the polarization triad 
 \begin{equation}
-	content...
+\alpha = \left\{
+\begin{array}{ll}
+(+1) & \text{for the Kidder  \cite{Kidder2008} convention} \\
+(-1) & \text{for the Blanchet et al. \cite{Blanchet_Iyer2008}   convention}\\
+\end{array} 
+\right.
 \end{equation}
-(\ashok{Need to add}).\\
+
 The results of polarization modes in terms of $\hat{H}_{l0}$ are:
-\begin{widetext}
+\begin{subequations}
 \begin{align}\label{eq:H20}
@@
 \end{align}
-\end{widetext}
+\end{subequations}
 We can see that all contribution to $m=0$, even-$l$ modes arise solely from the Christodoulou memory piece of $U_{lm}$.
@@
 We can combine the above with Eq 2.16 and explicitly compute the memory contribution to the + waveform polarization.This can be done in the following way
-\begin{widetext}
+
 \begin{equation}
@@
 Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$.
-It has long been noted that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift.
+It has long been noted that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}.
 The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation.
@@
 As a result in numerical simulation, the finite distance calculation introduces a systematic error.
-In more recent simulations [added ref], which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
-\par It has been argued in [add ref] that an important source of unphysical non-linear drift in numerical computation of gravitational wave strain lies in the transformation of the measured data to the strain $h$  which generally involves an integration in time.
+In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
+\par It has been argued in ref.
+\cite{Reisswig_Pollney2011} that an important source of unphysical non-linear drift in numerical computation of gravitational wave strain lies in the transformation of the measured data to the strain $h$  which generally involves an integration in time.
 The output of the numerical simulation is a discretely sampled time series of finite duration, incorporating some component of unresolved frequencies due to numerical error.
 This can lead to an uncontrollable non-linear drift if the integration is performed in the time domain.
-The unfiltered GW waveform is shown in fig [add fig], which clearly shows an oscillating waveform which settles to a nonzero value.
-The method described in ref[add ref] gives us a way to get rid of this unphysical behaviour
-\par The memory contribution to $h_{+}$ is given by
+The method described in ref \cite{Reisswig_Pollney2011} gives us a way to get rid of this unphysical behavior by eliminating the lower frequencies.
+\par The memory contribution to $h_{+}$ is given by \cite{Favata2010}
 
-\begin{equation}\label{3.2}
+\begin{equation}\label{eq:3.2}
 h_{+}^{(mem)} \approx \frac{R}{192 \, \pi}s_{\Theta}^{2}\left(17 +c_{\Theta}^{2} \right)\int_{-\infty}^{T_R}|\dot{h}_{22}|^2 dt
@@
 The above equation requires a time integration of absolute value of $\dot{h}_{22}$, but this is not typically the quantity which is directly computed in numerical simulation.
-The output of numerical simulation is usually in the form of component of curvature tensor components, or Zerilli-Moncrief-type variables defined relative to a background.
+The output of numerical simulation is usually in the form of component of curvature tensor, or Zerilli-Moncrief-type variables defined relative to a background.
 \subsection{Evaluating Bondi News from numerical data}
-The Bondi News is determined by
+The Bondi News which also the first derivative of gravitational wave strain is defined as
 \begin{equation}\label{3.3}
@@
 \end{equation}
-The gravitational wave memory contribution to $h_{+}$ polarization can be computed from this bondi news, by integrating the absolute value of bondi news as shown in equation \ref{hmem}.
-The memory contribution evaluated using Eq.\ref{hmem} will contain contribution from numerical noise, which is unphysical and must be removed from the data.
+The gravitational wave memory contribution to $h_{+}$ polarization can be computed from this bondi news, by integrating the absolute value of bondi news as shown in equation \ref{eq:3.2}.
+The memory contribution evaluated using Eq.\ref{eq:3.2} will contain contribution from numerical noise, which is unphysical and must be removed from the data.
+We apply the method described in the following section to remove the unphysical part of GW memory.
 \subsection{Integration of finite length signals in frequency domain}
-The unphysical part of mermory essential lies in low frequency. If we remove this low frequency part from data, we can get rid of memory contribution due to numerical noise. 
+The unphysical part of mermory essentially lies in lower frequency.
+If we remove this low frequency part from data such that the physical memory is still preserved, we can get rid of memory contribution due to numerical noise.
 Consider the Fourier transform, $\mathcal{F}$, applied to an absolutely integrable function $f(t)$,
@@
 The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
-
+\subsection{Compleate Christodoulou memory during BBH coalescence} 
+The PN waveform are given as a function of parameter $x$ which is related to time by equation \ref{eq:FbydE}.
+This equation can be integrated to obtain a relation between time $t$ and parameter $x$.
+The Christodoulou memory reaches it's peak value at the merger.
+We choose the values of integration constants such that this peak in $h_+^{(mem)}$ is reached at time $t=0$.
+For comparison we show the time derivative of $h_+^{(mem)}$ from Numerically obtained results and the analytical expressions in figure \ref{fig:derivative}.
+The slope as seen from figure \ref{fig:derivative} rises asymptotically.
+We use our PN results for computing the memory during early inspiral and attach them to NR results.
+We compare this memory for different mass ratio and different values of dimensionless spin parameter.
+Figure \ref{fig:differentmassratio} show the Christodoulou memory for different mass ratios, which shows that the Christodoulou tends to decrease with higher mass ration.
+The effect of dimensionless spin is shown in figure \ref{fig:differntchivalues}.
+It can be seen that the spin magnitude has significant effect on nonlinear memory, higher spin magnitude tends to increase the memory contribution.
+\begin{figure}
+	\includegraphics[width=4.0in]{/home/ashok/gravitational_wave_memory_project/plots/DerivativeComparisionNRvsPN0p800.pdf}
+	\caption{The figure above shows a comparison for time derivatives Christodoulou memory contribution to $h_+$ polarization between NR data and PN results }
+	\label{fig:derivative}
+\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \includegraphics[width=4.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
-\caption{NonSpinning case}
-\label{fig:q7}
+\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
+This figure shows a comparison for different mass ratios.
+It can be seen that higher mass ration tends to decrease the Chistodoulou memory.
+The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using Numerical Relativity data}
+\label{fig:differentmassratio}
 \end{figure}
@@
 \includegraphics[width=4.0in]{../plots/MemoryPlot_alignedSpin/0p95.pdf}
-\caption{NonSpinning case}
-\label{fig:0p95}
+\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
+This figure shows a comparison for different values of dimensionless spin parameter $\chi$.
+These results are for BBH system with equal spin magnitude for individual black hole and spins aligned with orbital angular momentum.
+It can be seen that higher spin magnitude tend to increase the Christodoulou memory.
+}
+\label{fig:differntchivalues}
 \end{figure}
@@
 
-\section{Angular Integral Of The Triple Product Of Spin-Weighted Spherical Harmonics}
+\section{Angular Integral Of The Triple Product Of Spin-Weighted Spherical Harmonics}\label{Appendix A}
 The following integral appears in the evaluation of nonlinear memory contribution to the radiative mass-multipole moments $U_{lm}$. Equation 2.4 can be rewritten as
@@
 
-\section{Complete 3.5 PN expressions}
+\section{Complete 3.5 PN expressions for Energy and Flux}\label{Appendix:C}
 
-\begin{align}\label{eq:Ef}
+\begin{align}\label{Appendix:C1}
 E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right.
@@
 \\
-\label{eq:Flux}
+\label{Appendix:C2}
 \mathcal{L}&=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right]\right.\nonumber\\
@@
 \\
-\label{eq:dxdt}
+\label{Appendix:C3}
 \frac{dx}{dt}&=\frac{64\, x^5  \, \eta}{5}\left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
@@
 
+\section{$h_{lm}$ modes for non-precessing binaries}\label{Appendix:D}
+The GW modes for precessing binaries with small inclination angle on nearly circular orbits through 1.5 PN order are given in section IV of ref \cite{Arun2009}.
+Setting $\iota = 0, \alpha = \pi$ in equation 4.16 and 4.17 gives the result for non-precessing case.
+In that case the spins are aligned along $z$ axis, so we set $\chi_a^x=\chi_s^x=\chi_a^y=\chi_s^y=0$.
+We can now write the expression as
+\begin{equation}
+	h_{lm} = (-1)^{m+1}\frac{(2M\eta v^2)}{D_L}\sqrt{\frac{16\pi}{5}}e^{\mathit{i}\,m\,\psi}\hat{h}_{lm},
+\end{equation}
+and $\hat{h}_{lm}$ modes are given by
+
+\begin{subequations}
+\begin{align}
+	\hat{h}_{22} &= 1 + x\left(-\frac{107}{42} + \frac{55\eta}{42}\right) + x^{3/2}\left(2\pi - \frac{4\,\delta\,\chi^{z}_{a}}{3} + \frac{4}{3}\left(-1+\eta\right)\chi^z_s\right)\\
+	\hat{h}_{21} &= \frac{\mathit{i}\delta}{3}\left[x^{1/2} - x\frac{3}{2\delta}\left(\chi^{z}_{a} + \delta\chi^{z}_{s}\right) + x^{3/2}\left(-\frac{17}{28} + \frac{5\eta}{7}\right)\right]\\
+	\hat{h}_{32} &= \frac{9}{8}\sqrt{\frac{5}{7}}\left[\frac{8}{27}x\left(1-3\nu\right) + \frac{32}{27}x^{3/2}\eta\,\chi^z_s\right]
+\end{align}
+\end{subequations}
+$\hat{h}_{lm}$ shown above are the only modes which has spin contributions.
+We only include spin contribution upto 1.5PN but keep the remaining expression upto 3PN.
+The complete expression upto 3.5PN are given in ref \cite{Blanchet_Iyer2008}
 \section*{References}

### 2019-01-18 notes/Paper.tex
undates notes

@@
 \end{subequations}
+\begin{figure}
+	\includegraphics[width=3.0in]{../plots/PNmemoryContributiontoHplus.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemoryContributionfunofTheta.pdf}
+	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
+The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4 $ .
+The plot to the right shows $x H_+$ as a function of parameter x.
+Increasing the magnitude of 𝜒 does not change the angular dependence, but does tend to increase the magnitude of the memory.
+		}
+	\label{fig:0p95}
+\end{figure}
 \end{widetext}
@@
 The unphysical part of mermory essentially lies in lower frequency.
+A method to remove this unphysical part is given in ref \cite{Reisswig_Pollney2011}.
 If we remove this low frequency part from data such that the physical memory is still preserved, we can get rid of memory contribution due to numerical noise.
@@
 We used this data and  performed a frequency domain time integration as described in equation \ref{3.6}, while removing the unphysical frequency by choosing a lower cutoff as show in section \ref{FFI}.
-The cutoff frequency in chosen as $\omega_0/10$ where $\omega_0$ is the instantaneous GW angular frequency.
-The first begin separating the complex Newman-Perose scalar $\psi_4$ data into phase and amplitude part.
+The cutoff frequency is chosen as $\omega_0/10$ where $\omega_0$ is the instantaneous GW angular frequency.
+We first begin by separating the complex Newman-Perose scalar $\psi_4$ data into phase and amplitude part.
 We compute instantaneous GW angular frequency from this by taking the first time derivative of phase corresponding to $\psi_4$.
 The instantaneous GW frequency often looks noisy at the beginning of the simulations due to numerical issues.
-We remove this noise data choose a lower and upper cutoff time.
+We remove this noise data by choosing a lower and upper cutoff time.
 Once we find the upper and lower cutoff time, we can apply a frequency domain time integration for $\psi_4$ by choosing a cutoff frequency.
-After filtering the noise creating frequency we can use equation \ref{3.2} to compute the memory contribution to "plus" polarization of GW signal.
+After filtering the noise creating frequency we can use equation \ref{eq:3.2} to compute the memory contribution to "plus" polarization of GW signal.
 The SXS catalog gives data for different parameter values like mass ratio, dimensionless spin parameter etc.
-The Christodoulou memory computed from Numerical Relativity data starts from zeros using equation \ref{3.2} due to finite simulation time.
+The Christodoulou memory computed from Numerical Relativity data starts from zeros using equation \ref{eq:3.2} due to finite simulation time.
 Chrostodoulou memory is hereditary, hence it's entire past should be included.
-This can be done by combining the post-Newtonian expression during early inspiral and numerical relativity results for merger part.
+This can be done by combining the PN expression during early inspiral and numerical relativity results for merger part.
 The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
 \subsection{Compleate Christodoulou memory during BBH coalescence} 
-The PN waveform are given as a function of parameter $x$ which is related to time by equation \ref{eq:FbydE}.
+The PN waveform are given as a function of parameter $x$ which is related to time by equation \ref{eq:2.14}.
 This equation can be integrated to obtain a relation between time $t$ and parameter $x$.
 The Christodoulou memory reaches it's peak value at the merger.
-We choose the values of integration constants such that this peak in $h_+^{(mem)}$ is reached at time $t=0$.
+We choose the values of integration constants such that the peak in $h_+^{(mem)}$ is reached at time $t=0$.
 For comparison we show the time derivative of $h_+^{(mem)}$ from Numerically obtained results and the analytical expressions in figure \ref{fig:derivative}.
@@
 The effect of dimensionless spin is shown in figure \ref{fig:differntchivalues}.
-It can be seen that the spin magnitude has significant effect on nonlinear memory, higher spin magnitude tends to increase the memory contribution.
+It can be seen that the spin magnitude has significant effect on nonlinear memory, higher spin magnitude, in case of spin aligned with orbital angular momentum tends to increase the memory contribution.
+When the spins are anti aligned with orbital angular momentum, the memory tend to decrease the Christodoulou memory.
+These results show it important to consider the spin contribution when studying the Christodoulou memory in gravitational wave signals.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
@@
 It can be seen that higher mass ration tends to decrease the Chistodoulou memory.
-The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using Numerical Relativity data}
+The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using Numerical Relativity data.
+The red dashed line shows result from PN expression}
  	\label{fig:differentmassratio}
@@
 \begin{figure}
-\includegraphics[width=4.0in]{../plots/MemoryPlot_alignedSpin/0p95.pdf}
+	\includegraphics[width=4.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
+	\includegraphics[width=4.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
-This figure shows a comparison for different values of dimensionless spin parameter $\chi$.
+The top figure shows a comparison for different values of dimensionless spin parameter $\chi$.
 These results are for BBH system with equal spin magnitude for individual black hole and spins aligned with orbital angular momentum.
 It can be seen that higher spin magnitude tend to increase the Christodoulou memory.
+The bottom figure shows the anti aligned case.
 }
@@
 \end{figure}
-\begin{figure}
-	\includegraphics[width=3.0in]{../plots/PNmemoryContributiontoHplus.pdf}
-	\includegraphics[width=3.0in]{../plots/PNmemoryContributionfunofTheta.pdf}
-	\caption{NonSpinning case}
-	\label{fig:0p95}
-\end{figure}
+
 \newpage

### 2019-01-18 notes/Paper.tex
updated notes

@@
 where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times.
-Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion [add ref], asymmetric mass loss due to neutrino emission and gamma-ray-burst jets, \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}
+Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion, asymmetric mass loss due to neutrino emission and gamma-ray-burst jets, \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}
 \par The linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by the following general formula for by Thorn \cite{Thorne_Braginskii1987}

### 2019-01-27 notes/Paper.tex
updated notes

@@
 \begin{align}\label{eq:F}
-\mathcal{L} &=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right.\right]\nonumber\\
+\mathcal{L} &=\frac{32}{5}\eta^2x^5\left\{1 + x\left[-\frac{1247}{336}-\frac{35\,\eta}{12}\right] + x^{3/2}\left[4\pi-\left\{\left(\frac{11}{4}-3\eta\right)\chisdL +\frac{11}{4}\delta\chiadL\right\}\right]\right\}\nonumber\\
 \end{align}
@@
 \frac{dx}{dt} & = \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
-+ x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
++ x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right\} \nonumber \\ 
 \end{align}
@@
 The results upto 3.5 PN are shows in Appendix \ref{Appendix:C}.
-The equation \ref{eq:2.14} can be used change the integration with respect to time into integration with respect to $x$ as shown in equation in \ref{eq:2.11}.
+The equation \ref{eq:2.14} can be used to change the integration with respect to time into integration with respect to $x$ as shown in equation in \ref{eq:2.11}.
 The expressions for $h_{lm}$ whose first time derivatives appear equation \ref{eq:2.10} are directly taken from reference \cite{Arun2009}.
@@
 The structure of modes for non-precessing binaries can be obtained by aligning the orbital angular momentum with the total angular momentum($\iota = 0, \alpha = \pi$).
-These results for a special case of non-precessing binaries is shown in Appendix \ref{Appendix:D}.
+These results for a special case of non-precessing binaries, which contain spin contribution at 1.5PN order are shown below
+\begin{subequations}
+	\begin{align}\label{eq:2.15}
+	\hat{h}_{22} &= 1 + x\left(-\frac{107}{42} + \frac{55\eta}{42}\right) + x^{3/2}\left(2\pi - \frac{4\,\delta\,\chi^{z}_{a}}{3} + \frac{4}{3}\left(-1+\eta\right)\chi^z_s\right)\\
+	\hat{h}_{21} &= \frac{\mathit{i}\delta}{3}\left[x^{1/2} - x\frac{3}{2\delta}\left(\chi^{z}_{a} + \delta\chi^{z}_{s}\right) + x^{3/2}\left(-\frac{17}{28} + \frac{5\eta}{7}\right)\right]\\
+	\hat{h}_{32} &= \frac{9}{8}\sqrt{\frac{5}{7}}\left[\frac{8}{27}x\left(1-3\nu\right) + \frac{32}{27}x^{3/2}\eta\,\chi^z_s\right]
+	\end{align}
+\end{subequations}
+The remaining $h_{lm}$ expressions does not contain spin terms and the complete expressions upto 3PN terms can be directly taken from Ref.\cite{Blanchet_Iyer2008}
  We show the $h_{lm}$ expressions for positive $m$ values, but the results for negative $m$ values can be obtained by the relation, $h_{l, -m} = (-1)^l \bar{h}_{lm}$.
 This results is not in general true for precessing binaries, but it continues to hold true of non-pressing case.
+With these modes we use equation \ref{eq:2.10} to evaluate the first time derivative of the memory mass-multipole moment, which is further integrated using equation \ref{eq:2.11}.
+The results after integration are show in next section.
 
@@
 \begin{equation}
-	h_{l0}^{(mem)} = \frac{\alpha}{\sqrt{2} \, R} U_{l0}^{(mem)} = 8 \sqrt{\frac{\pi}{5}}\frac{\eta M x}{R} 
+	h_{l0}^{(mem)} = \frac{\alpha}{\sqrt{2} \, R} U_{l0}^{(mem)} = 8 \sqrt{\frac{\pi}{5}}\frac{\eta M x}{R}\hat{H}_{lm}
 \end{equation}
@@
 \begin{align}\label{eq:H20}
-\hat{H}_{20} &= \alpha \frac{5}{14\sqrt{6}}\left\{ 1 + x\left[-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right]+ x^{3/2}\left[\left(\frac{7}{480} +\frac{27 \, \delta}{10} - \frac{7 \,\eta}{120}\right)\chiadL+\left(\frac{27}{10} +\frac{7 \, \delta^3}{480} - \frac{2 \, \eta}{15}\right)\chisdL\right]\right.
+&\hat{H}_{20} = \alpha \frac{5}{14\sqrt{6}}\left\{ 1 + x\left[-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right]+ x^{3/2}\left[\left(\frac{7}{480} +\frac{27 \, \delta}{10} - \frac{7 \,\eta}{120}\right)\chiadL+\left(\frac{27}{10} +\frac{7 \, \delta^3}{480} - \frac{2 \, \eta}{15}\right)\chisdL\right]\right.
 \nonumber \\
@@
 \begin{align}\label{eq:H40}
-	\hat{H}_{40} &=\alpha\frac{1}{504\sqrt{2}}\left\{1 + x\left[-\frac{180101}{29568} + \frac{27227\,\eta}{1056}\right] + x^{3/2}\left[\left(\frac{7}{40} + \frac{27\,\delta}{10} -\frac{7\,\eta}{10}\right)\chiadL+\left(\frac{27}{10}+\frac{7\,\delta^3}{40} +\frac{46\,\eta}{5}\right)\chisdL\right]\right.\nonumber\\
+	&\hat{H}_{40} =\alpha\frac{1}{504\sqrt{2}}\left\{1 + x\left[-\frac{180101}{29568} + \frac{27227\,\eta}{1056}\right] + x^{3/2}\left[\left(\frac{7}{40} + \frac{27\,\delta}{10} -\frac{7\,\eta}{10}\right)\chiadL+\left(\frac{27}{10}+\frac{7\,\delta^3}{40} +\frac{46\,\eta}{5}\right)\chisdL\right]\right.\nonumber\\
 	&\qquad x^{5/2}\left[-\frac{13565\,\pi}{1232} + \frac{13565\,\pi\,\eta}{308} + \left\{\frac{379}{2688}-\frac{1457321\,\delta}{68992}+\left(-\frac{59}{336} + \frac{3094367\,\delta}{22176}\right)\eta -\frac{87\,\eta^2}{56}\right\}\chiadL\right.\nonumber\\
@@
 \begin{align}\label{eq:H60}
-	\hat{H}_{60} &=-\alpha\frac{4195}{1419264\sqrt{273}}x\left\{1-\frac{3612\,\eta}{839}+ x\left[-\frac{45661561}{6342840} + \frac{101414\,\eta}{2517} - \frac{48118\,\eta^2}{839}\right]\right.
+	&\hat{H}_{60} =-\alpha\frac{4195}{1419264\sqrt{273}}x\left\{1-\frac{3612\,\eta}{839}+ x\left[-\frac{45661561}{6342840} + \frac{101414\,\eta}{2517} - \frac{48118\,\eta^2}{839}\right]\right.
 + x^{3/2} \left[\frac{1248\,\pi}{839} -\frac{4992\,\pi\,\eta}{839}\right.\nonumber\\
@@
 \begin{align}\label{eq:H80}
-	\hat{H}_{80} &=\alpha\frac{75601}{213497856\sqrt{119}}x^2\left[1-\frac{452070\,\eta}{75601} + \frac{733320\,\eta^2}{75601} + x\left(-\frac{265354303}{33869248} + \frac{18177343651\,\eta}{321757856} - \frac{5057483259\,\eta^2}{40219732} + \frac{1828935981\,\eta^3}{20109866}\right)\right]\nonumber\\
+	&\hat{H}_{80} =\alpha\frac{75601}{213497856\sqrt{119}}x^2\left[1-\frac{452070\,\eta}{75601} + \frac{733320\,\eta^2}{75601} + x\left(-\frac{265354303}{33869248} + \frac{18177343651\,\eta}{321757856} - \frac{5057483259\,\eta^2}{40219732} + \frac{1828935981\,\eta^3}{20109866}\right)\right]\nonumber\\
 \end{align}
@@
 There are no hereditary-memory contribution to the radiative current multipole moments $V_{lm}$\\
-We can combine the above with Eq 2.16 and explicitly compute the memory contribution to the + waveform polarization.This can be done in the following way
+We can combine the above with equation \ref{eq:2.1} and explicitly compute the memory contribution to the + waveform polarization.This can be done in the following way
 
@@
 \begin{align}
-	H_{+}^{(0,mem)}&=\alpha\frac{1}{96}s_{\Theta}^2\left(17 + c_{\Theta}^2\right),\\
-	H_{+}^{(0.5,mem)}&=0,\\
-	H_{+}^{(1,mem)}&=\alpha s_{\Theta}^2\left[-\frac{354241}{2064384} - \frac{62059}{1032192}c_{\Theta}^2 - \frac{4195}{688128}c_{\Theta}^4 + \left(\frac{15607}{73728} + \frac{9373}{36864}c_{\Theta}^2 +\frac{215}{8192}c_{\Theta}^4\right)\right],\\
-	H_{+}^{(1.5,mem)}&=\alpha s_{\Theta}^2\left\{\left[\frac{3}{1280}  + \frac{7}{3840}c_{\Theta}^2 +  \delta\left(\frac{153}{320} + \frac{9}{320}c_{\Theta}^2\right) +\eta \left(-\frac{3}{320}-\frac{7}{960}c_{\Theta}^2\right)\right]\chiadL\right.\nonumber\\
-	& \quad+\left.\left[\left(\frac{153}{320} + \frac{9}{320}c_{\Theta}^2\right) + \delta\left(\frac{3}{1280} + \frac{7}{3840}c_{\Theta}^2\right) + \eta\left(-\frac{3}{80} +\frac{23}{240}c_{\Theta}^2\right) + \delta\eta\left(-\frac{3}{320}-\frac{7}{960}\right)\right]\chisdL\right\}\\
-	H_{+}^{(2,mem)} &=\alpha s_{\Theta}^2\left\{-\frac{3968456539}{9364045824} + \frac{570408173}{4682022912}c_{\Theta}^2 + \frac{122166887}{3121348608}c_{\Theta}^4 + \frac{7560}{15925248}c_{\Theta}^6 \right.\nonumber\\
-	& \quad + \eta\left[-\frac{7169749}{18579456} - \frac{13220477}{18579456}c_{\Theta}^2 -\frac{1345405}{6193152}c_{\Theta}^4 - \frac{25115}{884736}c_{\Theta}^6\right] + \eta^2\left[\frac{10097}{147456}\right]
+	& H_{+}^{(0,mem)}=\alpha\frac{1}{96}s_{\Theta}^2\left(17 + c_{\Theta}^2\right),\\
+	& H_{+}^{(0.5,mem)}=0,\\
+	& H_{+}^{(1,mem)}=\alpha s_{\Theta}^2\left[-\frac{354241}{2064384} - \frac{62059}{1032192}c_{\Theta}^2 - \frac{4195}{688128}c_{\Theta}^4 + \eta\left(\frac{15607}{73728} + \frac{9373}{36864}c_{\Theta}^2 +\frac{215}{8192}c_{\Theta}^4\right)\right],\\
+	& H_{+}^{(1.5,mem)}=\alpha s_{\Theta}^2\left\{\left[\frac{3}{1280}  + \frac{7}{3840}c_{\Theta}^2 +  \delta\left(\frac{153}{320} + \frac{9}{320}c_{\Theta}^2\right) +\eta \left(-\frac{3}{320}-\frac{7}{960}c_{\Theta}^2\right)\right]\chiadL\right.\nonumber\\
+	& \qquad\qquad+\left.\left[\left(\frac{153}{320} + \frac{9}{320}c_{\Theta}^2\right) + \delta\left(\frac{3}{1280} + \frac{7}{3840}c_{\Theta}^2\right) + \eta\left(-\frac{3}{80} +\frac{23}{240}c_{\Theta}^2\right) + \delta\eta\left(-\frac{3}{320}-\frac{7}{960}\right)\right]\chisdL\right\}\nonumber\\
+	\\
+	& H_{+}^{(2,mem)} =\alpha s_{\Theta}^2\left\{-\frac{3968456539}{9364045824} + \frac{570408173}{4682022912}c_{\Theta}^2 + \frac{122166887}{3121348608}c_{\Theta}^4 + \frac{7560}{15925248}c_{\Theta}^6 \right.\nonumber\\
+	& \qquad\qquad + \eta\left[-\frac{7169749}{18579456} - \frac{13220477}{18579456}c_{\Theta}^2 -\frac{1345405}{6193152}c_{\Theta}^4 - \frac{25115}{884736}c_{\Theta}^6\right] + \eta^2\left[\frac{10097}{147456} + \frac{5179}{36864}c_{\Theta}^2 + \frac{44765}{147456}c_{\Theta}^4\right.\nonumber\\
+	& \qquad\qquad\quad \left.+ \frac{3395}{73728}c_{\Theta}^6\right] + \left[-\frac{1385}{4608}-\frac{85}{4608}c_{\Theta}^2 +\eta\left(\frac{19}{16} + \frac{7}{96}c_{\Theta}^2\right)\right]\chiaSqr + \left[-\frac{1385\sqrt{1-4\eta}}{2304} - \frac{85\sqrt{1-4\eta}}{2304}c_{\Theta}^2\right.\nonumber\\
+	& \qquad\qquad\quad \left.
+\left.+ \eta\left(\frac{\sqrt{1-4\eta}}{72} + \frac{\sqrt{1-4\eta}}{144}c_{\Theta}^2\right)\right]\chisDchia + \left[-\frac{1385}{4608}-\frac{85}{4608}c_{\Theta}^2 + \eta^2\left(-\frac{1}{36}-\frac{c_{\Theta}^2}{72}\right)+ \eta\left(\frac{11}{384} + \frac{c_{\Theta}^2}{128}\right)\right]\chisSqr\right\}\nonumber\\
+	\\
+	& H_{+}^{(2.5,mem)} =\alpha s_{\Theta}^2\left\{-\frac{2545\pi}{21504} - \frac{295\pi}{2688}c_{\Theta}^2 -\frac{65\pi}{7168}c_{\Theta}^4 + \eta\left[\frac{2545\pi}{5376} + \frac{295\pi}{672}c_{\Theta}^2 + \frac{65\pi}{1792}c_{\Theta}^4\right] + \left[\frac{11867}{3612672} + \frac{8156279\sqrt{1-4\eta}}{4816896}\right.\right.\nonumber\\
+	& \qquad\qquad\quad + \frac{163}{112896}c_{\Theta}^2 - \frac{484979\sqrt{1-4\eta}}{2408448}c_{\Theta}^2 + \frac{55}{1204224}c_{\Theta}^4 -\frac{70645\sqrt{1-4\eta}}{2064384}c_{\Theta}^4 + \eta^2\left(-\frac{3797}{150528}-\frac{411}{25088}c_{\Theta}^2\right.\nonumber\\
+	& \qquad\qquad\quad \left.+ \frac{55}{150528}c_{\Theta}^4\right) + \eta^2\left(-\frac{3797}{150528}-\frac{411}{25088}c_{\Theta}^2+\frac{55}{150528}c_{\Theta}^4\right) + \eta\left(-\frac{12343}{1806336}-\frac{4105391\sqrt{1-4\eta}}{10838016}-\frac{1517}{903168}c_{\Theta}^2 \right.\nonumber\\ &\qquad\qquad\quad\left.\left.+\frac{7446571\sqrt{1-4\eta}}{5419008}-\frac{55}{200704}c_{\Theta}^4 + \frac{58395\sqrt{1-4\eta}}{401408}c_{\Theta}^4\right)\right]\chiadL + \left[\frac{8156279}{4816896}+ \frac{11867\sqrt{1-4\eta}}{3612672}+ \frac{484979}{2408448}c_{\Theta}^2 \right.\nonumber\\
+	&\qquad\qquad\quad + \frac{163\sqrt{1-4\eta}}{112896}c_{\Theta}^2 - \frac{70645}{2064384}c_{\Theta}^4 + \frac{55\sqrt{1-4\eta}}{1204224}c_{\Theta}^4 + \eta\left(-\frac{7272431}{2709504}-\frac{12343\sqrt{1-4\eta}}{1806336} + \frac{2001211c_{\Theta}^2}{1354752}\right.\nonumber\\
+	&\qquad\qquad\quad\left.
+-\frac{1517\sqrt{1-4\eta}}{903168}c_{\Theta}^2 + \frac{24725}{129024}c_{\Theta}^4 - \frac{55\sqrt{1-4\eta}}{200704}c_{\Theta}^4\right) + \eta^2\left(\frac{1965709}{2709504}-\frac{3797\sqrt{1-4\eta}}{150528}-\frac{1072289}{1354752}c_{\Theta}^2\right.\nonumber\\
+	&\qquad\qquad\quad\left.\left.\left.
+-\frac{411\sqrt{1-4\eta}}{25088}c_{\Theta}^2 - \frac{16785}{100325}c_{\Theta}^4 + \frac{55\sqrt{1-4\eta}}{150528}c_{\Theta}^4\right)\right]\chisdL\right\}\nonumber\\
+	\\
+	& H_{+}^{(3,mem)} =\alpha s_{\Theta}^2\left\{-\frac{69549014242469}{46146017820672} + \frac{6093995955001}{23073008910336}c_{\Theta}^2 - \frac{1416977186081}{15382005940224}c_{\Theta}^4-\frac{2455652411}{78479622144}c_{\Theta}^6-\frac{9979199}{2491416576}c_{\Theta}^8\right.\nonumber\\
+	&\qquad\qquad\quad + \eta^2\left[\frac{111707903}{594542592} - \frac{792654827}{3269984256}c_{\Theta}^2 -\frac{9226880251}{6539968512}c_{\Theta}^4 - \frac{71269747}{148635648}c_{\Theta}^6-\frac{7835}{98304}c_{\Theta}^8\right] + \eta\left[\frac{1355497792045}{149824733184}\nonumber\right.\\
+	&\qquad\qquad\quad \left.- \frac{3485\pi^2}{9216}-\frac{41463325921}{51502252032}c_{\Theta}^2 -\frac{205\pi^2}{9216} + \frac{347235904277}{549357355008}c_{\Theta}^4 + \frac{788232313}{3567255552}c_{\Theta}^6 + \frac{302431}{9437184}c_{\Theta}^8\right]+\eta^3\left[\frac{10034527}{445906944}\right.\nonumber\\
+	&\qquad\qquad\quad\left.
++ \frac{57932071}{1226244096}c_{\Theta}^2 + \frac{29847079}{544997376}c_{\Theta}^4 + \frac{8170691}{24772608}c_{\Theta}^6 + \frac{9065}{131072}c_{\Theta}^8\right] + \left[-\frac{14443855}{66060288}  + \frac{113\sqrt{1-4\eta}}{8192} + \frac{3716075}{33030144}c_{\Theta}^2\right.\nonumber\\
+	& \qquad\qquad\quad\left. + \frac{791\sqrt{1-4\eta}}{73728} + \frac{113265}{7340032}c_{\Theta}^4 + \eta\left(\frac{32497919}{16515072} - \frac{113\sqrt{1-4\eta}}{2048}\right)\right]
 \end{align}
@@
 This can be done by combining the PN expression during early inspiral and numerical relativity results for merger part.
-The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
+The method u	sed to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
 \subsection{Compleate Christodoulou memory during BBH coalescence} 

### 2019-01-28 notes/Paper.tex
updated paper

@@
 The equation \ref{eq:2.14} can be used to change the integration with respect to time into integration with respect to $x$ as shown in equation in \ref{eq:2.11}.
-The expressions for $h_{lm}$ whose first time derivatives appear equation \ref{eq:2.10} are directly taken from reference \cite{Arun2009}.
-The PN expression for $h_{lm}$ there are computed for precessing binaries(Equations 4.17 in ref \cite{Arun2009}), but here we are only interested in aligned spin case.
-The structure of modes for non-precessing binaries can be obtained by aligning the orbital angular momentum with the total angular momentum($\iota = 0, \alpha = \pi$).
-These results for a special case of non-precessing binaries, which contain spin contribution at 1.5PN order are shown below
+The expressions for $h_{lm}$ whose first time derivatives appear equation \ref{eq:2.10} can be obtained from Ref\cite{Arun2009}.
+The GW modes for precessing binaries with small inclination angle on nearly circular orbits through 1.5 PN order are given in section IV of ref \cite{Arun2009}.
+Setting $\iota = 0, \alpha = \pi$ in equation 4.16 and 4.17 gives the result for non-precessing case.
+In that case the spins are aligned along $z$ axis, so we set $\chi_a^x=\chi_s^x=\chi_a^y=\chi_s^y=0$.
+We can now write the expression as
+\begin{equation}
+h_{lm} = (-1)^{m+1}\frac{(2M\eta v^2)}{D_L}\sqrt{\frac{16\pi}{5}}e^{\mathit{i}\,m\,\psi}\hat{h}_{lm},
+\end{equation}
+and $\hat{h}_{lm}$ modes are given by 
 \begin{subequations}
@@
  We show the $h_{lm}$ expressions for positive $m$ values, but the results for negative $m$ values can be obtained by the relation, $h_{l, -m} = (-1)^l \bar{h}_{lm}$.
-This results is not in general true for precessing binaries, but it continues to hold true of non-pressing case.
+This results is not in general true for precessing binaries, but it continues to hold true of non-precessing case.
+The time derivative of $h_{lm}$ modes can be obtained by following the procedure shown in Appendix \ref{Appendix B}.
 With these modes we use equation \ref{eq:2.10} to evaluate the first time derivative of the memory mass-multipole moment, which is further integrated using equation \ref{eq:2.11}.
@@
 \begin{figure}
-	\includegraphics[width=3.0in]{../plots/PNmemoryContributiontoHplus.pdf}
-	\includegraphics[width=3.0in]{../plots/PNmemoryContributionfunofTheta.pdf}
+	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpin.pdf}
+	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}
+	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpin.pdf}
+	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
@@
 This can be done by combining the PN expression during early inspiral and numerical relativity results for merger part.
-The method u	sed to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
+The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
 \subsection{Compleate Christodoulou memory during BBH coalescence} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}
-	\includegraphics[width=4.0in]{/home/ashok/gravitational_wave_memory_project/plots/DerivativeComparisionNRvsPN0p800.pdf}
-	\caption{The figure above shows a comparison for time derivatives Christodoulou memory contribution to $h_+$ polarization between NR data and PN results }
-	\label{fig:derivative}
-\end{figure}
-\begin{figure}
- 	\includegraphics[width=4.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
+ 	\includegraphics[width=3.5in]{../plots/MemoryPlot_nonSpining/q7.pdf}
  	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 \end{figure}
-
 \begin{figure}
-	\includegraphics[width=4.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
-	\includegraphics[width=4.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 
-\section{Calculation Of The $\dot{h}_{lm}$ Modes}
+\section{Calculation Of The $\dot{h}_{lm}$ Modes}\label{Appendix B}
 The $h_{lm}$ modes are written as 
-\begin{equation}
+\begin{equation}\label{eq:B1}
 	h_{lm}=8\sqrt{\frac{\pi}{5}}\frac{\eta\,M}{R}x\hat{H}_{lm}(x)e^{\mathit{i}m\psi}
 \end{equation}
-where the $\hat{H}_{lm}$ given in ref[add ref] for the case of precessing binaries.
+where the $\hat{H}_{lm}$ are obtained by taking the non-precessing limit of expressions for precessing binaries given in Ref.\cite{Arun2009} .
 In this paper we are interested in nonprecessing case, so we rewrite the expressions such that spin are aligned with the total orbital angular momentum.
@@
 where $\ln x_0 = \frac{11}{18} - \frac{2}{3}\gamma_E -\frac{4}{3}\ln2 + \frac{2}{3}\ln\left(\frac{M}{r_0}\right)$ is related to the arbitrary constant $r_0$ appearing in the coordinate transformation of Eq.
-(2.20) or ref[add ref].
-The orbital phase can be expressed a function of x for nonprecessing binaries, to 3.5PN order.
+(2.20) or ref\cite{Favata2009}.
+The orbital phase can be expressed a function of x for nonprecessing binaries, to 3.5PN order.This expression is given in Eq(3.4) of Ref.
+\cite{Ajith_2011}
 \begin{align}
@@
 where $\varphi_0$ is a certain reference phase.
+The first time derivative of modes can be obtained by using equation \ref{eq:B1} and substituting the expression for $\dot{x}$ as a function of $x$ using Eq.
+\ref{eq:C3}.
+We use the expression for time derivative as show in equation below
+\begin{equation}
+	\dot{h}_{lm} = 8\sqrt{\frac{\pi}{5}}\frac{\eta\,M}{R}\dot{x}e^{-\mathit{i}m\psi}\left(\tilde{H}_{lm}' - \mathit{i}m\psi'\tilde{H}_{lm} \right)
+\end{equation}
+where the prime denotes the derivative with respect to $x$.
 
@@
 \\
-\label{Appendix:C3}
+\label{eq:C3}
 \frac{dx}{dt}&=\frac{64\, x^5  \, \eta}{5}\left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
@@
 
-\section{$h_{lm}$ modes for non-precessing binaries}\label{Appendix:D}
-The GW modes for precessing binaries with small inclination angle on nearly circular orbits through 1.5 PN order are given in section IV of ref \cite{Arun2009}.
-Setting $\iota = 0, \alpha = \pi$ in equation 4.16 and 4.17 gives the result for non-precessing case.
-In that case the spins are aligned along $z$ axis, so we set $\chi_a^x=\chi_s^x=\chi_a^y=\chi_s^y=0$.
-We can now write the expression as
-\begin{equation}
-	h_{lm} = (-1)^{m+1}\frac{(2M\eta v^2)}{D_L}\sqrt{\frac{16\pi}{5}}e^{\mathit{i}\,m\,\psi}\hat{h}_{lm},
-\end{equation}
-and $\hat{h}_{lm}$ modes are given by
-
-\begin{subequations}
-\begin{align}
-	\hat{h}_{22} &= 1 + x\left(-\frac{107}{42} + \frac{55\eta}{42}\right) + x^{3/2}\left(2\pi - \frac{4\,\delta\,\chi^{z}_{a}}{3} + \frac{4}{3}\left(-1+\eta\right)\chi^z_s\right)\\
-	\hat{h}_{21} &= \frac{\mathit{i}\delta}{3}\left[x^{1/2} - x\frac{3}{2\delta}\left(\chi^{z}_{a} + \delta\chi^{z}_{s}\right) + x^{3/2}\left(-\frac{17}{28} + \frac{5\eta}{7}\right)\right]\\
-	\hat{h}_{32} &= \frac{9}{8}\sqrt{\frac{5}{7}}\left[\frac{8}{27}x\left(1-3\nu\right) + \frac{32}{27}x^{3/2}\eta\,\chi^z_s\right]
-\end{align}
-\end{subequations}
-$\hat{h}_{lm}$ shown above are the only modes which has spin contributions.
-We only include spin contribution upto 1.5PN but keep the remaining expression upto 3PN.
-The complete expression upto 3.5PN are given in ref \cite{Blanchet_Iyer2008}
 \section*{References}

### 2019-02-11 notes/Paper.tex
updated paper

@@
 
-Gravitational wave memory effect results in a permanent displacement of test masses after the gravitational wave has passed through an "ideal" GW detector.
-There are two kinds of gravitational wave memory effects: linear and nonlinear.
-The linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments.
+Gravitational waves, when they pass through a pair of freeling falling test masses cause an oscillatory motion.
+This oscillatory motion after the gravitational wave has passed through it, does not bring back the pair of test particle to their original location.
+This permanent displacement is called gravitational wave memory effect.
+This displacement is caused by a permanent deformation of spacetime geometry caused by passing gravitational waves.
+The memory effect associated with gravitational waves is accumulated effect over the entire lifehistory of the source.
+The permanent deformation that it leaves in spacetime geometry can be described in two ways: The Linear memory, which is caused by net change in the time derivative of source of one more multipole moment and the Nonlinear memory, which is caused by radiatiated gravitational waves acting as source for generating secondary GWs.
+The linear memory is important for unbound astrophysical system and have been studied in context of supernova explosion, asymmetric mass loss due to neutrino emission, gamma-ray-burst jets etc \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}.
 At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
@@
 where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times. 
-Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion, asymmetric mass loss due to neutrino emission and gamma-ray-burst jets, \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}
-\par The linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by the following general formula for by Thorn \cite{Thorne_Braginskii1987}
-
+\par A general formula for the linear memory produced by an unbound system of N particle as a result of  change in  it’s mass $M_A$ and velocity configuration $v_A$ is given by Thorn \cite{Thorne_Braginskii1987} as shown below.
 \begin{equation}\label{eq:1.2}
@@
 
-where the masses are unbound in their initial and final states(or both).
  Here $\Delta$ refers to take difference between the final and initial values of the summations, and $\boldsymbol{N}$ is the unit vector that points from the source to the observer.  
@@
 The other kind of memory is the nonlinear memory, which is also called Christodoulou memory.
-This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}.
-Christodoulou arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave.
-It can be understood in the following way.
-Consider the Einstein field equation in harmonic gauge.
+It arises as results of change in the radiative multipole moment that is sourced by the energy flux of the radiated GWs.
+It can be understood by looking at Eisntein's equations in harmonic gauge and considering the leading order nonlinear terms.
+The equation below shows a wave equation for spacetime metric in harmonic gauge with different source terms appearing on the right.
 \cite{Thorne1980}
@@
 where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$.
-When we apply standard Green's function to right hand side of equation \ref{eq:3.1a}, we obtain the following correction term to GW field \cite{Wiseman1991}
+The standard procedure to solve a wave equation with source is by using Green’s function.
+When applied to  Eq.
+\ref{eq:3.1a} obtains a correction term to gravitational field sourced by the Landau-Lifshitz term is shown below \cite{Wiseman1991}
 \begin{equation}\label{eq:1.6}
@@
 This equation shows that part of the distant GW is sourced by loss of GW energy.
+This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}.
 The nonlinear memory occurs in any system that radiates GWs.
@@
 The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992}
-\par The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity, it arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments.
-The nonlinear memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order.
-This memory effect causes a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral.
+\par The Christodoulou memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order.
+The effect of Christodoulou memory on GW waveform is a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral.
 The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work.
@@
 Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.
-These $m=0$ modes tend to be much smaller in magnitude and depends more sensitively on	initial conditions than other modes.
+These $m=0$ modes are much smaller in magnitude and depends more sensitively on	initial conditions than other modes.
 For detailed discussion on this see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009}
@@
 \end{equation} 
-In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$.
-There are many other difficulties too in calculating memory from numerical relativity data.
+In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.
+The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$.
+There are other difficulties too in calculating memory from numerical relativity data.
 Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified.
-Choosing these integration constant to be zero \cite{Reisswig_Pollney2011} causes an artificial memory and a slope in the $h_{lm}$ modes.
-This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.
+Choosing these integration constant to be zero \cite{Reisswig_Pollney2011}  does not make the signal oscillate around zero.
+So there is an artificial memory in the $h_{lm}$ modes which is caused by numerical noise.
+This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.
+The initial burst of junk radiation also contribute to artificial memory.
 \par Numerical relativity simulations starts at a finite radius, so although it contains most part of gravitational wave memory, PN results should be used for early inspiral phase.
@@
 \par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata \cite{Favata2009}.
-However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})
+However the effect of binary spins have not been explored in these results.
+Astrophysical black holes are formed by gravitation collapse of massive star when when they run out of their nuclear fuel in the core.
+These massive stars are likely to posses spin angular momentum, as a result the black hole formed would also be spinning.
+The black holes can also accrete matter from surrounding and as a result spin up during the process.
+It is highly likely that there is a large population of spinning binaries in the universe.
 \par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.
@@
 
-The mode decomposition of gravitational wave polarization is given by
+The mode decomposition in terms spin-weighted spherical harmonics for gravitational wave polarization is given by
 \begin{equation}\label{eq:2.1}
@@
 + \frac{57932071}{1226244096}c_{\Theta}^2 + \frac{29847079}{544997376}c_{\Theta}^4 + \frac{8170691}{24772608}c_{\Theta}^6 + \frac{9065}{131072}c_{\Theta}^8\right] + \left[-\frac{14443855}{66060288}  + \frac{113\sqrt{1-4\eta}}{8192} + \frac{3716075}{33030144}c_{\Theta}^2\right.\nonumber\\
-	& \qquad\qquad\quad\left. + \frac{791\sqrt{1-4\eta}}{73728} + \frac{113265}{7340032}c_{\Theta}^4 + \eta\left(\frac{32497919}{16515072} - \frac{113\sqrt{1-4\eta}}{2048}\right)\right]
+	& \qquad\qquad\quad\left.
++ \frac{791\sqrt{1-4\eta}}{73728} + \frac{113265}{7340032}c_{\Theta}^4 + \eta\left(\frac{32497919}{16515072} - \frac{113\sqrt{1-4\eta}}{2048} -\frac{8200699}{8257536}c_{\Theta}^2 - \frac{791\sqrt{1-4\eta}}{18432}c_{\Theta}^2 - \frac{701315}{5505024}c_{\Theta}^4\right)\nonumber\right.\\
+	& \qquad\qquad\quad\left. + \eta^2\left(-\frac{142301}{36864} + \frac{40501}{18432}c_{\Theta}^2 + \frac{1075}{4096}c_{\Theta}^4\right)\right]\chiaSqr + 
+	\left[\frac{4675\pi}{1152}-\frac{9\pi\sqrt{1-4\pi}}{2048} + \frac{275\pi}{1152}c_{\Theta}^2 - \frac{7\pi\sqrt{1-4\pi}}{2048}c_{\Theta}^2 + \eta\left(-\frac{947\pi}{288}\right.\right.\nonumber\\
+	& \qquad\qquad\quad + \left.\left.\frac{9\pi\sqrt{1-4\eta}}{512} -\frac{91\pi}{288}c_{\Theta}^2 + \frac{7\pi\sqrt{1-4\eta}}{512}c_{\Theta}^2\right)\right]\chisdL + \left[-\frac{14443855}{66060288} + \frac{113\sqrt{1-4\eta}}{8192} + \frac{3716075}{33030144}c_{\Theta}^2 + \frac{791\sqrt{1-4\eta}}{73728}c_{\Theta}^2\right.\nonumber\\
+	& \qquad\qquad\quad + \frac{113265}{7340032}c_{\Theta}^4 + \eta^3\left(-\frac{11}{192} - \frac{11}{384}c_{\Theta}^2\right) + \eta\left(\frac{15798397}{2752512} - \frac{33\sqrt{1-4\eta}}{512} + \frac{921469}{4128768}c_{\Theta}^2 - \frac{77\sqrt{1-4\eta}}{1536} - \frac{184955}{2752512}c_{\Theta}^4\right)\nonumber\\
+	& \qquad\qquad\quad \left.+ \eta^2\left(-\frac{9507487}{4128768} + \frac{19\sqrt{1-4\eta}}{512} - \frac{676133}{2064384}c_{\Theta}^2 +\frac{133\sqrt{1-4\eta}}{4608}c_{\Theta}^2+ \frac{11745}{458752}c_{\Theta}^4\right)\right]\chisSqr + \left[-\frac{9\pi}{2048} + \frac{4675\pi\sqrt{1-4\eta}}{1152}\right.\nonumber\\
+	& \qquad\qquad\quad - \frac{7\pi}{2048}c_{\Theta}^2 + \frac{275\pi\sqrt{1-4\eta}}{1152}c_{\Theta}^2 + \eta\left(\frac{9\pi}{512} + \frac{7\pi}{512}c_{\Theta}^2\right) + \left(\frac{113}{4096} -\frac{14443855\sqrt{1-4\eta}}{33030144} + \frac{791}{36864}c_{\Theta}^2 + \frac{3716075\sqrt{1-4\eta}}{16515072}c_{\Theta}^2\right.\nonumber\\
+	 & \qquad\qquad\quad +\frac{113265\sqrt{1-4\eta}}{3670016}c_{\Theta}^4 + \eta^2\left(\frac{33}{128} + \frac{11\sqrt{1-4\eta}}{384} + \frac{77}{384}c_{\Theta}^2 + \frac{11\sqrt{1-4\eta}}{768}c_{\Theta}^2\right) + \eta\left(-\frac{179}{1024} + \frac{56422223\sqrt{1-4\eta}}{8257536}\right.\nonumber\\
+	 & \qquad\qquad\quad \left.\left.\left.\left.
+-\frac{1253}{9216}c_{\Theta}^2 -\frac{440281\sqrt{1-4\eta}}{1376256}c_{\Theta}^2 -\frac{17415\sqrt{1-4\eta}}{131072}c_{\Theta}^4\right)\right)\chisdL\right]\right\}
 \end{align}
@@
 	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpin.pdf}
-	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}
+	\includegraphics[width=3.5in]{../plots/PNmemoryFavata.pdf}
 	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpin.pdf}
-	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
@@
 \end{widetext}
+
+\begin{figure}
+	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}
+	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}
+	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
+The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4 $ .
+The plot to the right shows $x H_+$ as a function of parameter x.
+Increasing the magnitude of 𝜒 does not change the angular dependence, but does tend to increase the magnitude of the memory.
+	}
+\end{figure}
 \section{Memory calculation using Numerical relativity data}
@@
 Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$.
-It has long been noted that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}.
+It has long been known that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}.
 The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation.
@@
 We used this data and  performed a frequency domain time integration as described in equation \ref{3.6}, while removing the unphysical frequency by choosing a lower cutoff as show in section \ref{FFI}.
-The cutoff frequency is chosen as $\omega_0/10$ where $\omega_0$ is the instantaneous GW angular frequency.
+The cutoff frequency is chosen as $\omega_0/10$ where $\omega_0$ is the minimum instantaneous GW angular frequency.
 We first begin by separating the complex Newman-Perose scalar $\psi_4$ data into phase and amplitude part.
@@
 \subsection{Compleate Christodoulou memory during BBH coalescence} 
-The PN waveform are given as a function of parameter $x$ which is related to time by equation \ref{eq:2.14}.
+The PN waveform for Christodoulou are given as a function of parameter $x$ which is related to time by equation \ref{eq:2.14}.
 This equation can be integrated to obtain a relation between time $t$ and parameter $x$.
@@
 We choose the values of integration constants such that the peak in $h_+^{(mem)}$ is reached at time $t=0$.
-For comparison we show the time derivative of $h_+^{(mem)}$ from Numerically obtained results and the analytical expressions in figure \ref{fig:derivative}.
-The slope as seen from figure \ref{fig:derivative} rises asymptotically.
-We use our PN results for computing the memory during early inspiral and attach them to NR results.
+We use our PN results for computing the memory during early inspiral and attach them to NR results near the late inspiral.
 We compare this memory for different mass ratio and different values of dimensionless spin parameter.
@@
 It can be seen that the spin magnitude has significant effect on nonlinear memory, higher spin magnitude, in case of spin aligned with orbital angular momentum tends to increase the memory contribution.
-When the spins are anti aligned with orbital angular momentum, the memory tend to decrease the Christodoulou memory.
-These results show it important to consider the spin contribution when studying the Christodoulou memory in gravitational wave signals.
+When the spins are anti aligned with orbital angular momentum, it tend to decrease the Christodoulou memory.
+These results show that it is important to consider the spin contribution when studying the Christodoulou memory in gravitational wave signals.
+\subsection{CONCLUSIONS}
+We find that spin effects are significantly important for observations that aims at detecting Chrisodoulou memory.
+
+\section{Acknowledgments}
+We thank the West Virginia Univerisity. This work was support by ...
 
@@
 \begin{figure}
- 	\includegraphics[width=3.5in]{../plots/MemoryPlot_nonSpining/q7.pdf}
+ 	\includegraphics[width=4.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
  	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.

### 2019-02-12 notes/Paper.tex
updated paper

@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
-	\includegraphics[width=3.5in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
+	\includegraphics[width=4.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
+	\includegraphics[width=4.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.

### 2019-02-13 notes/Paper.tex
updated notes

@@
 
-Gravitational waves, when they pass through a pair of freeling falling test masses cause an oscillatory motion.
+Gravitational waves(GW), when they pass through a pair of freeling falling test masses cause an oscillatory motion.
 This oscillatory motion after the gravitational wave has passed through it, does not bring back the pair of test particle to their original location.
@@
 The memory effect associated with gravitational waves is accumulated effect over the entire lifehistory of the source.
-The permanent deformation that it leaves in spacetime geometry can be described in two ways: The Linear memory, which is caused by net change in the time derivative of source of one more multipole moment and the Nonlinear memory, which is caused by radiatiated gravitational waves acting as source for generating secondary GWs.
+The permanent deformation that it leaves in spacetime geometry can be described in two ways: The Linear memory, which is caused by net change in the time derivative of source of one more multipole moment and the Nonlinear memory, which is caused by radiatiated gravitational waves acting as source for generating secondary gravitational waves.
 The linear memory is important for unbound astrophysical system and have been studied in context of supernova explosion, asymmetric mass loss due to neutrino emission, gamma-ray-burst jets etc \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}.
@@
 is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator.
-The Landau-Lifshitz term is the Gravitational wave stress tensor~\cite{Isaacson1968}.
+We can see that there are many source terms.The Landau-Lifshitz term is the Gravitational wave stress tensor which is the source of nonlinear memory contribution to GW waveform.~\cite{Isaacson1968}.
 \begin{equation}
@@
 where $T_{R}$ is the retarded time.
-This equation shows that part of the distant GW is sourced by loss of GW energy.
+The above equation \ref{eq:1.6} gives the nonlinear Landau-Lifshitz pseudotensor term contribution to gravitational wave memory.
 This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}.
@@
 \par The Christodoulou memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order.
-The effect of Christodoulou memory on GW waveform is a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral.
+The $+$ polarization of the gravitational wave signal starts to ossillates around zero and slowly starts to deviate dues to Chritodoulou memory effect.
+This oscillatory shift in the amplitude of $+$ polarization starts at early times and slowly grows during inspiral phase.
 The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work.
-In this case the memory must be extracted from Numerical relativity simulations.
+In this case the memory must be extracted from Numerical relativity (NR) simulations.
 However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes.
-Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.
+NR simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.
 These $m=0$ modes are much smaller in magnitude and depends more sensitively on	initial conditions than other modes.
@@
 Choosing these integration constant to be zero \cite{Reisswig_Pollney2011}  does not make the signal oscillate around zero.
+Although it should not if memory is present, but this nonlinear growth is not solely due to actual memory effect.
+This can also be caused by numerical noise.
 So there is an artificial memory in the $h_{lm}$ modes which is caused by numerical noise.
@@
 \par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.
+For PN calculations, this paper is just an extension of Favata's\cite{Favata2009} paper, where we include the spin contribution upto leading (1.5 PN) order for nonprecessing binaries.
+The nonlinear memory contribution near merger is computed using NR data from publicly available SXS catalog, where we process the data to remove the low frequency numerical noise.
 In Sec.
@@
 \end{equation}
-where we have followed the notation of Sec. 9 of Ref. \cite{Blanchet_Iyer2008}. The notational parameter $\alpha$ accounts for the two commonly used choices for the polarization triad 
+where we have followed the notation of Sec. 9 of Ref. \cite{Blanchet_Iyer2008}. The parameter $\alpha$ takes the value as shown below, depending upon the polarization triad chosen. 
 \begin{equation}
@@
 \end{subequations}
+The effect of spin on Christodoulou memory can be seen in fig \ref{fig:HpVsX}, which shows $xH_+$ as a function of $x$.
+The top left plot in fig \ref{fig:HpVsX} show $xH_+$ as a function of $x$ for different values of dimensionless antisymmetric spin parameters $\chis$ values.
+It can be seen that with increasing the $\chis$ values for aligned spin case tends to increase the Christodoulou memory.
+It was shown by Favata \cite{Favata2009} that increasing the PN order tends to decrease the memory contribution.
+For comparison we have shown the plot in Favata \cite{Favata2009} on the top right in fig \ref{fig:HpVsX}.
+The bottom plot in fig \ref{fig:HpVsX} show the effect of $\chis$ in anti aligned case.
 \begin{figure}
@@
 	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpin.pdf}
-	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
-The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4 $ .
-The plot to the right shows $x H_+$ as a function of parameter x.
-Increasing the magnitude of 𝜒 does not change the angular dependence, but does tend to increase the magnitude of the memory.
+	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the orbital separation.
+The plots above show the memory contribution to $H_+$  as a function of $x$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $\eta=1/4 $ .
+The top left plot shows $x H_+$ as a function of parameter x which is a function of time.
+In plot we consider aligned spin case, where the individual spins are aligned with orbital angular momentum.
+It can be seen that increasing the spin magnitude tend to increase the nonlinear memory contribution.
+The top right figure shows the effect of including the higher PN terms to memory.
+this is just for comparison with Favata's results \cite{Favata2009}.
+The bottom figure show the effect of spin when it is anti aligned with orbital angular momentum.
 		}
-	\label{fig:0p95}
+	\label{fig:HpVsX}
 \end{figure}
@@
 These results show that it is important to consider the spin contribution when studying the Christodoulou memory in gravitational wave signals.
-\subsection{CONCLUSIONS}
-We find that spin effects are significantly important for observations that aims at detecting Chrisodoulou memory.
 
-\section{Acknowledgments}
-We thank the West Virginia Univerisity. This work was support by ...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{figure}
+
+\subsection{CONCLUSIONS}
+Observation of Christodoulou memory would be an important test of general relativity in strong field regime.
+Effort to detect this memory effect in low frequency band from super massive black holes have been undertaken by pulsar timing array experiments.
+In these experiment the spin effect have been neglected for modeling the signal.
+It is highly likely that super massive black holes have significant spin due to accretion of matter from surrounding.
+We have shown in this paper that spin contribution can significantly change the Christodoulou memory.
+Hence it is important to consider the spin effect while modeling the signals.
+
+\section{Acknowledgments}
+We thank the West Virginia Univerisity. This work was support by ...
 

### 2019-02-14 notes/Paper.tex
updated references

@@
 Observation of Christodoulou memory would be an important test of general relativity in strong field regime.
-Effort to detect this memory effect in low frequency band from super massive black holes have been undertaken by pulsar timing array experiments.
+Effort to detect this memory effect in low frequency band from super massive black holes have been undertaken by pulsar timing array experiments \cite{Pshirkov2010}.
 In these experiment the spin effect have been neglected for modeling the signal.

### 2019-02-18 notes/Paper.tex
update notes

@@
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
-The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4 $ .
-The plot to the right shows $x H_+$ as a function of parameter x.
-Increasing the magnitude of 𝜒 does not change the angular dependence, but does tend to increase the magnitude of the memory.
-	}
+The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4$.}
 \end{figure}

### 2019-03-22 notes/Paper.tex
some edits

@@
-\documentclass[prd,preprintnumbers,twocolumn,eqsecnum,floatfix,letter]{revtex4}
+\documentclass[twocolumn,showpacs,aps,prd,nobibnotes,floatfix]{revtex4-1}
+\usepackage[hmargin=1.5cm,vmargin=1.5cm,bindingoffset=0.0cm]{geometry}
+\usepackage{longtable}
+%\usepackage{graphics,epsfig,placeins,subfigure,wrapfig}
+\usepackage{graphics,placeins,subfigure,wrapfig}
+\usepackage[usenames]{color}
+\usepackage[export]{adjustbox}
 \usepackage{color}
@@
 \usepackage[varg]{txfonts}
+\usepackage{soul}
 \usepackage[colorlinks, pdfborder={0 0 0}]{hyperref}
@@
 \allowdisplaybreaks
-\textwidth 7.5 in 
-\hoffset -1 cm 
+%\textwidth 7.5 in 
+%\hoffset 0 cm 
 \newcommand{\RomanNumeralCaps}[1]{\MakeUppercase{\romannumeral #1}}
@@
 \newcommand{\ashok}[1]{\textcolor{red}{\textit{Ashok:#1}}}
-\newcommand{\sean}[1]{\textcolor{cyan}{\textit{Sean:#1}}}
+\newcommand{\stm}[1]{\textcolor{blue}{\textit{Sean:#1}}}
 
@@
 
-\title{Effect of Black Hole Spin on Christodoulou Memory}
+\title{Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers}
 \author{Ashok Choudhary}\email{aschoudhary@mix.wvu.edu}
@@
 \affiliation{Department of Physics and Astronomy, West Virginia University, Morgantown, WV 26506, USA}
+\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Building, Morgantown, WV 26505, USA}
 
+\stm{note everywhere I have added ``the'', that ``upto'' should be ``up to'', etc}
 \begin{abstract}
-	The Christodoulou memory, which is a nonlinear memory effect sourced by the gravitational wave stress tensor, produce a growing, nonoscillatory change in the gravitational wave "plus" polarization.
+	The Christodoulou memory, which is a nonlinear memory effect sourced by the oscillatory component of the gravitational wave stress tensor, produces a growing, nonoscillatory change in the gravitational wave \st{"plus" polarization} strain \stm{the ``plus'' polarization isn't a fundamental quantity, its a gauge choice}.
 This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed.
-While the memory contribution during early in-spiral is well described by Post-Newtonian approximation, the contribution at merger must be obtained from Numerical relativity simulations.
-The Post-Newtonian corrections upto 3PN order to the gravitational wave memory for quasicircular, inspiralling compact binaries for non spinning has been computed by Favata et al.[Physical Review D 80, 024002 (2009)].
-Here we include the Spin contribution at leading order and calculate the memory effect upto 3PN order.
-We also compute memory contribution at merger from Numerical relativity using Weyl curvature component $\psi_4$ after removing the unphysical low frequency contribution due to spectral leakage.
+While the memory contribution during the early inspiral is well described by the post-Newtonian (PN) approximation, the contribution at merger must be obtained using numerical relativity simulations.
+The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata et al.
+\stm{is it Favata et al, or just Favata by himself?
+use bibtex for references, get style file for PRD so that they appear correctly}.
+Here we include the contribution from the individual black-hole spins at leading order \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order.
+We also compute the memory contribution during
+the merger by using the methods presented in \stm{cite Favata} to calculate the memory content using the oscillatory radiation as input, and
+applying that method using publicly available numerical relativity simulations of spinning binaries.
+\st{using Weyl curvature component $\psi_4$ after removing the unphysical low frequency contribution due to spectral leakage.
 We use the method described in [Class.
 Quantum Grav.
-28 195015] for integration which control the amplification of the unphysical frequency resulting from spectral leakage.
+28 195015] for integration which control the amplification of the unphysical frequency resulting from spectral leakage.}\stm{This is too much detail form the abstract, belongs in the body of the paper}
 \end{abstract}
@@
 
-Gravitational waves(GW), when they pass through a pair of freeling falling test masses cause an oscillatory motion.
-This oscillatory motion after the gravitational wave has passed through it, does not bring back the pair of test particle to their original location.
-This permanent displacement is called gravitational wave memory effect.
-This displacement is caused by a permanent deformation of spacetime geometry caused by passing gravitational waves.
-The memory effect associated with gravitational waves is accumulated effect over the entire lifehistory of the source.
-The permanent deformation that it leaves in spacetime geometry can be described in two ways: The Linear memory, which is caused by net change in the time derivative of source of one more multipole moment and the Nonlinear memory, which is caused by radiatiated gravitational waves acting as source for generating secondary gravitational waves.
-The linear memory is important for unbound astrophysical system and have been studied in context of supernova explosion, asymmetric mass loss due to neutrino emission, gamma-ray-burst jets etc \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}.
-At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
-\begin{equation}
-\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
-\end{equation} 
-where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times. 
-\par A general formula for the linear memory produced by an unbound system of N particle as a result of  change in  it’s mass $M_A$ and velocity configuration $v_A$ is given by Thorn \cite{Thorne_Braginskii1987} as shown below.
-\begin{equation}\label{eq:1.2}
-\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
-\end{equation}
+Gravitational waves (GW), when they pass through a pair of freeling falling test masses, induce an oscillatory motion.
+This motion is incredibly small for astrophysical sources.
+Nonetheless, the signals generated by ten black-hole binaries and one neutron-star binary were detected by the Advanced LIGO and Advanced Virgo \stm{cite the catalog paper}.
+For sources whose strength changes significantly with time, such as the final merger of compact binaries like those that have already been observed, there is an additional non-oscillatory signal that results in a permanent displacement of
+freely falling test masses.
+This permanent displacement, which is called the Christodoulou memory effect \cite{Christodoulou1991}, is sourced by the oscillatory part of the signal, and is therefore a fundamental manifestation of nonlinearity in general relativity.
+This memory effect accumulates over the entire history of the source.
+This nonlinear memory effect is distinct from linear memory, which is due to a change in the time derivative of one or more multipole moments of a source \stm{give a more specific definition}.
+The linear memory is important for unbound astrophysical systems and have been studied in context of supernova explosions \stm{put one or more citations after each item}, asymmetric mass loss due to neutrino emission \stm{isn't this also in the context of supernovae?
+if not, what source is this?}, and gamma-ray-burst jets \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}.
 
- Here $\Delta$ refers to take difference between the final and initial values of the summations, and $\boldsymbol{N}$ is the unit vector that points from the source to the observer.  
-\\
-The other kind of memory is the nonlinear memory, which is also called Christodoulou memory.
-It arises as results of change in the radiative multipole moment that is sourced by the energy flux of the radiated GWs.
-It can be understood by looking at Eisntein's equations in harmonic gauge and considering the leading order nonlinear terms.
-The equation below shows a wave equation for spacetime metric in harmonic gauge with different source terms appearing on the right.
-\cite{Thorne1980}
+\stm{I commented out the detailing of linear memory, since the paper is about nonlinear. The above mention of linear is sufficient.}
+%At leading order in a post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
+%\begin{equation}
+%\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
+%\end{equation} 
+%where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times. 
+%\par A general formula for the linear memory produced by an unbound system of N particle as a result of  change in  it’s mass $M_A$ and velocity configuration $v_A$ is given by Thorn \cite{Thorne_Braginskii1987} as shown below.
+%\begin{equation}\label{eq:1.2}
+%\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
+%\end{equation}
+
+% Here $\Delta$ refers to take difference between the final and initial values of the summations, and $\boldsymbol{N}$ is the unit vector that points from the source to the observer.
+
+Christodoulou memory arises from a change in the radiative multipole moment, and is therefore sourced by the energy flux of radiated GWs.
+It can be understood by considering Einstein's equation in the harmonic gauge, and calculating the leading order nonlinear contribution.
+The resulting wave equation for the spacetime metric can be expressed as \cite{Thorne1980}
 \begin{subequations}
@@
 \end{equation}
-is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator.
-We can see that there are many source terms.The Landau-Lifshitz term is the Gravitational wave stress tensor which is the source of nonlinear memory contribution to GW waveform.~\cite{Isaacson1968}.
+is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the d'Alembertian, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is the flat-space Laplacian.
+
+This wave equation has several source terms. The Landau-Lifshitz term represents the contribution of GWs to the stress tensor,
+and therefore serves as the source of the nonlinear memory contribution to GW waveform \cite{Isaacson1968},
+\stm{You appear to change the meanings of T and t, which you shouldn't do, but please verify that my correction is what you actually meant.
+Also, since the contravariant form appears in the above equation, you should give that here, rather than the covariant form.}
 \begin{equation}
-	T^{gw}_{jk}=\frac{1}{32\pi}\langle h^{TT}_{ab,j}h^{TT}_{ab,k}\rangle\approx T^{gw}_{00}n_{j}n_{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n_{j}n_{k}
+	t^{LL}_{jk}\equiv \frac{1}{32\pi}\langle h^{TT}_{ab,j}h^{TT}_{ab,k}\rangle\approx t^{LL}_{00}n_{j}n_{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n_{j}n_{k}
 \end{equation}
-where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$.
-The standard procedure to solve a wave equation with source is by using Green’s function.
-When applied to  Eq.
-\ref{eq:3.1a} obtains a correction term to gravitational field sourced by the Landau-Lifshitz term is shown below \cite{Wiseman1991}
+where $TT$ refers to the transverse-traceless gauge, $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-brackets correspond to an average over several wavelengths, \stm{what does this last one mean?} $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$.
+The standard procedure to solve a wave equation with a source is to use the Green's function.
+When this approach is applied to Eq.
+\ref{eq:3.1a}, we obtain a correction term to the gravitational field sourced by the Landau-Lifshitz term, given by \cite{Wiseman1991}
 \begin{equation}\label{eq:1.6}
@@
 where $T_{R}$ is the retarded time.
-The above equation \ref{eq:1.6} gives the nonlinear Landau-Lifshitz pseudotensor term contribution to gravitational wave memory.
-This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}.
-The nonlinear memory occurs in any system that radiates GWs.
-These systems are usually considered to have a vanishing linear memory.
-The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992}
-\par The Christodoulou memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order.
-The $+$ polarization of the gravitational wave signal starts to ossillates around zero and slowly starts to deviate dues to Chritodoulou memory effect.
-This oscillatory shift in the amplitude of $+$ polarization starts at early times and slowly grows during inspiral phase.
-The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work.
-In this case the memory must be extracted from Numerical relativity (NR) simulations.
-However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes.
-NR simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode.
-These $m=0$ modes are much smaller in magnitude and depends more sensitively on	initial conditions than other modes.
-For detailed discussion on this see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009}
-\par The NR simulations mostly easily compute  the curvature scalar $\Psi_4$ and decompose its values at large $R$ into a sum over spin-weighted spherical harmonic modes:
+Eq.~\ref{eq:1.6} is the nonlinear Landau-Lifshitz pseudotensor describing gravitational wave memory \cite{Christodoulou1991, Payne1894, Blanchet_Damour1992}.
+The nonlinear memory occurs in any system that radiates GWs, whereas these systems are usually considered to have a vanishing linear memory, although we can apply the formalism for linear memory to nonlinear memory if
+we consider the radiated gravitons to be the unbound masses \cite{Thorne1992}.
+%The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992}
+
+\par The Christodoulou memory effect arises from nonlinear interactions at 2.5PN order, but it affects the GWs at leading (0PN) order.
+In the transverse-traceless gauge, the $+$ polarization of the GW signal no longer oscillates around zero when memory is included, but instead accumulates a nonzero drift that remains small compared to the amplitude of the oscillatory contribution to the GWs throughout the inspiral, and becomes largest during the final merger.
+We can use PN methods to model the slow growth of memory during the inspiral phase, but the largest memory contribution that occurs during the merger requires information from numerical relativity (NR), since the PN approach breaks down at that stage.
+In principle, the memory can be extracted directly from NR simulations.
+However, it is challenging to extract the memory accurately.
+NR simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform, but the nonlinear memory is present in only $m=0$ mode.
+These $m=0$ modes are much smaller in magnitude and can be masked by numerical noise as well as subtleties of the numerical techniques used to extract the radiation.
+For a detailed discussion of these issues, see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009} \stm{the main reference for this is Reisswig and Pollney, and you should actually read that and include a little more detail here}
+\par NR simulations that use the most widely adopted methods directly compute the Weyl curvature scalar, $\Psi_4$, and decompose its values at large radii $R$ into a sum over spin-weighted spherical harmonic modes:
 \begin{equation}
@@
 The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$.
-There are other difficulties too in calculating memory from numerical relativity data.
+There are other difficulties too in calculating memory from NR data.
 Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified.
@@
 The initial burst of junk radiation also contribute to artificial memory.
-\par Numerical relativity simulations starts at a finite radius, so although it contains most part of gravitational wave memory, PN results should be used for early inspiral phase.
+\par NR simulations starts at a finite radius, so although it contains most part of gravitational wave memory, PN results should be used for early inspiral phase.
 The significance of memory accumulated during the inspiral phase can be seen by looking at the largest memory mode $h_{20}$.
@@
 \end{figure}
-\section{Memory calculation using Numerical relativity data}
-Numerical relativity simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by
+\section{Memory calculation using NR data}
+NR simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by
 \begin{equation}
@@
 The strain $h$ is only related to $\psi_4$ at an infinite distance and also strictly valid in a particular gauge.
-As a result in numerical simulation, the finite distance calculation introduces a systematic error.
+As a result in NR simulation, the finite distance calculation introduces a systematic error.
 In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
@@
 \cite{Reisswig_Pollney2011} that an important source of unphysical non-linear drift in numerical computation of gravitational wave strain lies in the transformation of the measured data to the strain $h$  which generally involves an integration in time.
-The output of the numerical simulation is a discretely sampled time series of finite duration, incorporating some component of unresolved frequencies due to numerical error.
+The output of the NR simulation is a discretely sampled time series of finite duration, incorporating some component of unresolved frequencies due to numerical error.
 This can lead to an uncontrollable non-linear drift if the integration is performed in the time domain.
@@
 The above equation requires a time integration of absolute value of $\dot{h}_{22}$, but this is not typically the quantity which is directly computed in numerical simulation.
-The output of numerical simulation is usually in the form of component of curvature tensor, or Zerilli-Moncrief-type variables defined relative to a background.
+The output of NR simulation is usually in the form of component of curvature tensor, or Zerilli-Moncrief-type variables defined relative to a background.
 \subsection{Evaluating Bondi News from numerical data}
@@
 The SXS catalog gives data for different parameter values like mass ratio, dimensionless spin parameter etc.
-The Christodoulou memory computed from Numerical Relativity data starts from zeros using equation \ref{eq:3.2} due to finite simulation time.
+The Christodoulou memory computed from NR data starts from zeros using equation \ref{eq:3.2} due to finite simulation time.
 Chrostodoulou memory is hereditary, hence it's entire past should be included.
-This can be done by combining the PN expression during early inspiral and numerical relativity results for merger part.
+This can be done by combining the PN expression during early inspiral and NR results for merger part.
 The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section.
@@
 It can be seen that higher mass ration tends to decrease the Chistodoulou memory.
-The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using Numerical Relativity data.
+The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using NR data.
 The red dashed line shows result from PN expression}
@@
 
-\bibliographystyle{apsrev4-1}
+%\bibliographystyle{apsrev4-1}
+\bibliographystyle{prsty}
 \bibliography{references}

### 2019-03-27 notes/Paper.tex
some corrections

@@
 \begin{abstract}
-	The Christodoulou memory, which is a nonlinear memory effect sourced by the oscillatory component of the gravitational wave stress tensor, produces a growing, nonoscillatory change in the gravitational wave \st{"plus" polarization} strain \stm{the ``plus'' polarization isn't a fundamental quantity, its a gauge choice}.
+	The Christodoulou memory, which is a nonlinear memory effect sourced by the oscillatory component of the gravitational wave stress tensor, produces a growing, nonoscillatory change in the gravitational wave strain.
 This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed.
 While the memory contribution during the early inspiral is well described by the post-Newtonian (PN) approximation, the contribution at merger must be obtained using numerical relativity simulations.
-The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata et al.
-\stm{is it Favata et al, or just Favata by himself?
-use bibtex for references, get style file for PRD so that they appear correctly}.
-Here we include the contribution from the individual black-hole spins at leading order \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order.
+The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata \cite{Favata2009}.
+Here we include the contribution from the individual black-hole spins up to 1.5PN \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order.
 We also compute the memory contribution during
-the merger by using the methods presented in \stm{cite Favata} to calculate the memory content using the oscillatory radiation as input, and
+the merger by using the methods presented in \cite{Favata2009} to calculate the memory content using the oscillatory radiation as input, and
 applying that method using publicly available numerical relativity simulations of spinning binaries.
-\st{using Weyl curvature component $\psi_4$ after removing the unphysical low frequency contribution due to spectral leakage.
-We use the method described in [Class.
-Quantum Grav.
-28 195015] for integration which control the amplification of the unphysical frequency resulting from spectral leakage.}\stm{This is too much detail form the abstract, belongs in the body of the paper}
 \end{abstract}
@@
 This motion is incredibly small for astrophysical sources.
-Nonetheless, the signals generated by ten black-hole binaries and one neutron-star binary were detected by the Advanced LIGO and Advanced Virgo \stm{cite the catalog paper}.
+Nonetheless, the signals generated by ten black-hole binaries and one neutron-star binary were detected by the Advanced LIGO and Advanced Virgo\cite{Ligocatalog2018} \stm{cite the catalog paper}.
 For sources whose strength changes significantly with time, such as the final merger of compact binaries like those that have already been observed, there is an additional non-oscillatory signal that results in a permanent displacement of
@@
 This memory effect accumulates over the entire history of the source.
-This nonlinear memory effect is distinct from linear memory, which is due to a change in the time derivative of one or more multipole moments of a source \stm{give a more specific definition}.
-The linear memory is important for unbound astrophysical systems and have been studied in context of supernova explosions \stm{put one or more citations after each item}, asymmetric mass loss due to neutrino emission \stm{isn't this also in the context of supernovae?
-if not, what source is this?}, and gamma-ray-burst jets \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}.
+\ashok{This nonlinear memory effect is distinct from linear memory, which arises from the non-oscillatory motion of unbounded source} \stm{give a more specific definition}.
+\ashok{The linear memory is important for unbound astrophysical systems and have been studied in context of supernova explosions resulting in neutron star kicks \cite{Burrows1996}\cite{Muller2004}\stm{put one or more citations after each item}, asymmetric mass loss due to neutrino emission \cite{COtt2009} \stm{isn't this also in the context of supernovae?
+if not, what source is this?}, and gamma-ray-burst jets \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}}.
 
@@
 \begin{equation}
-	t^{LL}_{jk}\equiv \frac{1}{32\pi}\langle h^{TT}_{ab,j}h^{TT}_{ab,k}\rangle\approx t^{LL}_{00}n_{j}n_{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n_{j}n_{k}
+	t_{LL}^{jk}\equiv \frac{1}{32\pi}\langle h_{TT}^{ab,j}h_{TT}^{ab,k}\rangle\approx t_{LL}^{00}n^{j}n^{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n^{j}n^{k}
 \end{equation}
@@
 
-%\bibliographystyle{apsrev4-1}
-\bibliographystyle{prsty}
+\bibliographystyle{apsrev4-1}
 \bibliography{references}

### 2019-03-27 notes/Paper.tex
some corrections

@@
 \par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.
-For PN calculations, this paper is just an extension of Favata's\cite{Favata2009} paper, where we include the spin contribution upto leading (1.5 PN) order for nonprecessing binaries.
+For PN calculations, this paper is just an extension of Favata's\cite{Favata2009} paper, where we include the spin contribution up to leading (1.5 PN) order for nonprecessing binaries.
 The nonlinear memory contribution near merger is computed using NR data from publicly available SXS catalog, where we process the data to remove the low frequency numerical noise.
@@
 The 3.5PN orbital energy and flux are given in Appendix C of reference \cite{Arun2009}.
-Here we who only upto 1.5 PN terms, the complete expression are show in Appendix \ref{Appendix:C1}.
+Here we who only up to 1.5 PN terms, the complete expression are show in Appendix \ref{Appendix:C1}.
 The 1.5 PN expression for energy is given by
@@
 
-Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, and expanding upto 1.5 PN gives the following expression.
+Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, and expanding up to 1.5 PN gives the following expression.
 
@@
 \end{widetext}
-The results upto 3.5 PN are shows in Appendix \ref{Appendix:C}.
+The results up to 3.5 PN are shows in Appendix \ref{Appendix:C}.
 The equation \ref{eq:2.14} can be used to change the integration with respect to time into integration with respect to $x$ as shown in equation in \ref{eq:2.11}.
@@
 \end{subequations}
-The remaining $h_{lm}$ expressions does not contain spin terms and the complete expressions upto 3PN terms can be directly taken from Ref.\cite{Blanchet_Iyer2008}
+The remaining $h_{lm}$ expressions does not contain spin terms and the complete expressions up to 3PN terms can be directly taken from Ref.\cite{Blanchet_Iyer2008}
  We show the $h_{lm}$ expressions for positive $m$ values, but the results for negative $m$ values can be obtained by the relation, $h_{l, -m} = (-1)^l \bar{h}_{lm}$.

### 2019-03-27 notes/Paper.tex
some corrections

@@
 These $m=0$ modes are much smaller in magnitude and can be masked by numerical noise as well as subtleties of the numerical techniques used to extract the radiation.
-For a detailed discussion of these issues, see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009} \stm{the main reference for this is Reisswig and Pollney, and you should actually read that and include a little more detail here}
+For a detailed discussion of these issues, see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009} \ashok{NR simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by
+	\begin{equation}
+	\psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_{\times}
+	\end{equation}\
+	Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$.
+It has long been known that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}.
+The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation.
+This nonlinearity is potentially caused by the fact that $\psi_4$ is typically extracted at a finite distance from the gravitating source.
+The strain $h$ is only related to $\psi_4$ at an infinite distance and also strictly valid in a particular gauge.
+As a result in NR simulation, the finite distance calculation introduces a systematic error.
+In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.}\stm{the main reference for this is Reisswig and Pollney, and you should actually read that and include a little more detail here}
 \par NR simulations that use the most widely adopted methods directly compute the Weyl curvature scalar, $\Psi_4$, and decompose its values at large radii $R$ into a sum over spin-weighted spherical harmonic modes:
@@
 \section{Memory calculation using NR data}
-NR simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by
-\begin{equation}
-	\psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_{\times}
-\end{equation}\
-Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$.
-It has long been known that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}.
-The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation.
-This nonlinearity is potentially caused by the fact that $\psi_4$ is typically extracted at a finite distance from the gravitating source.
-The strain $h$ is only related to $\psi_4$ at an infinite distance and also strictly valid in a particular gauge.
-As a result in NR simulation, the finite distance calculation introduces a systematic error.
-In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
+
 \par It has been argued in ref.

### 2019-03-29 tests/test_computeBWMResedual.py
added code to look at reseduals

@@ -1,0 +1,115 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+def step(x,x0, amp):
+	if x<=x0:
+		ans=0
+	else:
+		ans=amp
+	return ans 
+
+def s(t, hpmem, mu, phi):
+	dt =t[1]-t[0]
+	int_hp = np.cumsum(hpmem)*dt
+	
+	st = (1.0/2.0)*(1.0+mu)*(int_hp*np.cos(2*phi))
+
+	return st
+
+
+#data location
+file_location ='/home/ashok/gravitational_wave_memory_project/data/SamMassRatio_differentSzSpinOnly/Memory_data/'
+#filename_vec=['0p2', '0p43', '0p600', '0p800', '0p95']
+filename='0p43'
+
+datafile_hMemNR='rMPsi4_AlignedSpin_Sz_'+filename+'_q1dataClean_hMemNR.dat'
+timeNR, hmem, h_mem_plus = np.loadtxt(file_location+datafile_hMemNR, unpack=True)
+
+#Reproduce fig1 from Pshirkov paper
+
+time = np.linspace(0,10)
+hp = np.array([])
+for i in range(len(time)):
+	hp=np.append(hp,step(time[i],5,1))
+
+res = s(time, hp, 0, 0) 
+res_quadfit = np.polyfit(time, res, 2)
+res1d = np.poly1d(res_quadfit)
+res_quadSubtract = res-res1d(time)
+
+plt.plot(time, res1d(time), "r--")
+plt.plot(time, res, "g--")
+plt.plot(time, res_quadSubtract, "b")
+plt.show()
+
+
+####take numerical FFT and plot
+#hpf = np.fft.fft(hp)
+#freq = np.fft.fftfreq(len(hpf),d=time[2]-time[1])
+#plt.plot(freq, hpf)
+#plt.show()
+#################################################
+
+##construction a step function which looks like hmem from SXS data
+maxhmemSXS = max(hmem)
+timeleft = np.linspace(timeNR[0], timeNR[-1], len(timeNR))
+hmemStepleft = np.array([])
+for i in range(len(timeNR)):
+	hmemStepleft=np.append(hmemStepleft,step(timeNR[i],timeNR[-100], maxhmemSXS))
+
[…]

### 2019-04-08 notes/Paper.tex
updated notes

--- 2019-03-27 notes/Paper.tex
@@
 These $m=0$ modes are much smaller in magnitude and can be masked by numerical noise as well as subtleties of the numerical techniques used to extract the radiation.
-For a detailed discussion of these issues, see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009} \stm{the main reference for this is Reisswig and Pollney, and you should actually read that and include a little more detail here}
+For a detailed discussion of these issues, see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009} \ashok{The Christodoulou memory from spinning binaries have been studied by Reisswig and Pollney \cite{Reisswig2011} in numerical simulations, where they carry out fully relativistic binary BH simulations to directly measure the non-linear GW memory through inspiral, merger and ringdown in a gauge invariant way.
+In these studies, the authors have considered the equal mass spinning case, where the memory is contained in the $_{-2}Y_{l0}$ spin-weighted spherical harmonic components of the GW strain, $h_{+}$.
+The authors have found that the largest nonlinear memory is observed in case of merger of the aligned, maximally spinning BHs.
+Another interesting results in their studies is the case of the merger of maximally spinning anti aligned BHs, where they observed the strongest oscillating ringdown signal.}
 \par NR simulations that use the most widely adopted methods directly compute the Weyl curvature scalar, $\Psi_4$, and decompose its values at large radii $R$ into a sum over spin-weighted spherical harmonic modes:
@@
 \psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_{\times}
-\end{equation}\
+\end{equation}
 Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$.
@@
 In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
+
 \par It has been argued in ref.

### 2019-04-18 notes/Paper.tex
Ashok made some corrections in the paper

@@
 The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata \cite{Favata2009}.
-Here we include the contribution from the individual black-hole spins up to 1.5PN \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order.
+Here we include the contribution from the individual black-hole spins at leading order which appears at 1.5PN \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order.
 We also compute the memory contribution during
@@
 The 3.5PN orbital energy and flux are given in Appendix C of reference \cite{Arun2009}.
-Here we who only up to 1.5 PN terms, the complete expression are show in Appendix \ref{Appendix:C1}.
+Here we show only up to 1.5 PN terms, the complete expression are show in Appendix \ref{Appendix:C1}.
 The 1.5 PN expression for energy is given by
@@
 Setting $\iota = 0, \alpha = \pi$ in equation 4.16 and 4.17 gives the result for non-precessing case.
-In that case the spins are aligned along $z$ axis, so we set $\chi_a^x=\chi_s^x=\chi_a^y=\chi_s^y=0$.
+In our case the spins are aligned along $z$ axis, so we set $\chi_a^x=\chi_s^x=\chi_a^y=\chi_s^y=0$.
 We can now write the expression as
 \begin{equation}
-h_{lm} = (-1)^{m+1}\frac{(2M\eta v^2)}{D_L}\sqrt{\frac{16\pi}{5}}e^{\mathit{i}\,m\,\psi}\hat{h}_{lm},
+h_{lm} = (-1)^{m+1}\frac{(2M\eta x)}{D_L}\sqrt{\frac{16\pi}{5}}e^{\mathit{i}\,m\,\psi}\hat{h}_{lm},
 \end{equation}
@@
 A method to remove this unphysical part is given in ref \cite{Reisswig_Pollney2011}.
-If we remove this low frequency part from data such that the physical memory is still preserved, we can get rid of memory contribution due to numerical noise.
+If we remove this low frequency part from the data such that the physical memory is still preserved, we can get rid of memory contribution due to numerical noise.
 Consider the Fourier transform, $\mathcal{F}$, applied to an absolutely integrable function $f(t)$,
@@
 We compare this memory for different mass ratio and different values of dimensionless spin parameter.
-Figure \ref{fig:differentmassratio} show the Christodoulou memory for different mass ratios, which shows that the Christodoulou tends to decrease with higher mass ration.
+Figure \ref{fig:differentmassratio} show the Christodoulou memory for different mass ratios, which shows that the Christodoulou tends to decrease with higher mass ratios.
 The effect of dimensionless spin is shown in figure \ref{fig:differntchivalues}.

### 2019-07-12 scripts/Memory_Mass_Vs_Spin_v2.py
added code to make memory growth plots

@@ -4,6 +4,8 @@ import matplotlib.cm as cm
 import plotsettings
 from matplotlib.font_manager import FontProperties
 from matplotlib import colors, cm
+from matplotlib.ticker import LogLocator
+from matplotlib.colors import LogNorm
 
 def find_nearest_idx(array, value):
     array = np.asarray(array)
@@ -70,13 +72,15 @@ for i in range(len(Spin_array)):
 		Memory_growth_M_vs_Spin[i][j]  = 4.4*(10**(Mass_array[j]-23))*Memory_growth_in_two_weeks(Mass_array[j], Spin_array[i])
 
 fig, ax = plt.subplots()
-im = plt.imshow(Memory_growth_M_vs_Spin , interpolation='bilinear', cmap=cm.RdYlGn, origin='lower', extent=[-0.90, 0.99, 8, 10], vmax=abs(Memory_growth_M_vs_Spin).max(), vmin=-abs(Memory_growth_M_vs_Spin).max())
+im = plt.imshow(Memory_growth_M_vs_Spin, interpolation='bilinear', cmap=cm.RdYlGn, origin='lower', extent=[-0.90, 0.99, 8, 10], norm=LogNorm(vmin=3.261800077256976e-18, vmax=8.031504325972437e-16))
 plt.xlabel('Spin')
 plt.ylabel('Log{(M)')
 plt.clim(abs(Memory_growth_M_vs_Spin).min(),abs(Memory_growth_M_vs_Spin).max())
 plt.colorbar()
 fig.tight_layout()
-plt.show()	
+plt.savefig("/home/ashok/Desktop/gravitational_wave_memory_project/plots/Memory_Spin_vs_Mass.pdf")	
+plt.show()
+
 
 print abs(Memory_growth_M_vs_Spin).min(),abs(Memory_growth_M_vs_Spin).max()
 
@@ -113,7 +117,7 @@ def compute_rms_reseduals(log_SolarMass, Spin):
 	
 	idx = find_nearest_idx(timeNR, t_initial)
 	
-	hmem_two_weeks = hmem[idx:find_nearest_idx(timeNR, 25)]
+	hmem_two_weeks =  4.4*(10**(log_SolarMass-23))*hmem[idx:find_nearest_idx(timeNR, 25)]
 	timeNR_two_weeks = timeNR[idx:find_nearest_idx(timeNR, 25)]
 
[…]

### 2019-07-12 notes/Paper.tex
updated notes

@@
 \end{figure}
+\section{Observing Memory from merger of Super-massive black holes  }
+Nonlinear memory from the merger of super massive black hole binary can be potentially observed in pulsar timing measurement.
+Gravitation wave, when they pass through the space between earth and pulsar effect the time of arrival of the radio pulses at earth.
+The growth in the nonlinear memory signal amplitude from super massive black hole occurs on the time scale of a few days so the effect on timing residuals can be seen if observed over a few weeks period of time.
+It can be seen from equation [Eqn 26] that the memory amplitude depends on the total mass and distance from the source.
+The details regarding the calculation of the timing residuals can be seen in \cite{Pshirkov2010}, where they consider the memory event as a step function which occurs in the middle of an observation time of 10 years.
+The method described in \cite{Pshirkov2010} is general and can be applied to case where the memory jump is not at step function.
+We have seen that the spin can significantly change memory amplitude as a result it can change the timing residuals.
 
-\subsection{CONCLUSIONS}
+\begin{figure}
+	\includegraphics[width=4.0in]{../plots/Memory_Spin_vs_Mass.pdf}
+	\includegraphics[width=4.0in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
+	\caption{The top figure shows the Christodolou memory growth during two weeks of observation time as a function of Mass and Spin magnitude $\chi$.
+The here in M time the mass of Sun.
+The bottom plot show a the rms timing residuals during two week period of time.}
+	\label{fig:MemoryMassVsSpin}
+\end{figure}
+\section{CONCLUSIONS}
 Observation of Christodoulou memory would be an important test of general relativity in strong field regime.
@@
 Hence it is important to consider the spin effect while modeling the signals.
+The important thing for pulsar timing measurements would be the amount of memory changes over a single two weeks period of time which is shown the figure \ref{fig:MemoryMassVsSpin} as function of total mass of the binary and the spin magnitude.
+For a given mass and over the allowed spin range the net change in the memory amplitude can be as big as a factor of 5 which would change the SNR, or the expected event rate within a given volume, by a factor of $5^{3}=125$
 

### 2019-07-15 notes/Paper.tex
fixed bibtex

@@
 The growth in the nonlinear memory signal amplitude from super massive black hole occurs on the time scale of a few days so the effect on timing residuals can be seen if observed over a few weeks period of time.
-It can be seen from equation [Eqn 26] that the memory amplitude depends on the total mass and distance from the source.
+A method for observing Nonlinear memory signal from super massive black hole mergers using pulsar timing has been proposed in ref\cite{Pshirkov2010}.
+The authors have shown that the present day pulsar timing precision allows one to detect BWM due to  SMBH merger at distance upto 1 Gpc (for case of equal mass $10^{8}$ solar mass).
 The details regarding the calculation of the timing residuals can be seen in \cite{Pshirkov2010}, where they consider the memory event as a step function which occurs in the middle of an observation time of 10 years.
 The method described in \cite{Pshirkov2010} is general and can be applied to case where the memory jump is not at step function.
-We have seen that the spin can significantly change memory amplitude as a result it can change the timing residuals.
+In our analysis we explore how the nonlinear memory grows during two weeks interval to time for different ranges of mass and spin parameter values.
+We also calculate the timing residuals as done in \cite{Pshirkov2010} and explore how it changes as a function of total mass and spin of black holes as shown in Fig.
+\ref{fig:MemoryMassVsSpin}.
 
@@
 It is highly likely that super massive black holes have significant spin due to accretion of matter from surrounding.
-We have shown in this paper that spin contribution can significantly change the Christodoulou memory.
+We have seen that the spin can significantly change memory amplitude as a result it can change the timing residuals.
 Hence it is important to consider the spin effect while modeling the signals.

### 2019-08-19 scripts/MemoryGrowthIntenYears.py
updated scripts

@@ -1,0 +1,126 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+from matplotlib import colors, cm
+from matplotlib.ticker import LogLocator
+from matplotlib.colors import LogNorm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+
+def find_nearest_idx(array, value):
+    array = np.asarray(array)
+    idx = (np.abs(array - value)).argmin()
+    return idx
+                                         
+# defining intial time for given solar mass for a given fixed time interval of two weeks
+def time_initial(x, time_final):
+	return time_final - 2.6*pow(10, 11-x)
+
+#function that integrate the memory signal
+def s(t, hpmem, mu, phi):
+	dt =t[1]-t[0]
+	int_hp = np.cumsum(hpmem)*dt
+	
+	st = (1.0/2.0)*(1.0+mu)*(int_hp*np.cos(2*phi))
+
+	return st
+
+
+# Function that compute memory growth during two weeks interval of time for a given mass
+
+def Memory_growth_in_two_weeks(log_SolarMass, Spin):
+		
+	if Spin < 0:
+		file_location_hmem ='/home/ashok/gravitational_wave_memory_project/data/SXSdata/Spinning_binary_with_SpinAntialigned_27Dec/Memory_data/'
+		filename = 'm0p'+str(Spin)[3:5]
+		
+		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
+		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
+	
+		datafile_hNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hNR.dat'
+		timeNR, hp, hc = np.loadtxt(file_location_hmem+datafile_hNR, unpack=True)
+	
+
+		
+	if Spin == 0.0:
[…]

### 2019-08-19 scripts/Memory_Mass_Vs_Spin_tenYear.py

@@ -1,0 +1,343 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+from matplotlib import colors, cm
+from matplotlib.ticker import LogLocator
+from matplotlib.colors import LogNorm
+
+numer_of_observation_days=5.0*365.0	
+
+def find_nearest_idx(array, value):
+    array = np.asarray(array)
+    idx = (np.abs(array - value)).argmin()
+    return idx
+                                         
+# defining intial time for given solar mass for a given fixed time interval of two weeks
+def time_initial(log_solarMass, time_final, numer_of_observation_days):
+	return time_final - 1.76*pow(10, 10-log_solarMass)*numer_of_observation_days
+
+#function that integrate the memory signal
+def s(t, hpmem, mu, phi):
+	dt =t[1]-t[0]
+	int_hp = np.cumsum(hpmem)*dt
+	
+	st = (1.0/2.0)*(1.0+mu)*(int_hp*np.cos(2*phi))
+
+	return st
+
+
+# Function that compute memory growth during two weeks interval of time for a given mass
+
+def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
+		
+	if Spin < 0:
+		file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAntialigned_27Dec/Memory_data/'
+		filename = 'm0p'+str(Spin)[3:5]
+		
+		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
+		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
+	
+		datafile_PN_mem='PN_data_'+filename+'.txt'
+		timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
+	
+
+		
+	if Spin == 0.0:
+		file_location_hmem ='../data/NonSpinning_differentMassRatio/Memory_data/'
+
[…]

### 2019-08-19 scripts/ResdualGrowthIntenyear_v2.py

@@ -220,7 +220,7 @@ plt.plot(time_PN, hp_mem_PN, 'r:', label = "PN memory")
 plt.plot(timeNR, hmem,'k--', label = 'NR memory')
 
 #plt.xlim(-5000, 25000)
-plt.ylim(0.0, 0.1)
+plt.ylim(0.0, 0.2)
 plt.xlabel(r'$t/M$')
 plt.ylabel(r'$(R/M)\,h^{(mem)}_{+}$')
 plt.legend(loc=2)
@@ -241,7 +241,7 @@ ax2.plot(time_two_weeks, hmem_two_weeks)
 plt.xlim(22500, 22530)
 plt.ylim(0.0379, 0.068)
 '''
-plt.savefig("../plots/MemoryGrowthtwoweeks.pdf")
+#plt.savefig("../plots/MemoryGrowthtwoweeks.pdf")
 plt.show()
 
 
@@ -402,7 +402,7 @@ for i in Spin_array:
 		timeNR_two_weeks = timeNR_two_weeks - timeNR_two_weeks[0] +  k*shift + l*shiftSpin
 		plt.plot(timeNR_two_weeks, res_two_weeks,'k--')
 		plt.plot(timeNR_two_weeks, res_quad_fit_two_weeks ,'r--')
-		plt.plot(timeNR_two_weeks, res_quadSubtract_two_weeks , label='$M=10 ^{'+ str(round(j,1))+'}M_\odot\t\, \, S ='+str(round(i, 2))+'$')
+		plt.plot(timeNR_two_weeks, res_quadSubtract_two_weeks , label='M=10**(' + str(round(j,1))+')\, S ='+ str(round(i, 2)))
 		#time_last = timeNR_two_weeks[-1]+25
 		k+=1		
 	l+=1
@@ -416,6 +416,6 @@ plt.xlabel(r'$time \, (days)$')
 plt.ylabel(r'$Residuals \, (fs)$')
 plt.legend(loc=2)
 fontP.set_size('12.')
-plt.savefig("../plots/ResedualGrowthtwoweeks.pdf")
+#plt.savefig("../plots/ResedualGrowthtwoweeks.pdf")
 plt.show()
 

### 2019-08-19 scripts/plotsettings.py

@@ -11,7 +11,7 @@ rc('axes', labelsize=14) #24
 rc("axes", linewidth=0.5) #2)
 rc('xtick', labelsize=18)
 rc('ytick', labelsize=18)
-rc('legend', fontsize=16) #16
+rc('legend', fontsize=10) #16
 rc('xtick.major', pad=8) #8)
 rc('ytick.major', pad=8) #8)
 rc('xtick.major', size=13) #8)

### 2019-08-19 notes/Paper.tex
updated notes

@@
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the orbital separation.
-The plots above show the memory contribution to $H_+$  as a function of $x$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $\eta=1/4 $ .
+The plots above show the memory contribution to $H_+$  as a function of $x$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at $\Theta$ = 0 points along the binary orbital angular momentum) for $\eta=1/4 $ .
 The top left plot shows $x H_+$ as a function of parameter x which is a function of time.
@@
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
-The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4$.}
+The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at $\Theta$ = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4$.}
 \end{figure}
@@
 \begin{figure}
- 	\includegraphics[width=4.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
+ 	\includegraphics[width=3.5in]{../plots/MemoryPlot_nonSpining/q7.pdf}
  	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 \begin{figure}
-	\includegraphics[width=4.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
-	\includegraphics[width=4.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 In our analysis we explore how the nonlinear memory grows during two weeks interval to time for different ranges of mass and spin parameter values.
-We also calculate the timing residuals as done in \cite{Pshirkov2010} and explore how it changes as a function of total mass and spin of black holes as shown in Fig.
-\ref{fig:MemoryMassVsSpin}.
-
+We also calculate the timing residuals as done in \cite{Pshirkov2010} and explore how it changes as a function of total mass and spin of black holes as shown in Fig[\ref{fig:MemoryMassVsSpin}].
+\par 
+Future observations with increased precision would be able to look at the memory growth on even shorter time of few weeks.
+The memory growth on the time scale of two weeks does not look like a step function, instead it's shape depends upon the spins of the merging black holes.
+In Fig[\ref{fig:resedualGrowth}].we show the memory growth interval in two weeks interval of time with solid lines for different range of spin and total mass of the binary.
+The dashed line show the complete memory growth in growth in geometrical units.
+It can be seen the as the total mass of binary becomes larger the two weeks interval of time shrinks in geometrical units.
+We have chosen this two weeks interval of time such that the change in memory during this time interval is maximum.
+For the pulsar timing observation we compute the residuals during this two week interval of time following the method described in \cite{Pshirkov2010}.
+We show a figure similar to
 \begin{figure}
-	\includegraphics[width=4.0in]{../plots/Memory_Spin_vs_Mass.pdf}
-	\includegraphics[width=4.0in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryGrowthtwoweeks.pdf}
+	\includegraphics[width=3.5in]{../plots/ResedualGrowthtwoweeks.pdf}
+	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
+The dashed line shows the complete memory growth.
+The dashed red line show the memory growth computed using PN}
+	\label{fig:resedualGrowth}
+\end{figure}
+\begin{figure}
+	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
 	\caption{The top figure shows the Christodolou memory growth during two weeks of observation time as a function of Mass and Spin magnitude $\chi$.
@@
 The bottom plot show a the rms timing residuals during two week period of time.}
+	\label{fig:MemoryMassVsSpin}
+\end{figure}
+\begin{figure}
+	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass_tenyears.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass_tenyears.pdf}
+	\caption{The top figure shows the Christodolou memory growth during ten years of observation time as a function of Mass and Spin magnitude $\chi$.
+The here in M time the mass of Sun.
+The bottom plot show a the rms timing residuals during ten years interval of time.}
 	\label{fig:MemoryMassVsSpin}

### 2019-08-22 scripts/Memory_Mass_Vs_Spin_tenYear.py
updated scripts

@@ -7,7 +7,7 @@ from matplotlib import colors, cm
 from matplotlib.ticker import LogLocator
 from matplotlib.colors import LogNorm
 
-numer_of_observation_days=5.0*365.0	
+numer_of_observation_days=50.0	
 
 def find_nearest_idx(array, value):
     array = np.asarray(array)
@@ -39,7 +39,7 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
 	
-		datafile_PN_mem='PN_data_'+filename+'.txt'
+		datafile_PN_mem='ResampledPN_data_'+filename+'.txt'
 		timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
 	
 
@@ -50,7 +50,7 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		filename_hmem = 'rMPsi4_noSpin_q1dataClean_hMemNR.dat'
 
 		datafile_hMemNR = filename_hmem
-		datafile_PN_mem='PN_data_0.txt'
+		datafile_PN_mem='ResampledPN_data_0.txt'
                 
                 timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
                 timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
@@ -62,10 +62,18 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
 
-		datafile_PN_mem='PN_data_'+filename+'.txt'
+		datafile_PN_mem='ResampledPN_data_'+filename+'.txt'
 		timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
 
[…]

### 2019-08-22 scripts/Memory_Mass_Vs_Spin_v4.py

@@ -7,6 +7,7 @@ from matplotlib import colors, cm
 from matplotlib.ticker import LogLocator
 from matplotlib.colors import LogNorm
 
+numer_of_observation_days=14.0	
 
 def find_nearest_idx(array, value):
     array = np.asarray(array)
@@ -29,7 +30,7 @@ def s(t, hpmem, mu, phi):
 
 # Function that compute memory growth during two weeks interval of time for a given mass
 
-def Memory_growth_in_two_weeks(log_SolarMass, Spin):
+def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		
 	if Spin < 0:
 		file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAntialigned_27Dec/Memory_data/'
@@ -38,22 +39,16 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin):
 		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
 	
-		datafile_hNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hNR.dat'
-		timeNR, hp, hc = np.loadtxt(file_location_hmem+datafile_hNR, unpack=True)
-	
 
 		
 	if Spin == 0.0:
 		file_location_hmem ='../data/NonSpinning_differentMassRatio/Memory_data/'
 
 		filename_hmem = 'rMPsi4_noSpin_q1dataClean_hMemNR.dat'
-		filename_h22 = 'rMPsi4_noSpin_q1dataClean_hNR.dat'
 
 		datafile_hMemNR = filename_hmem
-		datafile_hNR = filename_h22
 		
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
-		timeNR, hp, hc = np.loadtxt(file_location_hmem+datafile_hNR, unpack=True)	
 
 	if Spin > 0:
 		file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAligned_27Dec/Memory_data/'
[…]

### 2019-08-22 scripts/Resample_PN_data_t0LowResl.py

@@ -1,0 +1,143 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+def find_nearest_idx(array, value):
+    array = np.asarray(array)
+    idx = (np.abs(array - value)).argmin()
+    return idx
+
+def Resample_ten_years_memory_data(Spin):
+		
+    if Spin < 0:
+            file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAntialigned_27Dec/Memory_data/'
+            filename = 'm0p'+str(Spin)[3:5]
+                    
+            datafile_PN_mem='PN_data_'+filename+'.txt'
+            timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
+            
+    if Spin == 0.0:
+            file_location_hmem ='../data/NonSpinning_differentMassRatio/Memory_data/'
+
+            datafile_PN_mem='PN_data_0.txt'
+            
+            timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
+
+    if Spin > 0:
+            file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAligned_27Dec/Memory_data/'
+            filename = '0p'+str(Spin)[2:4]
+    
+            datafile_PN_mem='PN_data_'+filename+'.txt'
+            timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
+
+
+    return timePN, hmem_PN, file_location_hmem, datafile_PN_mem
+
+Spin=-0.941
+#Spin_array = np.array([0.99, 0.801, 0.601, 0.201, 0.0, -0.201, -0.431, -0.601, -0.801, -0.941])
+
+
+time = Resample_ten_years_memory_data(Spin)[0]
+hmem = Resample_ten_years_memory_data(Spin)[1]
+
+t0=time[0]
+tf_PN=time[-1]
+
+num_of_subintervl = 50.0
+len_each_intrvl = (tf_PN-t0)/num_of_subintervl
+
+
+dt=time[2]-time[1]
+
+time_intr = np.array([])
+
+ti=t0
+tf=t0+len_each_intrvl
+i=1
+
+while tf< -4000.0:
+    dx_sample = len_each_intrvl/(10*i)
+    print i, dx_sample, tf
[…]

### 2019-08-22 scripts/ResdualGrowthIntenyear.py

@@ -169,7 +169,7 @@ t_fPN = 550.0
 
 Spin_array = np.array([-0.941,0.0, 0.99])
 
-Mass_array = np.array([8.0, 9.0,10.0])
+Mass_array = np.array([8.0, 9.0,10.0, 11.0 ,12.0])
 
 shift = 2500
 shiftSpin=10000
@@ -361,7 +361,7 @@ def compute_rms_reseduals(log_SolarMass, Spin, numer_of_observation_days):
 
 		
 		
-	hmem_two_weeks =  4.6*(10**(log_SolarMass-23))*hmem_two_weeks_ip1
+	hmem_two_weeks =  4.9*(10**(log_SolarMass-23))*hmem_two_weeks_ip1
 	#Convert to femto seconds
 	hmem_two_weeks=pow(10,15)*hmem_two_weeks
 	

### 2019-08-22 scripts/ResdualGrowthIntenyear_v2.py

@@ -220,7 +220,7 @@ plt.plot(time_PN, hp_mem_PN, 'r:', label = "PN memory")
 plt.plot(timeNR, hmem,'k--', label = 'NR memory')
 
 #plt.xlim(-5000, 25000)
-plt.ylim(0.0, 0.2)
+plt.ylim(0.0, 0.1)
 plt.xlabel(r'$t/M$')
 plt.ylabel(r'$(R/M)\,h^{(mem)}_{+}$')
 plt.legend(loc=2)
@@ -241,7 +241,7 @@ ax2.plot(time_two_weeks, hmem_two_weeks)
 plt.xlim(22500, 22530)
 plt.ylim(0.0379, 0.068)
 '''
-#plt.savefig("../plots/MemoryGrowthtwoweeks.pdf")
+plt.savefig("../plots/MemoryGrowthtwoweeks.pdf")
 plt.show()
 
 
@@ -416,6 +416,6 @@ plt.xlabel(r'$time \, (days)$')
 plt.ylabel(r'$Residuals \, (fs)$')
 plt.legend(loc=2)
 fontP.set_size('12.')
-#plt.savefig("../plots/ResedualGrowthtwoweeks.pdf")
+plt.savefig("../plots/ResedualGrowthtwoweeks.pdf")
 plt.show()
 

### 2019-08-22 scripts/plotsettings.py

@@ -9,8 +9,8 @@ rc("lines", markeredgewidth=0.6)
 rc("lines", linewidth=2)
 rc('axes', labelsize=14) #24
 rc("axes", linewidth=0.5) #2)
-rc('xtick', labelsize=18)
-rc('ytick', labelsize=18)
+rc('xtick', labelsize=15)
+rc('ytick', labelsize=15)
 rc('legend', fontsize=10) #16
 rc('xtick.major', pad=8) #8)
 rc('ytick.major', pad=8) #8)

### 2019-08-23 notes/Paper.tex
Updated notes

@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
-	\includegraphics[width=3.5in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
+	\includegraphics[width=3.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
+	\includegraphics[width=3.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 Gravitation wave, when they pass through the space between earth and pulsar effect the time of arrival of the radio pulses at earth.
+The modulation in the observed frequency is given by
+\begin{equation}
+\frac{\Delta \nu}{\nu_{0}} = \frac{1}{2c}\int_{0}^{D}d\lambda \left(e^{i}e^{j}\frac{\partial h_{ij}}{\partial t}\right) \Bigg\vert_{path}
+\end{equation}
+
+The pulsar timing measurement primarily measures the difference between the actual pulse arrival time and the time predicted from the spin-down model of a pulsar, which is called the timing residuals.
+The timing residuals $s(t)$, accumulated during a time interval of length due to passing gravitational waves is given by the following expression
+\begin{equation}
+	s(t) = \frac{1}{2}(1+\mu)\left[\left(\int_{0}^{t}d\tau \, h_{+}(\tau)\right)\cos 2\phi + \left(\int_{0}^{t}d\tau \, h_{\times}(\tau)\right)\sin 2\phi\right]
+\end{equation}
+There are two unknown quantities associated with the timing residuals, the pulsar periods and it's spin down rate. We subtract these two quantities from the expression $s(t)$,
+
 The growth in the nonlinear memory signal amplitude from super massive black hole occurs on the time scale of a few days so the effect on timing residuals can be seen if observed over a few weeks period of time.
@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/MemoryGrowthtwoweeks.pdf}
-	\includegraphics[width=3.5in]{../plots/ResedualGrowthtwoweeks.pdf}
+	\includegraphics[width=3.0in]{../plots/MemoryGrowthtwoweeks.pdf}
+	\includegraphics[width=3.0in]{../plots/ResedualGrowthtwoweeks.pdf}
 	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass.pdf}
-	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.0in]{../plots/Memory_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.0in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
 	\caption{The top figure shows the Christodolou memory growth during two weeks of observation time as a function of Mass and Spin magnitude $\chi$.
@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass_tenyears.pdf}
-	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass_tenyears.pdf}
+	\includegraphics[width=3.0in]{../plots/Memory_Spin_vs_Mass_tenyears.pdf}
+	\includegraphics[width=3.0in]{../plots/MemoryRes_Spin_vs_Mass_tenyears.pdf}
 	\caption{The top figure shows the Christodolou memory growth during ten years of observation time as a function of Mass and Spin magnitude $\chi$.

### 2019-08-24 scripts/Memory_Mass_Vs_Spin_tenYear.py
updated scripts

@@ -7,7 +7,7 @@ from matplotlib import colors, cm
 from matplotlib.ticker import LogLocator
 from matplotlib.colors import LogNorm
 
-numer_of_observation_days=50.0	
+numer_of_observation_days=14.0	
 
 def find_nearest_idx(array, value):
     array = np.asarray(array)
@@ -68,18 +68,46 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 	hmem=17.0*hmem
         
         # Do interpolation on NR data to increase the resolution. Necessory because the loaded data could not handle 10^12 solar mass two weeks time interval.
-        dt_NR_intrp=(timeNR[2]-timeNR[1])/4.0
+        hmem_orgnl = hmem
+        timeNR_orgnl = timeNR
+
+        dt_NR_intrp=(timeNR[2]-timeNR[1])/100.0
         timeNR_intrp = np.arange(timeNR[0], timeNR[-1], dt_NR_intrp)
 
         hmem_intrp = np.interp(timeNR_intrp, timeNR, hmem)
+
+        #plt.plot(timeNR_intrp, hmem_intrp)
+        #plt.plot(timeNR, hmem, 'r--')
+        #plt.show()
+
         timeNR = timeNR_intrp
         hmem = hmem_intrp
 
+        #Set PN memory to same resolution as NR after interpolation 
+        '''
+        hmem_PN1 = hmem_PN[:-len(hmem_orgnl)]
+        timePN1 = timePN[:-len(hmem_orgnl)]
+
+        hmem_PN2 = hmem_PN[-len(hmem_orgnl):]
+        timePN2 = timePN[-len(hmem_orgnl):]
+
+        timePN2_intrp = np.arange(timePN2[0], timePN2[-1], dt_NR_intrp)
+        hmem_PN2_intrp = np.interp(timePN2_intrp, timePN2, hmem_PN2)
+
+        hmem_PN = np.append(hmem_PN1, hmem_PN2_intrp)
+        time_PN = np.append(timePN1, timePN2_intrp)
+
+        #plt.plot(timePN, hmem_PN)
+        #plt.plot(time_PNx, hmem_PNx, '--')
+        #plt.show()
+
+        '''
+
+      
[…]

### 2019-08-24 scripts/Resample_PN_data_t0LowResl.py

@@ -42,7 +42,7 @@ hmem = Resample_ten_years_memory_data(Spin)[1]
 t0=time[0]
 tf_PN=time[-1]
 
-num_of_subintervl = 50.0
+num_of_subintervl = 40.0
 len_each_intrvl = (tf_PN-t0)/num_of_subintervl
 
 
@@ -54,13 +54,14 @@ ti=t0
 tf=t0+len_each_intrvl
 i=1
 
-while tf< -4000.0:
+while tf< -9000.0:
     dx_sample = len_each_intrvl/(10*i)
-    print i, dx_sample, tf
     time_updated=np.arange(ti, tf, dx_sample)
     time_intr = np.append(time_intr, time_updated) 
     ti=tf
     tf=ti+len_each_intrvl
+    print i, dx_sample, tf
+
 
     i+=1
     

### 2019-08-24 scripts/ResdualGrowthIntwoWeeks_v3.py

@@ -11,6 +11,8 @@ from matplotlib.font_manager import FontProperties
 import PostNewtonianMemoryFnc as PNM
 from mpl_toolkits.axes_grid.inset_locator import (inset_axes, InsetPosition, mark_inset)
 
+numer_of_observation_days=2000.0	
+
 
 def find_nearest_idx(array, value):
     array = np.asarray(array)
@@ -32,7 +34,7 @@ def s(t, hpmem, mu, phi):
 
 # Function that compute memory growth during two weeks interval of time for a given mass
 
-def Memory_growth_in_two_weeks(log_SolarMass, Spin):
+def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		
 	if Spin < 0:
 		file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAntialigned_27Dec/Memory_data/'
@@ -41,22 +43,21 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin):
 		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
 	
-		datafile_hNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hNR.dat'
-		timeNR, hp, hc = np.loadtxt(file_location_hmem+datafile_hNR, unpack=True)
-	
+		datafile_PN_mem='ResampledPN_data_'+filename+'.txt'
+		timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
 
 		
 	if Spin == 0.0:
 		file_location_hmem ='../data/NonSpinning_differentMassRatio/Memory_data/'
 
 		filename_hmem = 'rMPsi4_noSpin_q1dataClean_hMemNR.dat'
-		filename_h22 = 'rMPsi4_noSpin_q1dataClean_hNR.dat'
-
-		datafile_hMemNR = filename_hmem
-		datafile_hNR = filename_h22
 		
+                datafile_hMemNR = filename_hmem
[…]

### 2019-08-25 scripts/ResdualGrowthIntwoWeeks_v3.py
updated script

@@ -3,15 +3,8 @@ import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 import plotsettings
 from matplotlib.font_manager import FontProperties
-from matplotlib import colors, cm
-from matplotlib.ticker import LogLocator
-from matplotlib.colors import LogNorm
-import plotsettings
-from matplotlib.font_manager import FontProperties
-import PostNewtonianMemoryFnc as PNM
-from mpl_toolkits.axes_grid.inset_locator import (inset_axes, InsetPosition, mark_inset)
 
-numer_of_observation_days=2000.0	
+numer_of_observation_days=14.0	
 
 
 def find_nearest_idx(array, value):
@@ -71,13 +64,40 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 
 	hmem=17.0*hmem
 
-        #Extend the time period when the memory settlles to make it look like a step function
+    #Extend the time period when the memory settlles to make it look like a step function
    
         timeNR_up=timeNR- timeNR[0]+timeNR[-1]
         timeNR=np.append(timeNR,timeNR_up)
 
         hmem_up=np.full(len(timeNR_up), max(hmem))
         hmem=np.append(hmem, hmem_up)
+
+        #Length of NR data
+        len_NR_data = len(timeNR)
+
+        # Create a comple memory waveform
+        idx_NRpeak = find_nearest_idx(timeNR, max(hmem)*0.8)
+        
+        #Shifting the Post-Newtonian part to where it blows up
+        
+        hp_mem_PN = hmem_PN 
+        hp_mem_PN_max_idx = np.argmax(hp_mem_PN)
+        hp_mem_PN = hp_mem_PN[:hp_mem_PN_max_idx]
+        time_PN = timePN[:hp_mem_PN_max_idx]
+                
+        time_PN=time_PN-time_PN[-1]
+        time_PN=time_PN+timeNR[idx_NRpeak]
+
+        #Find index of ti in NR on PN 
+        idx_ti = find_nearest_idx(time_PN, timeNR[0])
+        
[…]

### 2019-08-25 notes/Paper.tex
updated plots

@@
 \begin{figure}
-	\includegraphics[width=3.0in]{../plots/MemoryGrowthtwoweeks.pdf}
-	\includegraphics[width=3.0in]{../plots/ResedualGrowthtwoweeks.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryGrowthtwoweeks.pdf}
+	\includegraphics[width=3.5in]{../plots/ResedualGrowthtwoweeks.pdf}
 	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
@@
 \begin{figure}
-	\includegraphics[width=3.0in]{../plots/Memory_Spin_vs_Mass.pdf}
-	\includegraphics[width=3.0in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
 	\caption{The top figure shows the Christodolou memory growth during two weeks of observation time as a function of Mass and Spin magnitude $\chi$.
@@
 \begin{figure}
-	\includegraphics[width=3.0in]{../plots/Memory_Spin_vs_Mass_tenyears.pdf}
-	\includegraphics[width=3.0in]{../plots/MemoryRes_Spin_vs_Mass_tenyears.pdf}
+	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass_tenyears.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass_tenyears.pdf}
 	\caption{The top figure shows the Christodolou memory growth during ten years of observation time as a function of Mass and Spin magnitude $\chi$.

### 2019-09-30 notes/Paper.tex
Updated plots and notes

@@
 
-\title{Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers}
+\title{Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers and it's implication on Observing Gravitational Wave Bursts With Memory in Pulsar Timing Array}
 \author{Ashok Choudhary}\email{aschoudhary@mix.wvu.edu}
@@
 This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed.
-While the memory contribution during the early inspiral is well described by the post-Newtonian (PN) approximation, the contribution at merger must be obtained using numerical relativity simulations.
+Merger of supermassive black holes is a powerful source of gravitational wave with Chritodoulou  memory effect.
+This effect can be potentially seen in pulsar timing observations.
+Initiatives to observe Burst with memory signal have assumed the signal to be a step function of a given amplitude\cite{Pshirkov2010}, which is valid in  situations like supernova explosion.
+But events like merger of supermassive black holes happens on the time scale of few days depending upon the mass.
+With future pulsar timing observation, it would be possible to see the change in Chritodoulou memory over two weeks interval of time.
+We investigate how the effect of Spin changes the timing residuals over a two weeks period of interval.
+Since Chritodoulou memory is an accumulated effect during the entire history of the source and while memory contribution during the early inspiral is well described by the post-Newtonian (PN) approximation, the contribution at merger must be obtained using numerical relativity simulations.
 The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata \cite{Favata2009}.
@@
 	applying that method using publicly available numerical relativity simulations of spinning binaries.
+We Finally use this model to compute the timing residuals for a memory growth during two weeks interval of time following the method described in \cite{Pshirkov2010}
 \end{abstract}
@@
 We show a figure similar to
+\begin{widetext}
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/MemoryGrowthtwoweeks.pdf}
-	\includegraphics[width=3.5in]{../plots/ResedualGrowthtwoweeks.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksGeoUnits.pdf}
 	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
@@
 \end{figure}
+
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass.pdf}
-	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow8MSun.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow10MSun.pdf}
+	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
+The dashed line shows the complete memory growth.
+The dashed red line show the memory growth computed using PN}
+	\label{fig:resedualGrowth}
+\end{figure} 
+\end{widetext}
+
+\begin{widetext}
+\begin{figure}
+	\includegraphics[width=7.5in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksQuadFitInDays.pdf}
+	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
+The dashed line shows the complete memory growth.
+The dashed red line show the memory growth computed using PN}
+	\label{fig:resedualG}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=7.5in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays.pdf}
+	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
+The dashed line shows the complete memory growth.
+The dashed red line show the memory growth computed using PN}
+	\label{fig:resedualGrowth}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryResSpinvsMass14Days.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryResSpinvsMass1825Days.pdf}
 	\caption{The top figure shows the Christodolou memory growth during two weeks of observation time as a function of Mass and Spin magnitude $\chi$.
@@
 \end{figure}
-\begin{figure}
-	\includegraphics[width=3.5in]{../plots/Memory_Spin_vs_Mass_tenyears.pdf}
-	\includegraphics[width=3.5in]{../plots/MemoryRes_Spin_vs_Mass_tenyears.pdf}
-	\caption{The top figure shows the Christodolou memory growth during ten years of observation time as a function of Mass and Spin magnitude $\chi$.
-The here in M time the mass of Sun.
-The bottom plot show a the rms timing residuals during ten years interval of time.}
-	\label{fig:MemoryMassVsSpin}
-\end{figure}
+\end{widetext}
 \section{CONCLUSIONS}

### 2019-09-30 scripts/CleanSXSdata.py
Updated scripts

@@ -16,7 +16,7 @@ filename_list_nonspinning = ["rMPsi4_noSpin_q1data"]#, "rMPsi4_noSpin_q1p5data",
 for filename in filename_list_nonspinning:
 
 	#filename = "rMPsi4_noSpin_q1data"
-	timeNR, psi_plus, psi_cross = np.loadtxt("/home/ashok/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/"+filename+".txt", unpack=True)
+	timeNR, psi_plus, psi_cross = np.loadtxt("../data/NonSpinning_differentMassRatio/"+filename+".txt", unpack=True)
 
 	plt.plot(timeNR, psi_plus)
 	plt.plot(timeNR, psi_cross)
@@ -43,7 +43,7 @@ for filename in filename_list_nonspinning:
 	plt.plot(timeNR,instfreq)
 	plt.show()
 
-	f = open("/home/ashok/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/"+filename+"Clean.txt","w") 
+	f = open("../data/NonSpinning_differentMassRatio/"+filename+"Clean.txt","w") 
 
 	for i in range(len(timeNR)):
 		f.write("%E %E %E\n" % (timeNR[i], psi_plus[i], psi_cross[i])) 

### 2019-09-30 scripts/MakeReseddualPlot_from_MathematicaData.py

@@ -1,0 +1,15 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+file_location_hmem_Ingrl ='../data/dataForPaper/MemoryGrowthInDaysIngral/'
+datafile_hMem_Ingrl = "MemoryGrowthIngrlSpinm0p9InDays1825LogMass9p81818.txt"
+
+time_InDays, hMem_Ingrl = np.loadtxt(file_location_hmem_Ingrl+datafile_hMem_Ingrl, unpack=True)
+
+plt.plot(time_InDays, hMem_Ingrl)
+plt.show()
+
+

### 2019-09-30 scripts/Memory_Mass_Vs_Spin_v4.py

@@ -39,16 +39,21 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
 	
+		datafile_PN_mem='ResampledPN_data_'+filename+'.txt'
+		timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
 
 		
 	if Spin == 0.0:
 		file_location_hmem ='../data/NonSpinning_differentMassRatio/Memory_data/'
 
 		filename_hmem = 'rMPsi4_noSpin_q1dataClean_hMemNR.dat'
-
-		datafile_hMemNR = filename_hmem
 		
+                datafile_hMemNR = filename_hmem
+    		datafile_PN_mem='ResampledPN_data_0.txt'
+	
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
+                timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
+
 
 	if Spin > 0:
 		file_location_hmem ='../data/SXSdata/Spinning_binary_with_SpinAligned_27Dec/Memory_data/'
@@ -57,22 +62,86 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 		datafile_hMemNR='rMPsi4_Sz1_'+filename+'_Sz2_'+filename+'_q1p5dataN4Clean_hMemNR.dat'
 		timeNR, hmem, h_mem_plus = np.loadtxt(file_location_hmem+datafile_hMemNR, unpack=True)
 
+		datafile_PN_mem='ResampledPN_data_'+filename+'.txt'
+		timePN, hmem_PN = np.loadtxt(file_location_hmem+datafile_PN_mem, unpack=True)
 
-	hmem=17.0*hmem	
+	hmem=17.0*hmem
+        #Increase the resolution which works for 10^12 Msun data
+        #'''
+
+        if log_SolarMass > 10.0:
+
+            dt_NR_intrp=(timeNR[2]-timeNR[1])/10.0
[…]

### 2019-09-30 scripts/PlotMathematicaHmemIntgrldataIndays.py

@@ -1,0 +1,141 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+
+#Load saved memory data from mathematica in geometric units
+
+file_location_hmemGeoUnits = "/home/aschoudhary/gravitational_wave_memory_project/data/dataForPaper/MemoryGrowthInDaysIngral/"
+		
+
+# Load two weeks data for different mass values
+
+filename_Spin0p99_10Pow8MSun = 'MemoryGrowthIngrlSpin0p99InDays14LogMass8p.txt'
+filename_Spin0p0_10Pow8MSun = 'MemoryGrowthIngrlSpin0p0InDays14LogMass8p.txt'
+filename_Spinm0p94_10Pow8MSun = 'MemoryGrowthIngrlSpinm0p94InDays14LogMass8p.txt'
+
+filename_Spin0p99_10Pow10MSun = 'MemoryGrowthIngrlSpin0p99InDays14LogMass10p.txt'
+filename_Spin0p0_10Pow10MSun = 'MemoryGrowthIngrlSpin0p0InDays14LogMass10p.txt'
+filename_Spinm0p94_10Pow10MSun = 'MemoryGrowthIngrlSpinm0p94InDays14LogMass10p.txt'
+
+time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun = np.loadtxt(file_location_hmemGeoUnits+filename_Spin0p99_10Pow8MSun, unpack=True)
+time_Spin0p0_10Pow8MSun, hmem_Spin0p0_10Pow8MSun = np.loadtxt(file_location_hmemGeoUnits+filename_Spin0p0_10Pow8MSun, unpack=True)
+time_Spinm0p94_10Pow8MSun, hmem_Spinm0p94_10Pow8MSun = np.loadtxt(file_location_hmemGeoUnits+filename_Spinm0p94_10Pow8MSun, unpack=True)
+
+time_Spin0p99_10Pow10MSun, hmem_Spin0p99_10Pow10MSun = np.loadtxt(file_location_hmemGeoUnits+filename_Spin0p99_10Pow10MSun, unpack=True)
[…]

### 2019-09-30 scripts/PlotMathematicaHmemSpinVsMassDensity.py

@@ -1,0 +1,231 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+
+#Load saved memory data from mathematica in geometric units
+
+file_location_hmem = "/home/aschoudhary/gravitational_wave_memory_project/data/dataForPaper/MemoryGrowthInDaysIngral/"
+		
+# Load data for a given input values
+
+def Load_data(Spin, Log_Mass, days):
+
+    if Log_Mass == 8. or Log_Mass ==9. or  Log_Mass ==10. or Log_Mass ==11. or Log_Mass ==12.:
+        Log_Mass = str(Log_Mass).replace('.0', 'p')
+
+    else:
+        Log_Mass = str(Log_Mass).replace('.', 'p')
+
+    if Spin < 0.0:
+        Spin=str(Spin).replace('.', 'p')
+        Spin=str(Spin).replace('-', 'm')
+
+    Spin=str(Spin).replace('.', 'p')
+
+    filename = 'MemoryGrowthIngrlSpin'+Spin+'InDays'+str(days)+'LogMass'+Log_Mass+'.txt'
+
+    time, hmem_Intgrl = np.loadtxt(file_location_hmem+filename, unpack=True)
+    
+    return time, hmem_Intgrl 
+
+# define a funcion that computes the residuals
+
+def Compute_residuals(Spin, Log_Mass, days):
+    time = Load_data(Spin, Log_Mass, days)[0]
+    res = Load_data(Spin, Log_Mass, days)[1]
+
+    res_quadfit = np.polyfit(time, res, 2)
+
+    res_1d = np.poly1d(res_quadfit)
+
+    quadratic_fit_to_res = res_1d(time)
+
+    res_quadSubtract = res-quadratic_fit_to_res
+
+    # mean of reseduals
+    res_mean = np.sqrt(np.mean(res_quadSubtract**2))
+
+
+    return time, res, res_quadSubtract, res_mean
+
+
[…]

### 2019-09-30 scripts/PlotMathematicaHmemdata.py

@@ -1,0 +1,94 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+def find_nearest_idx(array, value):
+    array = np.asarray(array)
+    idx = (np.abs(array - value)).argmin()
+    return idx
+
+
+#Load saved memory data from mathematica in geometric units
+
+file_location_hmemGeoUnits = "/home/aschoudhary/gravitational_wave_memory_project/data/dataForPaper/MemoryGrowthGeoUnit/"
+		
+filename_Spin0p99 = 'datahmemGrowthGeoUnitSpin0p99.txt'
+filename_Spin0p0 = 'datahmemGrowthGeoUnitSpin0p0.txt'
+filename_Spinm0p94 = 'datahmemGrowthGeoUnitSpinm0p94.txt'
+
+time_Spin0p99, hmem_Spin0p99 = np.loadtxt(file_location_hmemGeoUnits+filename_Spin0p99, unpack=True)
+time_Spin0p0, hmem_Spin0p0 = np.loadtxt(file_location_hmemGeoUnits+filename_Spin0p0, unpack=True)
+time_Spinm0p94, hmem_Spinm0p94 = np.loadtxt(file_location_hmemGeoUnits+filename_Spinm0p94, unpack=True)
+
+# Load two weeks data for different mass values
+
+filename_Spin0p99_10Pow8MSun = 'datahmemGrowthGeoUnitSpin0p99TwoWeeks10pow8MSun.txt'
+filename_Spin0p0_10Pow8MSun = 'datahmemGrowthGeoUnitSpin0p0TwoWeeks10pow8MSun.txt'
+filename_Spinm0p94_10Pow8MSun = 'datahmemGrowthGeoUnitSpinm0p94TwoWeeks10pow8MSun.txt'
+
+filename_Spin0p99_10Pow9MSun = 'datahmemGrowthGeoUnitSpin0p99TwoWeeks10pow9MSun.txt'
+filename_Spin0p0_10Pow9MSun = 'datahmemGrowthGeoUnitSpin0p0TwoWeeks10pow9MSun.txt'
[…]

### 2019-09-30 scripts/PlotMathematicaHmemdataIndays.py

@@ -1,0 +1,89 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+def find_nearest_idx(array, value):
+    array = np.asarray(array)
+    idx = (np.abs(array - value)).argmin()
+    return idx
+
+
+#Load saved memory data from mathematica in geometric units
+
+file_location_hmemGeoUnits = "/home/aschoudhary/gravitational_wave_memory_project/data/dataForPaper/MemoryGrowthInDays/"
+		
+
+# Load two weeks data for different mass values
+
+filename_Spin0p99_10Pow8MSun = 'datahmemGrowthInDaysSpin0p99N14Days8MSun.txt'
+filename_Spin0p0_10Pow8MSun = 'datahmemGrowthInDaysSpin0p0N14Days8MSun.txt'
+filename_Spinm0p94_10Pow8MSun = 'datahmemGrowthInDaysSpinm0p94N14Days8MSun.txt'
+
+filename_Spin0p99_10Pow9MSun = 'datahmemGrowthInDaysSpin0p99N14Days9MSun.txt'
+filename_Spin0p0_10Pow9MSun = 'datahmemGrowthInDaysSpin0p0N14Days9MSun.txt'
+filename_Spinm0p94_10Pow9MSun = 'datahmemGrowthInDaysSpinm0p94N14Days9MSun.txt'
+
+filename_Spin0p99_10Pow10MSun = 'datahmemGrowthInDaysSpin0p99N14Days10MSun.txt'
+filename_Spin0p0_10Pow10MSun = 'datahmemGrowthInDaysSpin0p0N14Days10MSun.txt'
+filename_Spinm0p94_10Pow10MSun = 'datahmemGrowthInDaysSpinm0p94N14Days10MSun.txt'
+
+filename_Spin0p99_10Pow11MSun = 'datahmemGrowthInDaysSpin0p99N14Days11MSun.txt'
+filename_Spin0p0_10Pow11MSun = 'datahmemGrowthInDaysSpin0p0N14Days11MSun.txt'
[…]

### 2019-09-30 scripts/PlotMathematicaHmemdataloglog14vs5Year.py

@@ -1,0 +1,222 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import plotsettings
+from matplotlib.font_manager import FontProperties
+
+
+#Load saved memory data from mathematica in geometric units
+
+file_location_hmem = "/home/aschoudhary/gravitational_wave_memory_project/data/dataForPaper/MemoryGrowthInDaysIngral/"
+		
+# Load data for a given input values
+
+def Load_data(Spin, Log_Mass, days):
+
+    if Log_Mass == 8. or Log_Mass ==9. or  Log_Mass ==10. or Log_Mass ==11. or Log_Mass ==12.:
+        Log_Mass = str(Log_Mass).replace('.0', 'p')
+
+    else:
+        Log_Mass = str(Log_Mass).replace('.', 'p')
+
+    if Spin < 0.0:
+        Spin=str(Spin).replace('.', 'p')
+        Spin=str(Spin).replace('-', 'm')
+
+    Spin=str(Spin).replace('.', 'p')
+
+    filename = 'MemoryGrowthIngrlSpin'+Spin+'InDays'+str(days)+'LogMass'+Log_Mass+'.txt'
+
+    time, hmem_Intgrl = np.loadtxt(file_location_hmem+filename, unpack=True)
+    
+    return time, hmem_Intgrl 
+
+# Load two weeks data for different mass values
+
+
+
+
+### Making Plots ########
+legend_size = 5
+fig= plt.figure()
+fontP = FontProperties()
+fontP.set_size('5.0')
+plt.subplots_adjust(left=0.125, right=0.95, bottom=0.1, top=0.95, wspace= 0.2, hspace = 0.2)
+
+plt.subplot(2,2,1)
+
+plt.loglog(Load_data(0.99, 8.0, 14)[0], Load_data(0.99, 8.0, 14)[1], '--', alpha=0.7)
[…]

### 2019-09-30 scripts/ResdualGrowthIntwoWeeks_v3.py

@@ -64,8 +64,22 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
 
 	hmem=17.0*hmem
 
-    #Extend the time period when the memory settlles to make it look like a step function
-   
+        #Increase the resolution which works for 10^12 Msun data
+        #'''
+
+        if log_SolarMass > 10.0:
+
+            dt_NR_intrp=(timeNR[2]-timeNR[1])/100.0
+            timeNR_intrp = np.arange(timeNR[0], timeNR[-1], dt_NR_intrp)
+
+            hmem_intrp = np.interp(timeNR_intrp, timeNR, hmem)
+
+            timeNR = timeNR_intrp
+            hmem = hmem_intrp
+        #'''
+
+        #Extend the time period when the memory settlles to make it look like a step function
+
         timeNR_up=timeNR- timeNR[0]+timeNR[-1]
         timeNR=np.append(timeNR,timeNR_up)
 
@@ -100,23 +114,33 @@ def Memory_growth_in_two_weeks(log_SolarMass, Spin, numer_of_observation_days):
         hmem = np.append(hp_mem_PN[:idx_ti], hmem)
 
 	#Look where the memory growth looks greatest
-		
-	time_final_i = timeNR[-1]
 
-	t_initial_i =  time_initial(log_SolarMass, time_final_i, numer_of_observation_days)
+        if log_SolarMass > 10.0:
+            idx_NRpeak = find_nearest_idx(timeNR, max(hmem)*0.999)
+            time_final_i = timeNR[idx_NRpeak]
+        else:	
+            time_final_i = timeNR[-1]
+	
+        t_initial_i =  time_initial(log_SolarMass, time_final_i, numer_of_observation_days)
 
+        
 	idx_i = find_nearest_idx(timeNR, t_initial_i)
 
 	hmem_two_weeks_i = hmem[idx_i:find_nearest_idx(timeNR, time_final_i)]
 	time_two_weeks_i = timeNR[idx_i:find_nearest_idx(timeNR, time_final_i)]
         
-        '''
+        #'''
         #check if two weeks interval looks correct 
         plt.plot(time_two_weeks_i, hmem_two_weeks_i)
[…]

### 2019-09-30 scripts/h5toASIIdataNospin.py

@@ -12,7 +12,7 @@ filename_list_noSpin = ["rMPsi4_noSpin_q1data"]#, "rMPsi4_noSpin_q1p5data", "rMP
 
 for filename in filename_list_noSpin: 
 
-	file_name_psi="/home/ashok/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/"+filename+".h5"
+	file_name_psi="../data/NonSpinning_differentMassRatio/"+filename+".h5"
 	f_psi = h5py.File(file_name_psi,'r+')
 
 	# Reading the strain data
@@ -54,7 +54,7 @@ for filename in filename_list_noSpin:
 	d2h22_imag_by_dt2_intrp_data= (d2h22_amp_intrp*np.exp(1j*d2h22_phase_intrp)).imag
 
 
-	f = open("/home/ashok/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/"+filename+".txt","w") 
+	f = open("../data/NonSpinning_differentMassRatio/"+filename+".txt","w") 
 
 	for i in range(len(time)):
 		f.write("%E %E %E\n" % (time_intrp[i], d2h22_real_by_dt2_intrp_data[i], d2h22_imag_by_dt2_intrp_data[i])) 

### 2019-09-30 scripts/plotsettings.py

@@ -5,17 +5,17 @@ rc('font', family='serif')
 rc('font', serif='times')
 #rc('font', weight='bolder')
 rc('mathtext', default='sf')
-rc("lines", markeredgewidth=0.6)
-rc("lines", linewidth=2)
-rc('axes', labelsize=14) #24
+rc("lines", markeredgewidth=0.5)
+rc("lines", linewidth=1)
+rc('axes', labelsize=10) #24
 rc("axes", linewidth=0.5) #2)
-rc('xtick', labelsize=15)
-rc('ytick', labelsize=15)
-rc('legend', fontsize=10) #16
+rc('xtick', labelsize=14)
+rc('ytick', labelsize=14)
+rc('legend', fontsize=12) #16
 rc('xtick.major', pad=8) #8)
 rc('ytick.major', pad=8) #8)
-rc('xtick.major', size=13) #8)
-rc('ytick.major', size=13) #8)
+rc('xtick.major', size=8) #8)
+rc('ytick.major', size=8) #8)
 rc('xtick.minor', size=7) #8)
 rc('ytick.minor', size=7) #8)
 

### 2019-09-30 scripts/psi4tohAshok_v2.m

@@ -16,9 +16,9 @@ hintmeth = 3;    %if 1, integrate psi4 in time to 1st order in delta and filter 
 tstart = 100;    %time in M to start, ie length of time at beginning to cut out
 tstop = 50;      %time in M to stop, measured from end
 filename='rMPsi4_noSpin_q1dataClean';
-save_location='/home/ashok/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/Memory_data/';
+save_location='/home/aschoudhary/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/Memory_data/';
 filename_and_loc=strcat(save_location, filename);
-wave = load(strcat(strcat('/home/ashok/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/',filename),'.txt'));
+wave = load(strcat(strcat('/home/aschoudhary/gravitational_wave_memory_project/data/NonSpinning_differentMassRatio/',filename),'.txt'));
 timeNR = wave(:,1);
 psi_plus = wave(:,2);
 psi_cross = wave(:,3);

### 2019-10-01 notes/Paper.tex
Updated paper

@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpin.pdf}
-	\includegraphics[width=3.5in]{../plots/PNmemoryFavata.pdf}
-	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpin.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemorycontributionHpAlginedSpin.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemoryFavata.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemorycontributionHpAntiAlginedSpin.pdf}
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the orbital separation.
@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}
-	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}
+	\includegraphics[width=3.0in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}
 	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation.
@@
 \begin{figure}
- 	\includegraphics[width=3.5in]{../plots/MemoryPlot_nonSpining/q7.pdf}
+ 	\includegraphics[width=3.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
  	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 \begin{figure}
-	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksGeoUnits.pdf}
-	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
-The dashed line shows the complete memory growth.
-The dashed red line show the memory growth computed using PN}
-	\label{fig:resedualGrowth}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow8MSun.pdf}
-	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow10MSun.pdf}
-	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
-The dashed line shows the complete memory growth.
-The dashed red line show the memory growth computed using PN}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksGeoUnits.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow8MSun.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow10MSun.pdf}
+	\caption{The top figure shows the memory growth during merger of a binary with time measured in geometric units with total spin magnitude as indicated above.
+The two weeks interval of time for a given mass in indicated above.
+The above figure show the the memory growth in days.
+The left figure show this growth for $10^{8} M_{\odot}$ and the right figure show the the memory growth in days for $10^{10} M_{\odot}$}
 	\label{fig:resedualGrowth}
@@
 \begin{widetext}
+
 \begin{figure}
-	\includegraphics[width=7.5in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksQuadFitInDays.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays10pow8MSun.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays10pow8MSunQuadSub.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays10pow10MSun.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays10pow10MSunQuadSub.pdf}
 	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
@@
 \begin{figure}
-	\includegraphics[width=7.5in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays.pdf}
+	\includegraphics[width=7.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays.pdf}
 	\caption{The top figure the memory growth during two weeks interval of time with solid lines.

### 2019-10-03 scripts/PlotMathematicaHmemIntgrldataIndays.py
updated scripts with some correction

@@ -33,24 +33,23 @@ legend_size = 5
 fig= plt.figure()
 fontP = FontProperties()
 fontP.set_size('5.0')
-plt.subplots_adjust(left=0.125, right=0.95, bottom=0.1, top=0.95, wspace= 0.2, hspace = 0.2)
+#plt.subplots_adjust(left=0.125, right=0.95, bottom=0.1, top=0.95, wspace= 0.2, hspace = 0.2)
 
-plt.subplot(2,2,1)
+#plt.subplot(2,2,1)
 
-plt.plot(time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$')
-plt.plot(time_Spin0p0_10Pow8MSun, hmem_Spin0p0_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$')
-plt.plot(time_Spinm0p94_10Pow8MSun, hmem_Spinm0p94_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$')
+plt.plot(time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$', linewidth=2)
+plt.plot(time_Spin0p0_10Pow8MSun, hmem_Spin0p0_10Pow8MSun, '--',alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$', linewidth=2)
+plt.plot(time_Spinm0p94_10Pow8MSun, hmem_Spinm0p94_10Pow8MSun, '-.', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$', linewidth=2)
 
-plt.xlabel(r'$t$(days)')
-plt.ylabel(r'$Residuals$')
-plt.xticks([0, 5, 10, 15])
-#plt.yticks([2.5, 5])
-#plt.ylim(0, 0.1)
-plt.legend(loc='best', prop={'size':legend_size})
[…]

### 2019-10-03 scripts/PlotMathematicaHmemSpinVsMassDensity.py

@@ -60,110 +60,133 @@ legend_size = 5
 fig= plt.figure()
 fontP = FontProperties()
 fontP.set_size('5.0')
-plt.subplots_adjust(left=0.125, right=0.95, bottom=0.1, top=0.95, wspace= 0.2, hspace = 0.2)
+#plt.subplots_adjust(left=0.1, right=0.95, bottom=0.1, top=0.95, wspace= 0.2, hspace = 0.1)
 
-plt.subplot(2,2,1)
+#plt.subplot(2,2,1)
 
 plt.loglog(Load_data(0.99, 8.0, 14)[0], Load_data(0.99, 8.0, 14)[1], 'c--', alpha=0.7)
-plt.loglog(Load_data(0.0, 8.0, 14)[0], Load_data(0.0, 8.0, 14)[1], 'c',alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$')
+plt.loglog(Load_data(0.0, 8.0, 14)[0], Load_data(0.0, 8.0, 14)[1], 'c',alpha=0.7, label=r'$M=10 ^{8}\ M_\odot \ $')
 plt.loglog(Load_data(-0.94, 8.0, 14)[0], Load_data(-0.94, 8.0, 14)[1], 'c-.', alpha=0.7)
 
 plt.loglog(Load_data(0.99, 9.090910, 14)[0], Load_data(0.99, 9.090910, 14)[1], 'm--', alpha=0.7)
-plt.loglog(Load_data(0.0, 9.090910, 14)[0], Load_data(0.0, 9.090910, 14)[1],'m' ,alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$')
+plt.loglog(Load_data(0.0, 9.090910, 14)[0], Load_data(0.0, 9.090910, 14)[1],'m' ,alpha=0.7, label=r'$M=10 ^{9}\ M_\odot \ $')
 plt.loglog(Load_data(-0.94, 9.090910, 14)[0], Load_data(-0.94, 9.090910, 14)[1], 'm-.', alpha=0.7)
 
 
-plt.loglog(Load_data(0.99, 12.0, 14)[0], Load_data(0.99, 10.0, 14)[1], 'k--', alpha=0.7)
[…]

### 2019-10-03 scripts/PlotMathematicaHmemdata.py

@@ -62,12 +62,12 @@ fontP = FontProperties()
 
 plt.semilogy(time_Spin0p99, hmem_Spin0p99, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$')
 #plt.semilogy(time_Spin0p0, hmem_Spin0p0, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$')
-plt.semilogy(time_Spinm0p94, hmem_Spinm0p94, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$')
+plt.semilogy(time_Spinm0p94, hmem_Spinm0p94, '--', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$')
 
 
 plt.scatter([time_Spin0p99_10Pow8MSun[0], time_Spin0p99_10Pow8MSun[-1]], [hmem_Spin0p99_10Pow8MSun[0], hmem_Spin0p99_10Pow8MSun[-1]], marker='x', color='red', label=r'$M=10 ^{8}\ M_\odot \ $')
 plt.scatter([time_Spin0p99_10Pow9MSun[0], time_Spin0p99_10Pow9MSun[-1]], [hmem_Spin0p99_10Pow9MSun[0], hmem_Spin0p99_10Pow9MSun[-1]], marker='*', color='black', label=r'$M=10 ^{9}\ M_\odot \ $')
-plt.scatter([time_Spin0p99_10Pow10MSun[0], time_Spin0p99_10Pow10MSun[-1]], [hmem_Spin0p99_10Pow10MSun[0], hmem_Spin0p99_10Pow10MSun[-1]], marker='.', color='red', label=r'$M=10 ^{10}\ M_\odot \ $')
+plt.scatter([time_Spin0p99_10Pow11MSun[0], time_Spin0p99_10Pow11MSun[-1]], [hmem_Spin0p99_10Pow11MSun[0], hmem_Spin0p99_10Pow11MSun[-1]], marker='.', color='red', label=r'$M=10 ^{11}\ M_\odot \ $')
 
[…]

### 2019-10-03 scripts/PlotMathematicaHmemdataIndays.py

@@ -49,16 +49,19 @@ time_Spin0p99_10Pow11MSun, hmem_Spin0p99_10Pow11MSun = np.loadtxt(file_location_
 time_Spin0p0_10Pow11MSun, hmem_Spin0p0_10Pow11MSun = np.loadtxt(file_location_hmemGeoUnits+filename_Spin0p0_10Pow11MSun, unpack=True)
 time_Spinm0p94_10Pow11MSun, hmem_Spinm0p94_10Pow11MSun = np.loadtxt(file_location_hmemGeoUnits+filename_Spinm0p94_10Pow11MSun, unpack=True)
 
+#Making plots
 
 fig= plt.figure()
 fontP = FontProperties()
+fontP.set_size('20.')
+legend_size = 2
 
-plt.plot(time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$')
-plt.plot(time_Spin0p0_10Pow8MSun, hmem_Spin0p0_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$')
-plt.plot(time_Spinm0p94_10Pow8MSun, hmem_Spinm0p94_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$')
+plt.plot(time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$', linewidth=2)
+plt.plot(time_Spin0p0_10Pow8MSun, hmem_Spin0p0_10Pow8MSun, '--', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$', linewidth=2)
+plt.plot(time_Spinm0p94_10Pow8MSun, hmem_Spinm0p94_10Pow8MSun,'-.', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$', linewidth=2)
 
-plt.xlabel(r'$t(days)$')
[…]

### 2019-10-03 scripts/plotsettings.py

@@ -11,7 +11,7 @@ rc('axes', labelsize=10) #24
 rc("axes", linewidth=0.5) #2)
 rc('xtick', labelsize=14)
 rc('ytick', labelsize=14)
-rc('legend', fontsize=12) #16
+rc('legend', fontsize=16) #16
 rc('xtick.major', pad=8) #8)
 rc('ytick.major', pad=8) #8)
 rc('xtick.major', size=8) #8)

### 2019-10-03 notes/Paper.tex
made some changes in the notes

@@
 
-\title{Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers and it's implication on Observing Gravitational Wave Bursts With Memory in Pulsar Timing Array}
+\title{Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers and its implication on Observing Gravitational Wave Bursts With Memory in Pulsar Timing Array}
 \author{Ashok Choudhary}\email{aschoudhary@mix.wvu.edu}
@@
 \begin{figure}
-	\includegraphics[width=7.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDaysPreFit.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDaysPostfit.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/ResidualGrowth5YearsDiffSpinsInDaysPrefit.pdf}
+	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/ResidualGrowth5YearsDiffSpinsInDaysPostfit.pdf}
 	\caption{The top figure the memory growth during two weeks interval of time with solid lines.

### 2019-10-08 notes/Paper.tex
updated notes

@@
 	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/ResidualGrowth2weeksDiffSpinsInDays10pow10MSunQuadSub.pdf}
-	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
-The dashed line shows the complete memory growth.
-The dashed red line show the memory growth computed using PN}
+	\caption{The figure above shows residual growth in a given interval of time.
+The top left plot shows this residual growth during two weeks interval of time for different spin parameter values and the top right plot shows the residual growth after subtracting a quadratic fit for maximally spinning case and a binary with $M = 10^{8}\, M_{\odot}$.
+The bottom figure shows the similar plot for a binary with $M = 10^{11}\, M_{\odot}$.}
 	\label{fig:resedualG}
@@
 	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/ResidualGrowth5YearsDiffSpinsInDaysPostfit.pdf}
-	\caption{The top figure the memory growth during two weeks interval of time with solid lines.
-The dashed line shows the complete memory growth.
-The dashed red line show the memory growth computed using PN}
+	\caption{The figure above shows residual growth in a given interval of time.
+The top left plot shows this residual growth during two weeks interval of time and the top right plot shows the residual growth after subtracting a quadratic fit.
+The bottom figure shows the similar plot for 5 years interval of time.}
 	\label{fig:resedualGrowth}
@@
 	\includegraphics[width=3.5in]{../plots/PlotfromMathematicaData/MemoryResSpinvsMass1825Days.pdf}
-	\caption{The top figure shows the Christodolou memory growth during two weeks of observation time as a function of Mass and Spin magnitude $\chi$.
-The here in M time the mass of Sun.
-The bottom plot show a the rms timing residuals during two week period of time.}
+	\caption{The top figure shows the root mean square of post-fit timing residuals growth as a function of Mass and Spin magnitude $\chi$.
+The left plot shows this post-fit residual fit during two weeks interval of time and right plot shows this for 5 years interval.}
 	\label{fig:MemoryMassVsSpin}

### 2019-10-09 scripts/PlotMathematicaHmemIntgrldataIndays.py
updated scripts

@@ -70,9 +70,9 @@ fontP.set_size('5.0')
 
 #plt.subplot(2,2,2)
 
-plt.plot(time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun,'c', label='Residual', linewidth=2)
+plt.plot(time_Spin0p99_10Pow8MSun, hmem_Spin0p99_10Pow8MSun,'c', label='Pre-fit residual', linewidth=2)
 plt.plot(time_Spin0p99_10Pow8MSun, quadratic_fit_to_res0p99_10pow8, 'r--', label='Quadratic fit', linewidth=2)
-plt.plot(time_Spin0p99_10Pow8MSun, res0p99_10pow8_quadSubtract, 'm-.', label='Residual post quadratric subtraction', linewidth=2)
+plt.plot(time_Spin0p99_10Pow8MSun, res0p99_10pow8_quadSubtract, 'm-.', label='Post-fit residual', linewidth=2)
 
 plt.xlabel(r'$t$(days)', fontsize=18)
 plt.ylabel(r'$Residuals$', fontsize=18)
@@ -129,9 +129,9 @@ fontP.set_size('5.0')
 
 #plt.subplot(2,2,4)
 
-plt.plot(time_Spin0p99_10Pow10MSun, hmem_Spin0p99_10Pow10MSun, 'c', label='Residual', linewidth=2)
+plt.plot(time_Spin0p99_10Pow10MSun, hmem_Spin0p99_10Pow10MSun, 'c', label=' Pre-fit residual', linewidth=2)
 plt.plot(time_Spin0p99_10Pow10MSun, quadratic_fit_to_res0p99_10pow10, 'r--', label='Quadratic fit', linewidth=2)
-plt.plot(time_Spin0p99_10Pow10MSun, res0p99_10pow10_quadSubtract, 'm-.', label='Residual post quadratric subtraction', linewidth=2)
+plt.plot(time_Spin0p99_10Pow10MSun, res0p99_10pow10_quadSubtract, 'm-.', label='Post-fit residual ', linewidth=2)
 plt.xlabel(r'$t$(days)', fontsize=18)
 plt.ylabel(r'$Residuals$', fontsize=18)
 

### 2019-10-09 scripts/PlotMathematicaHmemSpinVsMassDensity.py

@@ -49,7 +49,7 @@ def Compute_residuals(Spin, Log_Mass, days):
     res_mean = np.sqrt(np.mean(res_quadSubtract**2))
 
 
-    return time, res, res_quadSubtract, res_mean
+    return time, res, res_quadSubtract, res_mean, quadratic_fit_to_res
 
 
 def Residual_growth14Days(Spin, Log_SolarMass):
@@ -80,7 +80,7 @@ plt.loglog(Load_data(-0.94, 11.0909, 14)[0], Load_data(-0.94, 11.0909, 14)[1], '
 
 
 plt.xlabel(r'$t$(days)', fontsize=18)
-plt.ylabel(r'$Residuals \, \ prefit$', fontsize=18)
+plt.ylabel(r'$prefit \, \ Residuals$', fontsize=18)
 #plt.xticks([0, 5, 10, 15])
 plt.yticks([pow(10, -24), pow(10, -18), pow(10, -12)])
 #plt.ylim(pow(10, -24), pow(10, -8))
@@ -111,7 +111,7 @@ plt.loglog(Load_data(-0.94, 11.0909, 1825)[0], Load_data(-0.94, 11.0909, 1825)[1
 
 
 plt.xlabel(r'$t$(days)', fontsize=18)
-plt.ylabel(r'$Residuals \, \ postfit$', fontsize=18)
+plt.ylabel(r'$postfit \, \ Residuals$', fontsize=18)
 #plt.xticks([0, 5, 10, 15])
 plt.yticks([pow(10, -19), pow(10, -16), pow(10, -13), pow(10, -10)])
 #plt.ylim(pow(10, -24), pow(10, -8))
@@ -147,7 +147,7 @@ plt.loglog(Compute_residuals(0.0, 12.0, 14)[0], abs(Compute_residuals(-0.94, 12.
 
 
 plt.xlabel(r'$t$(days)', fontsize=18)
-plt.ylabel(r'$Residuals \, \ prefit$', fontsize=18)
+plt.ylabel(r'$prefit \, \ Residuals$', fontsize=18)
 #plt.xticks([0, 5, 10, 15])
 plt.yticks([pow(10, -22), pow(10, -19), pow(10, -16)])
 plt.legend(loc='best', prop={'size':15})
@@ -180,7 +180,7 @@ plt.loglog(Compute_residuals(0.0, 12.0, 1825)[0], abs(Compute_residuals(-0.94, 1
 
 
[…]

### 2019-10-09 scripts/PlotMathematicaHmemdata.py

@@ -67,6 +67,7 @@ plt.semilogy(time_Spinm0p94, hmem_Spinm0p94, '--', alpha=0.7, label=r'$\mathbf{\
 
 plt.scatter([time_Spin0p99_10Pow8MSun[0], time_Spin0p99_10Pow8MSun[-1]], [hmem_Spin0p99_10Pow8MSun[0], hmem_Spin0p99_10Pow8MSun[-1]], marker='x', color='red', label=r'$M=10 ^{8}\ M_\odot \ $')
 plt.scatter([time_Spin0p99_10Pow9MSun[0], time_Spin0p99_10Pow9MSun[-1]], [hmem_Spin0p99_10Pow9MSun[0], hmem_Spin0p99_10Pow9MSun[-1]], marker='*', color='black', label=r'$M=10 ^{9}\ M_\odot \ $')
+plt.scatter([time_Spin0p99_10Pow10MSun[0], time_Spin0p99_10Pow10MSun[-1]], [hmem_Spin0p99_10Pow10MSun[0], hmem_Spin0p99_10Pow10MSun[-1]], marker='+', color='black', label=r'$M=10 ^{10}\ M_\odot \ $')
 plt.scatter([time_Spin0p99_10Pow11MSun[0], time_Spin0p99_10Pow11MSun[-1]], [hmem_Spin0p99_10Pow11MSun[0], hmem_Spin0p99_10Pow11MSun[-1]], marker='.', color='red', label=r'$M=10 ^{11}\ M_\odot \ $')
 
 #plt.scatter([time_Spin0p0_10Pow8MSun[0], time_Spin0p0_10Pow8MSun[-1]], [hmem_Spin0p0_10Pow8MSun[0], hmem_Spin0p0_10Pow8MSun[-1]], marker='x', color='red')
@@ -75,6 +76,7 @@ plt.scatter([time_Spin0p99_10Pow11MSun[0], time_Spin0p99_10Pow11MSun[-1]], [hmem
 
 plt.scatter([time_Spinm0p94_10Pow8MSun[0], time_Spinm0p94_10Pow8MSun[-1]], [hmem_Spinm0p94_10Pow8MSun[0], hmem_Spinm0p94_10Pow8MSun[-1]], marker='x', color='red')
[…]

### 2019-10-09 notes/Paper.tex
Updated plots in the paper

@@
 Nonlinear memory from the merger of super massive black hole binary can be potentially observed in pulsar timing measurement.
-Gravitation wave, when they pass through the space between earth and pulsar effect the time of arrival of the radio pulses at earth.
+Gravitational waves, when they pass through the space between earth and pulsar effect the time of arrival of the radio pulses at earth.
 The modulation in the observed frequency is given by
@@
 We show a figure similar to
-\begin{widetext}
+%%\begin{widetext}
 \begin{figure}
@@
 	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow8MSun.pdf}
-	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow10MSun.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow9MSun.pdf}
+	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow11MSun.pdf}
 	\caption{The top figure shows the memory growth during merger of a binary with time measured in geometric units with total spin magnitude as indicated above.
-The two weeks interval of time for a given mass in indicated above.
-The above figure show the the memory growth in days.
-The left figure show this growth for $10^{8} M_{\odot}$ and the right figure show the the memory growth in days for $10^{10} M_{\odot}$}
+The two weeks interval of time for a given mass as indicated above.
+The bottom three figures show the the memory growth in days for Mass $10^{8}M_{\odot}$, $10^{9}M_{\odot}$ and $10^{11}M_{\odot}$ respectively.}
 	\label{fig:resedualGrowth}
 \end{figure} 
-\end{widetext}
+%\end{widetext}
 

### 2019-10-09 notes/Paper.tex
updated paper

@@
 	applying that method using publicly available numerical relativity simulations of spinning binaries.
-We Finally use this model to compute the timing residuals for a memory growth during two weeks interval of time following the method described in \cite{Pshirkov2010}
+We finally use this model to compute the timing residuals for a memory growth during two weeks interval of time following the method described in \cite{Pshirkov2010}.
+We find that when the burst with memory( BWM) signals are not assumed to a step function and replaced with the actual memory growth functions can significantly chage the expected results.
+Our results shows that spin contribution plays an important role while computing the residuals.
 \end{abstract}
@@
 \end{equation}
-
 The pulsar timing measurement primarily measures the difference between the actual pulse arrival time and the time predicted from the spin-down model of a pulsar, which is called the timing residuals.
@@
 \end{equation}
-There are two unknown quantities associated with the timing residuals, the pulsar periods and it's spin down rate. We subtract these two quantities from the expression $s(t)$,
+There are two unknown quantities associated with the timing residuals, the pulsar periods and it's spin down rate. We subtract these two quantities from the expression $s(t)$.
 
@@
 \par 
-Future observations with increased precision would be able to look at the memory growth on even shorter time of few weeks.
+Future observations with increased precision would be able to look at the memory growth on even shorter time of two weeks.
 The memory growth on the time scale of two weeks does not look like a step function, instead it's shape depends upon the spins of the merging black holes.
-In Fig[\ref{fig:resedualGrowth}].we show the memory growth interval in two weeks interval of time with solid lines for different range of spin and total mass of the binary.
-The dashed line show the complete memory growth in growth in geometrical units.
+In Fig[\ref{fig:MemoryGrowthTwoweeks}].we show the memory growth during two weeks interval of time with markers on complete memory during for different range of spin and total mass of the binary.
+The solid line show the complete memory growth in growth in geometrical units.
 It can be seen the as the total mass of binary becomes larger the two weeks interval of time shrinks in geometrical units.
@@
 For the pulsar timing observation we compute the residuals during this two week interval of time following the method described in \cite{Pshirkov2010}.
-We show a figure similar to
+It can be seen in Fig[\ref{fig:MemoryGrowthTwoweeks}] that the memory growth during this two weeks interval of time look very for different mass values.
+For higher masses ($10^{10}M_{\odot}$), this looks like a straight line.
+For lower masses this looks like a nonlinear growth.
+In Fig[\ref{fig:resedualG}] we show the residual before(pre-fit) and after  (post-fit) a quadratic subtraction.
+It can be seen that the residual growth fits better to a pre-fit residuals
 %%\begin{widetext}
@@
 The bottom three figures show the the memory growth in days for Mass $10^{8}M_{\odot}$, $10^{9}M_{\odot}$ and $10^{11}M_{\odot}$ respectively.}
-	\label{fig:resedualGrowth}
+	\label{fig:resedualG}
 \end{figure} 

### 2019-10-09 notes/Paper.tex
updated paper

@@
 For lower masses this looks like a nonlinear growth.
-In Fig[\ref{fig:resedualG}] we show the residual before(pre-fit) and after  (post-fit) a quadratic subtraction.
-It can be seen that the residual growth fits better to a pre-fit residuals
+In Fig[\ref{fig:resedualGrowthTwoweeks}] we show the residual before(pre-fit) and after  (post-fit) a quadratic subtraction.
+It can be seen that the residual growth fits better to a pre-fit residuals for a higher mass supermassive black hole ($M=10^{11}M_{\odot}$), which obvious because when the memory growth is linear, it integral would give a quadratic function.
+However the memory growth for black holes with $M = 10^{9}M_{\odot}$ looks very  nonlinear, hence the resulting integral also looks like a function different from a quadratic polynomial.
+This makes the timing residual post quadratic subtraction depend upon the mass of super massive binary.
+Fig[\ref*{fig:resedualGrowthLogLog}] shows a log-log plot for pre-fit and post-fit residuals for different ranges of mass and spin parameters.
+It can be seen the residual growth amplitude when evaluated for a two weeks interval tends to increase initially with mass, reaches a maximum when $M=10^{9}M_{\odot}$ and then again decreases.
+This shows that when observed over a two weeks interval of time super massive black holes above a mass of $10^{10}M_{\odot}$ would be harder to detect.
+Fig[\ref*{fig:MemoryMassVsSpin}] show color plot for rms residuals for a range of spin and mass values.
 %%\begin{widetext}
@@
 The bottom three figures show the the memory growth in days for Mass $10^{8}M_{\odot}$, $10^{9}M_{\odot}$ and $10^{11}M_{\odot}$ respectively.}
-	\label{fig:resedualG}
+	\label{fig:MemoryGrowthTwoweeks}
 \end{figure} 
@@
 The bottom figure shows the similar plot for a binary with $M = 10^{11}\, M_{\odot}$.}
-	\label{fig:resedualG}
+	\label{fig:resedualGrowthTwoweeks}
 \end{figure}
@@
 The bottom figure shows the similar plot for 5 years interval of time.}
-	\label{fig:resedualGrowth}
+	\label{fig:resedualGrowthLogLog}
 \end{figure}
@@
 Observation of Christodoulou memory would be an important test of general relativity in strong field regime.
-Effort to detect this memory effect in low frequency band from super massive black holes have been undertaken by pulsar timing array experiments \cite{Pshirkov2010}.
-In these experiment the spin effect have been neglected for modeling the signal.
+Effort to detect this memory effect in low frequency band from super massive black holes have been undertaken by pulsar timing array experiments \cite{Arzoumanian2015}.
+The author in their analysis have looked for memory in a observation data for 5 years.
+In their analysis they have assumed the memory to a step function for a given amplitude.
+This amplitude is of the order $10^{-15}$, which is memory jump for a $10^{8}M_{\odot}$ binary.
+In these analysis the spin effect have been neglected for modeling the signal.
 It is highly likely that super massive black holes have significant spin due to accretion of matter from surrounding.
 We have seen that the spin can significantly change memory amplitude as a result it can change the timing residuals.
-Hence it is important to consider the spin effect while modeling the signals.
+In our analysis we observe that:(1) Growth of memory when looks over a shorter observing time scale does not look like a step function, instead it's shape depends upon the spin and mass of binary for a give interval of time.(2) The post-fit residual for a observation period of two weeks  .Hence it is important to consider the spin effect while modeling the signals.
 The important thing for pulsar timing measurements would be the amount of memory changes over a single two weeks period of time which is shown the figure \ref{fig:MemoryMassVsSpin} as function of total mass of the binary and the spin magnitude.

### 2019-10-09 notes/Paper.tex
updated paper

@@
 We use PN expressions for memory for early time and for late time we use memory extracted from NR simulations.
+In Sec.
+V we look at how these results can effect expected results in effort to detect the BWM from super massive binaries in pulsar timing array.
+We look a the timing residuals growth over a period of two weeks interval of time.
 
@@
 We have seen that the spin can significantly change memory amplitude as a result it can change the timing residuals.
-In our analysis we observe that:(1) Growth of memory when looks over a shorter observing time scale does not look like a step function, instead it's shape depends upon the spin and mass of binary for a give interval of time.(2) The post-fit residual for a observation period of two weeks  .Hence it is important to consider the spin effect while modeling the signals.
-The important thing for pulsar timing measurements would be the amount of memory changes over a single two weeks period of time which is shown the figure \ref{fig:MemoryMassVsSpin} as function of total mass of the binary and the spin magnitude.
+In our analysis we observe that:(1) Growth of memory when looked over a shorter observing time scale does not look like a step function, instead it's shape depends upon the spin and mass of binary for a give interval of time.(2) The post-fit rms residual for a observation period of two weeks reach a maximum for $10^{9}M_{\odot}$ and tend to decrease higher mass case.
+Hence it is important to consider the spin effect while modeling the BWM signals and consider a range of mass values.
+Another important thing for pulsar timing measurements would be the amount of memory changes over a single two weeks period of time which is shown the figure \ref{fig:MemoryMassVsSpin} as function of total mass of the binary and the spin magnitude.
 For a given mass and over the allowed spin range the net change in the memory amplitude can be as big as a factor of 5 which would change the SNR, or the expected event rate within a given volume, by a factor of $5^{3}=125$

### 2019-10-09 scripts/PlotMathematicaHmemSpinVsMassDensity.py
updated scripts for density plots

@@ -79,8 +79,8 @@ plt.loglog(Load_data(-0.94, 11.0909, 14)[0], Load_data(-0.94, 11.0909, 14)[1], '
 
 
 
-plt.xlabel(r'$t$(days)', fontsize=18)
-plt.ylabel(r'$prefit \, \ Residuals$', fontsize=18)
+plt.xlabel('t(days)', fontsize=18)
+plt.ylabel('pre-fit Residuals', fontsize=18)
 #plt.xticks([0, 5, 10, 15])
 plt.yticks([pow(10, -24), pow(10, -18), pow(10, -12)])
 #plt.ylim(pow(10, -24), pow(10, -8))
@@ -110,8 +110,8 @@ plt.loglog(Load_data(-0.94, 11.0909, 1825)[0], Load_data(-0.94, 11.0909, 1825)[1
 
 
 
-plt.xlabel(r'$t$(days)', fontsize=18)
-plt.ylabel(r'$postfit \, \ Residuals$', fontsize=18)
+plt.xlabel('t(days)', fontsize=18)
+plt.ylabel('post-fit Residuals', fontsize=18)
 #plt.xticks([0, 5, 10, 15])
 plt.yticks([pow(10, -19), pow(10, -16), pow(10, -13), pow(10, -10)])
 #plt.ylim(pow(10, -24), pow(10, -8))
@@ -146,8 +146,8 @@ plt.loglog(Compute_residuals(0.0, 11.0909, 14)[0], abs(Compute_residuals(-0.94, 
 plt.loglog(Compute_residuals(0.0, 12.0, 14)[0], abs(Compute_residuals(-0.94, 12.0, 14)[2]),'r', label=r'$M=10 ^{12}\ M_\odot \ $')
 
 
-plt.xlabel(r'$t$(days)', fontsize=18)
-plt.ylabel(r'$prefit \, \ Residuals$', fontsize=18)
+plt.xlabel('t(days)', fontsize=18)
+plt.ylabel('pre-fit Residuals', fontsize=18)
 #plt.xticks([0, 5, 10, 15])
 plt.yticks([pow(10, -22), pow(10, -19), pow(10, -16)])
 plt.legend(loc='best', prop={'size':15})
@@ -179,8 +179,8 @@ plt.loglog(Compute_residuals(0.0, 11.0909, 1825)[0], abs(Compute_residuals(-0.94
[…]

### 2019-10-10 scripts/MakeReseddualPlot_from_MathematicaData.py
updated script with new plots

@@ -1,15 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.cm as cm
-import plotsettings
-from matplotlib.font_manager import FontProperties
-
-file_location_hmem_Ingrl ='../data/dataForPaper/MemoryGrowthInDaysIngral/'
-datafile_hMem_Ingrl = "MemoryGrowthIngrlSpinm0p9InDays1825LogMass9p81818.txt"
-
-time_InDays, hMem_Ingrl = np.loadtxt(file_location_hmem_Ingrl+datafile_hMem_Ingrl, unpack=True)
-
-plt.plot(time_InDays, hMem_Ingrl)
-plt.show()
-
-

### 2019-10-10 scripts/Make_memoryCalculationplots_nonSpining_v4.py

@@ -127,82 +127,21 @@ def find_nearest_idx(array, value):
 
 #data location
 file_location ='../data/NonSpinning_differentMassRatio/Memory_data/'
-#import data
-mass_ratio_vec = [1, 1.5, 2, 2.5, 3, 4, 5, 6, 7, 8, 9.5]
-filename_vec=['q1', 'q1p5', 'q2', 'q2p5','q3','q4','q5','q6', 'q7', 'q8','q9p5']
-i=0
-
-for filename in filename_vec:
-	
-	datafile_hNRdot='rMPsi4_noSpin_'+filename+'dataClean_hdotNR.dat'
-	datafile_hNR='rMPsi4_noSpin_'+filename+'dataClean_hNR.dat'
-	datafile_hMemNR='rMPsi4_noSpin_'+filename+'dataClean_hMemNR.dat'
-
-	timeNR, hmem, h_mem_plus = np.loadtxt(file_location+datafile_hMemNR, unpack=True)
-	timeNR, hdot_plus, hdot_cross = np.loadtxt(file_location+datafile_hNRdot, unpack=True)
-	timeNR, h_plus, h_cross = np.loadtxt(file_location+datafile_hNR, unpack=True)
-
-
-	#Making plots
-	legend_size = 1
-	fig = plt.figure()
-	plt.suptitle('Memory Calculation for q ='+str(mass_ratio_vec[i]) ,fontsize = 15)
-	fontP = FontProperties()
-	fontP.set_size('10.')
-
-
-	legend = plt.legend(loc='best',prop={'size':legend_size})
-	plt.subplot(2,2,1)
-	plt.plot(timeNR, hdot_plus, 'r', label=r'$hdot_{+}$')
-	plt.plot(timeNR, hdot_cross, 'k--', label=r'$hdot_{\times}$')
-	plt.grid()
-	plt.xlabel(r'$time$')
-	plt.ylabel(r'$hdot_{+,\times}$')
-	plt.legend(loc=2)
-	fontP.set_size('10.')
-
-	plt.subplot(2,1,2)
-	plt.plot(timeNR, h_plus, 'r', label=r'$h_{+}$')
-	plt.plot(timeNR, h_cross, 'k--', label=r'$h_{\times}$')
-	plt.grid()
-	plt.xlabel(r'$time$')
[…]

### 2019-10-10 scripts/PlotMathematicaHmemSpinVsMassDensity.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 import plotsettings
 from matplotlib.font_manager import FontProperties
-
+from scipy.signal import savgol_filter as sv
 
 #Load saved memory data from mathematica in geometric units
 
@@ -36,6 +36,7 @@ def Load_data(Spin, Log_Mass, days):
 def Compute_residuals(Spin, Log_Mass, days):
     time = Load_data(Spin, Log_Mass, days)[0]
     res = Load_data(Spin, Log_Mass, days)[1]
+    res = sv(res, 51, 3)
 
     res_quadfit = np.polyfit(time, res, 2)
 
@@ -235,9 +236,9 @@ fontP.set_size('5.0')
 
 #fig, ax = plt.subplots()
 #plt.subplot(1,2,1)
-im = plt.imshow(Reseduals_M_vs_Spin14days , interpolation='bicubic', cmap=cm.plasma, extent=[8, 11.0, -0.94, 0.99], vmin=vmn14days, vmax=vmx14days)
+im = plt.imshow(Reseduals_M_vs_Spin14days , interpolation='quadric', cmap=cm.plasma, extent=[8, 11.0, -0.94, 0.99], vmin=vmn14days, vmax=vmx14days)
 
-plt.xlabel(r'$Log10{(M)}$', fontsize=18)
+plt.xlabel(r'$\log10{(M)}$', fontsize=18)
 plt.ylabel(r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N$', fontsize=18)
 plt.colorbar()
 fig.tight_layout()
@@ -252,7 +253,7 @@ fontP.set_size('5.0')
 im = plt.imshow(Reseduals_M_vs_Spin1825days , interpolation='bicubic', cmap=cm.plasma, extent=[8, 11.0, -0.94, 0.99], vmin=vmn1825days, vmax=vmx1825days)
 
 
-plt.xlabel(r'$Log10{(M)}$', fontsize=18)
+plt.xlabel(r'$\log10{(M)}$', fontsize=18)
 plt.ylabel(r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N$', fontsize=18)
 plt.colorbar()
 fig.tight_layout()

### 2019-10-10 scripts/PlotMathematicaHmemdataIndays.py

@@ -75,6 +75,27 @@ fontP = FontProperties()
 fontP.set_size('20.')
 legend_size = 2
 
+plt.plot(time_Spin0p99_10Pow9MSun, hmem_Spin0p99_10Pow9MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$', linewidth=2)
+plt.plot(time_Spin0p0_10Pow9MSun, hmem_Spin0p0_10Pow9MSun, '--', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$', linewidth=2)
+plt.plot(time_Spinm0p94_10Pow9MSun, hmem_Spinm0p94_10Pow9MSun, '-.', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$', linewidth=2)
+
+
+plt.xlabel(r'$t(days)$', fontsize=18)
+plt.ylabel(r'$(1Gpc/D_{L})h^{(mem)}_{+}$', fontsize=18)
+#plt.xticks([-3000, -2000, -1000, 0])
+#plt.ylim(0, 0.1)
+plt.legend()
+fig.tight_layout()
+
+plt.savefig('/home/aschoudhary/gravitational_wave_memory_project/plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow9MSun.pdf')
+plt.show()
+
+
+fig= plt.figure()
+fontP = FontProperties()
+fontP.set_size('20.')
+legend_size = 2
+
 plt.plot(time_Spin0p99_10Pow11MSun, hmem_Spin0p99_10Pow11MSun, alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.99$', linewidth=2)
 plt.plot(time_Spin0p0_10Pow11MSun, hmem_Spin0p0_10Pow11MSun, '--', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =0.0$', linewidth=2)
 plt.plot(time_Spinm0p94_10Pow11MSun, hmem_Spinm0p94_10Pow11MSun, '-.', alpha=0.7, label=r'$\mathbf{\chi}_{s} \cdot \hat{\mathbf{L}}_N =-0.94$', linewidth=2)
[…]

### 2019-10-10 notes/Paper.tex
Updated notes

@@
 These massive stars are likely to posses spin angular momentum, as a result the black hole formed would also be spinning.
-The black holes can also accrete matter from surrounding and as a result spin up during the process.
-It is highly likely that there is a large population of spinning binaries in the universe.
+The super black holes found at the galactic center can acquire significant spins due also accretion of matter from surrounding.
+It is highly likely that there is a large population of spinning super massive binaries in the universe.
 \par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.
-For PN calculations, this paper is just an extension of Favata's\cite{Favata2009} paper, where we include the spin contribution up to leading (1.5 PN) order for nonprecessing binaries.
+For PN calculations, we extend the calculations  done in Favata's\cite{Favata2009} paper, where we include the spin contribution up to leading (1.5 PN) order for nonprecessing binaries.
 The nonlinear memory contribution near merger is computed using NR data from publicly available SXS catalog, where we process the data to remove the low frequency numerical noise.
@@
 \begin{figure}
- 	\includegraphics[width=3.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
+ 	\includegraphics[width=3.5in]{../plots/MemoryPlot_nonSpining/q7.pdf}
  	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
@@
 \begin{figure}
-	\includegraphics[width=3.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
-	\includegraphics[width=3.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
+	\includegraphics[width=3.5in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce.
